### 2013-01-06 paper/ms.tex
Uploaded all of the code and the paper. Only some of the data.

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ Letters}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
\bibliographystyle{apj}
\usepackage{graphics}
\usepackage[dvips]{epsfig}

\newcommand{\kms}{\,km~s$^{-1}$}
\def\squig{\sim\!\!}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\Lsun}{\mbox{\,$L_{\odot}$}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\bftheta}{\mathbf{\Theta}}
\newcommand{\degree}{\ensuremath{^\circ}}
\def\spose#1{\hbox to 0pt{#1\hss}}
\def\simlt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13C$}}}
\def\simgt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13E$}}}
\font\smcap=cmcsc10
\def\caii{Ca\,{\smcap ii}}

\shorttitle{The not so common kinematics of the Local Group}
\shortauthors{Forero-Romero et al.}

\begin{document}
\title{The not so common kinematics of the Local Group}
\author{
J. E.\ Forero-Romero\altaffilmark{1}, 
Y. Hoffman\altaffilmark{2}, 
S.  Bustamante\altaffilmark{3}, 
S. Gottl\"ober\altaffilmark{4}, 
G. Yepes\altaffilmark{5}
}

\altaffiltext{1}{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia, \email{je.forero@uniandes.edu.co}}
\altaffiltext{2}{Racah Institute of Physics, The Hebrew University of Jerusalem, 91904 Jerusalem, Israel}
\altaffiltext{3}{Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle 67 No. 53-108, Medell\'{\i}n, Colombia}
\altaffiltext{4}{Leibniz-Institut f\"ur Astrophysik, Potsdam, An der Sternwarte 16, 14482 Potsdam, Germany}
\altaffiltext{5}{Grupo de Astrof\'{\i}sica, Departamento de F\'{\i}sica Te\'orica, Universidad Aut\'onoma de Madrid,
Cantoblanco E-280049, Spain}

\date{\today}

\begin{abstract}

Recent observations constrained the tangential velocity of M31 with respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm 4.4$\kms \citep{vanderMarel12}. In this study we use a large volume high resolution N-body cosmological simulation to statistically study this kinematics in the context of the $\Lambda$CDM cosmology. We define the Local Group (LG) to be composed by two dominant separate halos hosting the MW and the M31 galaxy fulfilling certain isolation criteria. We find that the most probable values for the tangential and radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-65\pm 5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm 5$\kms. Within a similar absolute uncertainty defined by observations the pairs centered around these values are $\sim5$ times more abundant than the pairs in the observational interval. Furthermore, we find that only $\sim8\%$ of the pairs show a ratio between its tangential and radial velocity consistent with observations. {\bf This tension increases if we make a narrower selection in the pair sample to match the observational constraints on pair separation and total mass. Expressing these results in terms of the angular momentum and mechanical energy per unit mass also reflects this disagreement albeit in a less strong fashion due to the larger uncertainties inferred for these physical quantities. We also include in our analysis results from constrained simulations and cosmic web charachterization techniques. The results of this study imply that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.}

\end{abstract}

\begin{keywords}
{galaxies: kinematics and dynamics, Local Group, methods:numerical}
\end{keywords}

\section{Introduction}

The Milky Way (MW) and Andromeda galaxy (M31) are the dominant galaxies in the Local Group (LG). Astronomical observations of their mass distribution impose constraints on the standard cosmological model. The satellite overabundance problem \citep{Klypin99,Moore99}, tidal disruption features \citep{pandas09} and the disk dominated morphology \citep{Kazantzidis2008} are examples on of how LG studies are linked to the cosmological context. Detailed studies on the Magellanic Clouds dynamics and their possible link to M31 add to the interest of understanding the details of the LG kinematics and dynamics \citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}. However, a general concern in the use of the LG as a tool for near-field cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG regarding the properties of interest \citep{Liu2011,ForeroRomero2011,Purcell2012}.

A new valuable piece of information in this issue is the recent observational determination of the proper-motion measurements of M31, which until recently had been out of reach \citep{vanderMarel12}. This measurement opens up the possibility to study in detail the dynamics of the LG in a cosmological context. The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 32$ km s$^{-1}$. Together with the values of the relative radial velocity of $v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative motion of the MW and Andromeda is consistent with a head-on collision. With this information it is possible to ask how common is this kinematic configuration in a $\Lambda$CDM Universe.

This Letter presents such study. We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar isolation criteria as inferred in the LG. We quantify these results in terms of the number of pairs with given radial and tangential velocities in the galactocentric rest frame. We also find the pairs that are consistent with a head on collision in terms of the ratio of the radial to tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.3$ and present these results in terms of the reduced angular momentum and mechanical energy.

In addition we make use of three N-body simulations with constrained initial conditions which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc. The special feature of these simulations is that each volume features a pair of halos with the right isolation and relative positioning characteristics to be considered Local Group halos.  Finally, we provide an interpretation of these results in terms of the large scale structure environment. We find what places in the cosmic web favor the presence of dark matter halos with kinematics similar to the LG.

This Letter is structured as follows. In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs and the method we use to characterize the cosmic web. In Section \ref{sec:results} we present the results for the dynamics in these pairs in terms of the tangential/radial velocities and the orbital angular momentum/mechanical energy. In the same section we investigate what kind of large scale environment can foster the production of pairs with dynamics similar to those observed in the LG. Finally, in the last section we comment and conclude about the implications of these results in the context of the $\Lambda$CDM model.

\section{Simulation and Environment}
\label{sec:methods}
\subsection{The Bolshoi and Constrained Simulations}

\label{subsec:lg-sample}
\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation.pdf}
\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass.pdf}
\caption{\label{fig:distros} Distributions of separations and total halo mass (Milky Way and M31) in the LG-like pairs selected from the Bolshoi Simulation.}
\end{center}
\end{figure*}

The Bolshoi simulation follows the non-linear evolution of the dark matter density field using N-body techniques. The simulation has a cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$ particles. The cosmological parameters used in the simulation are $\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and $n=0.95$, corresponding to the matter density, vacuum energy density, the normalization of the power spectrum, the reduced Hubble constant and the index of the slope in the initial power spectrum. This set of parameter is compatible with the analysis  of the seventh year of data from the Wilkinson Microwave Anisotropy Probe (WMAP) \citep{Jarosik2011}. A detailed description of this simulation can be found in \citep{Bolshoi}.

With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun. In this paper we use the halos obtained through a FOF algorithm with a linking length $0.17$ times the average interparticle distance. We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}. The database allows us to obtain the comoving positions, peculiar velocities and masses for all the halos in the simulation volume at $z=0$. The Hubble flow is taken into account to convert the relative velocities into physical velocities and allow for a comparison with observations.

The constrained simulations we use in this Letter are part of the Constrained Local UniversE Simulations (CLUES) project whose main objective is to reproduce the large scale structure in the Local Universe as accurately as observations permit. The algorithm and observational constraints to construct the initial conditions are described in \citep{clues2010}.  We use three dark matter only simulations, each has a cubic volume of $64$\hMpc on a side, with the density field sampled with $1024^3$ particles. The cosmological density parameter is $\Omega_m=0.28$, the cosmological constant $\Omega_{\Lambda}=0.72$, the dimensionless Hubble parameter $h=0.73$, the spectral index of the primordial density perturbations $n=0.96$ and the power spectrum normalization $\sigma_{8}=0.817$, also consistent with WMAP 7th year data. The halo catalogs are constructed with the same FOF code used on the Bolshoi Simulation.

\subsection{The LG-sample}
Based on the FOF catalogs in the Bolshoi simulation we construct a halo pairs sample with masses and isolation properties consistent with the inferred properties for the MW and M31. The characteristics we impose to define a LG-like halo pair are the following:

\begin{enumerate}
\item Each halo has a mass in the range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$.
\item With respect to each halo, there cannot be any other halo within the mass range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$ closer than its partner. It means that there cannot be ambiguity on the identity of the pair members.
\item The relative radial velocity between the two halos is negative \citep{vanderMarel12}.
\item The distance between the center of mass of the halos must be less than $0.7$\hMpc \citep{ribas05,vanderMarel08}.
\item There cannot be halos more massive than $5\times 10^{12}$\hMsun within a radius of $2$\hMpc with respect to every object centre \citep{Karachentsev04,Anton09}.
\item There cannot be halos more massive than $5\times 10^{13}$\hMsun with  a radius of $3$\hMpc with respect to every object centre \citep{Karachentsev04}.
\end{enumerate}

This sample in the Bolshoi simulation has $1256$ pairs. This is $\sim 30\%$ smaller than the the number of pairs we found in \cite{ForeroRomero2011}, which did not include the condition 5) in the construction of the sample. Additionally, there is a sample of three (3) pairs constructed from the three constrained realizations. These pairs fulfill all the above mentioned conditions and additionally are located in a place with the right distances with respect to the Virgo and Coma clusters in the simulation. Figure \ref{fig:distros} show the distribution of separations and total pair mass computed from the $1256$ pairs in the Bolshoi simulation.

{\bf Additional constraints will be set on this population in order to match the observational estimates for separation and total mass. The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}.}

\begin{table}
\caption{Summary of the kinematic properties for M31 in the galactocentric frame as reported by \citep{vanderMarel12}. Values in parenthesis correspond to vector components}
\begin{center}
\begin{tabular}{ccc}\hline\hline
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\\hline
\end{tabular}\\
\vspace{1mm}
Note: the observational uncertainties in the position vector correspond to a $5\%$ in each component consistent with the uncertainties in the distance \citep[see references in][]{vanderMarel08}
\end{center}
\label{table:1}
\end{table}

\subsection{The kinematic environment}
We implement a large scale environment classification to quantify the dynamical state of the dark matter field. A detailed description of the algorithm can be found in \cite{Hoffman12}. Here we briefly describe its main physical assumptions.

The classification is based on computing the following velocity shear over a grid:

\begin{equation}
\Sigma_{ij}=-\frac{1}{2H_{0}}\left(\frac{\partial v_{i}}{\partial r_{j}}+\frac{\partial v_{j}}{\partial r_{i}}\right),
\end{equation}

where $i,j=x,y,z$, $H_{0}$ is the Hubble constant and $v_{i}, r_{i}$ correspond to the peculiar velocities and comoving positions, respectively. The matrix  $\Sigma_{ij}$ is symmetric and real, allowing for its diagonalization with real eigenvalues $\lambda_{1}>\lambda_{2}>\lambda_{3}$. A positive value of $\lambda_{i}>0$ implies collapse in comoving coordinates along the direction defined the corresponding eigenvector $u_{i}$. This method allows for the classification of each point as a peak, filament, sheet and void, if there are three, two, one and zero eigenvalues larger than a given threshold $\lambda_{\rm th}$, respectively.

Environment classification is an intrinsic multi-scale problem. Morphological features can appear or disappear depending on the physical scale that defines the density and velocity field. In the case of the LG kinematics there is a physical scale defined by the typical separation between the MW and M31, which is on the order of $\sim 0.5$ \hMpc. For this reason, all the results for the environment that we present here are computed over a grid of $256^3$ cells with a spacing of $0.97$\hMpc, where a first CIC interpolation is performed and later a Gaussian smoothing with a variance scale of $0.97$\hMpc is applied. In this Letter we are interested in the possible range of $\lambda_{i}$ spanned by the halo pairs that display dynamics similar to those of the observed LG.

\section{Results}
\label{sec:results}

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt.pdf}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_narrow_rt.pdf}
\caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation. The left panel corresponds to the general LG-like pair simple, the right panel corresponds to the subsample with cuts on pair separation and total mass, matching the observational constraints. The squared panel shows as a shaded histogram the number of pairs on the radial and tangential velocities plane. The rectangular panels show the histograms when only one of the velocity components is used to bin the data. The half ellipsoid corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}. The circles represent the positions of the pairs from the constrained simulations. The highest number density in the plot is obtained around $v_{\rm rad,\Lambda CDM} = 65\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms}
\label{fig:rt}
\end{center}

\end{figure*}

\subsection{Radial and tangential velocities}

Figure \ref{fig:rt} summarizes the central finding of this Letter. Most of the LG-like pairs (left panel) in the Bolshoi Simulation and the pairs in the constrained simulation have radial and tangential velocities notably different from the observational constraints.  The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -65\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram. There are $45$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations. This makes the LG-like pairs consistent with observations $\sim5$ times less likely to be found than a pair with velocities around $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$.

From Figure \ref{fig:rt} it is also clear that there is a significant number of pairs with a high tangential-to-radial velocity ratio. The peak in the pair number density is located around a region of $f_{\rm t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 1$, while the observations suggest $f_{\rm t}<0.32$. In the Bolshoi Simulation, out of $1256$ LG-like pairs we find that $108$ pairs ($\sim 8\%$ of the total sample) have values for $f_{\rm t}$ consistent with that constraint. {\bf The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.}

The right panel in Figure \ref{fig:rt} shows the results of making narrower selection of the LG-like pairs to separations between $700-800$ kpc and a total mass in the two halos to be in the range $1-4 \times 10^{12}$\Msun. This change emphasizes the tension found from the full LG-like sample. The sample is smaller ($1256$ pairs are reduced to $180$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -36\pm 5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $10$ pairs around that value and zero pairs consistent with observations. This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to the separation and total mass conditions used to define the LG-like pairs. {\bf We note that this tighter selection also excludes 2 of the 3 pairs from the constrained simulations.}

\subsection{Reduced Angular Momentum and Energy}

To gain additional physical insight we consider the two halos as points of mass $m_{\rm M31}$ and $m_{\rm MW}$. In the center of mass their orbit has an angular momentum $L=\mu|{{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and mechanical energy $E=\frac{1}{2}\mu |{\bf v}_{\rm M31}|^2 - G\mu M/|{\bf r}_{\rm M31}|$, where the total mass is $M=m_{\rm M31}+m_{\rm MW}$, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the gravitational constant. We express the results of the previous section in terms of the reduced angular momentum and energy $l_{orb}=L/\mu$ and $e_{\rm tot}=E/\mu$.

{\bf This formulation has a clear theoretical advantage if one considers the angular momentum and the mechanical energy as quasi-conserved dynamical quantities. This means that, after some formation time, these quantities do not significantly evolve as it is the case for the radial and tangential velocities. But there is an observational disadvantage. These quantities have to be derived from observations, increasing the uncertainties in their determination.}

In Table \ref{table:1} we list the M31 position (${\bf r}_{\rm M31}$), M31 velocity (${\bf v}_{\rm M31}$) and total mass ($M$) that we use to estimate the uncertainties in the $e_{\rm tot}-l_{\rm orb}$ plane. To do this we generate $10^6$ pairs with different values for these three variables using a Gaussian distribution for each vector component with a variance equal to their uncertainties ${\sigma_{\bf r}}_{\rm M31}$ and ${\sigma_{\bf v}}_{\rm M31}$).. For each pair we calculate its $l_{\rm orb}$ and $e_{\rm tot}$ values.

The results are summarized in Figure \ref{fig:EJ} following the same plotting conventions from Figure \ref{fig:rt}. The closed line is the iso-density contour in the 2D histogram enclosing $68\%$ of the Monte Carlo generated pairs consistent with observations. The poor observational constraints on the $e_{\rm tot}-l_{\rm orb}$ plane in comparison to the radial-tangential velocity plane comes from the fact that we are adding up large uncertainties in three different aspects. First, the tangential velocity has a $100\%$ uncertainty, making the angular momentum compatible with very low and high values. Second, the uncertainty on the square of the norm of the velocity has an uncertainty of $40\%$, having an impact on the kinetic energy. Third, the uncertainty in the total mass of the Local Group is close to $20\%$, which has an impact on the potential energy.

In the right panel of Figure \ref{fig:EJ} we see that further selection of LG-like pairs to have similar separation and total mass makes them occupy a narrower region. This shows that there are preferred values for $e_{\rm tot}-l_{\rm orb}$. Under the assumption of quasi-conserved quantity it means that their formation and evolution conditions must be similar. In this case the agreement with observations is moderate. A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would clearly bring the observations in the region of low $l_{\rm orb}$, clearly below the average $\Lambda$CDM expectation.

\subsection{Environment}

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ.pdf}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_narrow_EJ.pdf}
\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses. The panel distribution is the same as in Figure \ref{fig:rt}. The contour line shows the $1-\sigma$ uncertainty estimated through Monte Carlo techniques from the observational values summarized in Table 1. The poorer constraints on this plane are due to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$). }
\label{fig:EJ}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.50\textwidth]{./figures/test_lambda.pdf}
\caption{Integrated probability distributions for the three eigenvalues in the V-web environment classification algorithms. The dark region marks the eigenvalue interval defined by the nine (9) pairs with similar kinematics as observed in the Local Group.}
\label{fig:lambda}
\end{center}
\end{figure}

We compute the V-web eigenvalue distribution for all halos more massive than $10^{11}$\hMsun in the Bolshoi Simulation. Figure \ref{fig:lambda} illustrates the results of this exercise. The solid line shows the integrated distribution for the three eigenvalues. The shadow regions are defined by the minimum and maximum of the eigenvalues in the nine (9) pairs in the LG-like sample that show similar kinematics as the observations.

There is a preference for the largest eigenvalue to be positive  and have an upper bound $\lambda_{1}<1$, the second largest eigenvalue is bounded between $-0.1 < \lambda_{2} < 0.3$, while the smallest eigenvalue is always negative with a lower bound $\lambda_{3}>-0.5$. Under these conditions the peak and the void environment are completely excluded as environmental hosts of LG-like pairs for reasonable values of the threshold eigenvalue $\lambda_{\rm th}\approx 0.2-0.4$ \citep{Hoffman12,Libeskind12}. All pairs with kinematics similar to observations must reside in filaments or sheets. The result is similar if we consider instead the $108$ pairs with a weak tangential velocity component $f_{\rm t}<0.32$.

However, the inverse condition is weak. Making a selection of pairs by constraining the eigenvalues $\lambda_{i}$ to the interval mentioned above, does not bias significantly the values for the radial and tangential velocities. This suggests that residing in filaments/sheets is a condition common to isolated pairs, regardless of their kinematics. A detailed study on the influence of environment in LG-like pairs will be presented in a forthcoming publication (Bustamante et al. {\it in prep}).

\section{Conclusions}
We have presented a comparison between the observed kinematics for the M31 in the galactocentric restframe and the expectations for a large N-body cosmological simulation in the $\Lambda$CDM cosmology. In the simulation we select a sample of halo pairs in the mass range $5\times 10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation conditions of the Local Group from other massive structures. While the observations show that M31 moves towards us with a highly radial velocity, the simulation shows that the most common configuration at $z=0$ has values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm5$ \kms.

Using the same absolute values for the uncertainty in the observed velocity components, we find that halos within the preferred $\Lambda$CDM values are five times more common than pairs compatible with the observational constraint.  The qualitative nature of these results is still valid after a narrower selection on separation and total pair mass. Additionally, pairs with a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar to observations) represent $8\%$ of the total sample of LG-like pairs. Making an tighter selection to match the observational constraints on the separation and total mass results in zero pairs compatible with observations.

Approximating the LG as two point masses we express the above mentioned results in terms of the orbital angular momentum $l_{\rm orb}$ and the mechanical energy $e_{\rm tot}$ per unit of reduced mass. We find that the uncertainties in the tangential velocity, the square of the norm of the velocity and the total mass in the LG produce poorer constraints on the number of simulated pairs that are consistent with the observations. Nevertheless, in the case of the LG-pair sample that also fulfills the separation and total mass criteria there is a slight tension between simulation and observation. A reduction by a factor of $2$ in the observational uncertainty on the radial velocity would clarify this issue.

{\bf In the three pairs from constrained simulations we find kinematics dominated by radial velocities. However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in modest concordance with observations. This conforms to the fact that the observational values do not seem to be common in a $\Lambda$CDM Universe. Additionally we have found that pairs with velocities dominated by the radial component (by the same amount allowed by observations) are preferably to be found in filaments and sheets of dark matter. }

{\bf In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and mechanical energy. These values are in tension with the observational results. Pairs with highly radial orbits are not common in $\Lambda$CDM. This hints that the formation and evolution of the Local Group had special characteristics with respect to the average expectations in $\Lambda$CDM. Understanding the origin and implications of this result deserve further study.}

\label{sec:conclusions}
\section*{Acknowledgments}  
J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de Investigaciones at Universidad de los Andes through its {\it Fondo de Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber Foundation through its fellowship administered by the International Astronomical Union.

\bibliographystyle{apj}
%\bibliography{references} 
\begin{thebibliography}{25}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Besla} {et~al.}(2007){Besla}, {Kallivayalil}, {Hernquist},
  {Robertson}, {Cox}, {van der Marel}, \& {Alcock}}]{Besla2007}
{Besla}, G., {Kallivayalil}, N., {Hernquist}, L., {Robertson}, B., {Cox},
  T.~J., {van der Marel}, R.~P., \& {Alcock}, C. 2007, \apj, 668, 949

\bibitem[{{Forero-Romero} {et~al.}(2011){Forero-Romero}, {Hoffman}, {Yepes},
  {Gottl{\"o}ber}, {Piontek}, {Klypin}, \& {Steinmetz}}]{ForeroRomero2011}
{Forero-Romero}, J.~E., {Hoffman}, Y., {Yepes}, G., {Gottl{\"o}ber}, S.,
  {Piontek}, R., {Klypin}, A., \& {Steinmetz}, M. 2011, \mnras, 417, 1434

\bibitem[{{Fouquet} {et~al.}(2012){Fouquet}, {Hammer}, {Yang}, {Puech}, \&
  {Flores}}]{Fouquet2012}
{Fouquet}, S., {Hammer}, F., {Yang}, Y., {Puech}, M., \& {Flores}, H. 2012,
  \mnras, 427, 1769

\bibitem[{{Freeman} \& {Bland-Hawthorn}(2002)}]{Freeman2002}
{Freeman}, K., \& {Bland-Hawthorn}, J. 2002, \araa, 40, 487

\bibitem[{{Gottloeber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
  {Yepes}}]{clues2010}
{Gottloeber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, ArXiv e-prints

\bibitem[{{Hoffman} {et~al.}(2012){Hoffman}, {Metuki}, {Yepes},
  {Gottl{\"o}ber}, {Forero-Romero}, {Libeskind}, \& {Knebe}}]{Hoffman12}
{Hoffman}, Y., {Metuki}, O., {Yepes}, G., {Gottl{\"o}ber}, S., {Forero-Romero},
  J.~E., {Libeskind}, N.~I., \& {Knebe}, A. 2012, \mnras, 425, 2049

\bibitem[{{Jarosik} {et~al.}(2011){Jarosik}, {Bennett}, {Dunkley}, {Gold},
  {Greason}, {Halpern}, {Hill}, {Hinshaw}, {Kogut}, {Komatsu}, {Larson},
  {Limon}, {Meyer}, {Nolta}, {Odegard}, {Page}, {Smith}, {Spergel}, {Tucker},
  {Weiland}, {Wollack}, \& {Wright}}]{Jarosik2011}
{Jarosik}, N., {Bennett}, C.~L., {Dunkley}, J., {Gold}, B., {Greason}, M.~R.,
  {Halpern}, M., {Hill}, R.~S., {Hinshaw}, G., {Kogut}, A., {Komatsu}, E.,
  {Larson}, D., {Limon}, M., {Meyer}, S.~S., {Nolta}, M.~R., {Odegard}, N.,
  {Page}, L., {Smith}, K.~M., {Spergel}, D.~N., {Tucker}, G.~S., {Weiland},
  J.~L., {Wollack}, E., \& {Wright}, E.~L. 2011, \apjs, 192, 14

\bibitem[{{Karachentsev} {et~al.}(2004){Karachentsev}, {Karachentseva},
  {Huchtmeier}, \& {Makarov}}]{Karachentsev04}
{Karachentsev}, I.~D., {Karachentseva}, V.~E., {Huchtmeier}, W.~K., \&
  {Makarov}, D.~I. 2004, \aj, 127, 2031

\bibitem[{{Kazantzidis} {et~al.}(2008){Kazantzidis}, {Bullock}, {Zentner},
  {Kravtsov}, \& {Moustakas}}]{Kazantzidis2008}
{Kazantzidis}, S., {Bullock}, J.~S., {Zentner}, A.~R., {Kravtsov}, A.~V., \&
  {Moustakas}, L.~A. 2008, \apj, 688, 254

\bibitem[{{Klypin} {et~al.}(1999){Klypin}, {Kravtsov}, {Valenzuela}, \&
  {Prada}}]{Klypin99}
{Klypin}, A., {Kravtsov}, A.~V., {Valenzuela}, O., \& {Prada}, F. 1999, \apj,
  522, 82

\bibitem[{{Klypin} {et~al.}(2011){Klypin}, {Trujillo-Gomez}, \&
  {Primack}}]{Bolshoi}
{Klypin}, A.~A., {Trujillo-Gomez}, S., \& {Primack}, J. 2011, \apj, 740, 102

\bibitem[{{Knebe} {et~al.}(2011){Knebe}, {Libeskind}, {Doumler}, {Yepes},
  {Gottl{\"o}ber}, \& {Hoffman}}]{Knebe2011}
{Knebe}, A., {Libeskind}, N.~I., {Doumler}, T., {Yepes}, G., {Gottl{\"o}ber},
  S., \& {Hoffman}, Y. 2011, \mnras, 417, L56

\bibitem[{{Libeskind} {et~al.}(2012){Libeskind}, {Hoffman}, {Forero-Romero},
  {Gottl{\"o}ber}, {Knebe}, {Steinmetz}, \& {Klypin}}]{Libeskind12}
{Libeskind}, N.~I., {Hoffman}, Y., {Forero-Romero}, J., {Gottl{\"o}ber}, S.,
  {Knebe}, A., {Steinmetz}, M., \& {Klypin}, A. 2012, ArXiv e-prints

\bibitem[{{Liu} {et~al.}(2011){Liu}, {Gerke}, {Wechsler}, {Behroozi}, \&
  {Busha}}]{Liu2011}
{Liu}, L., {Gerke}, B.~F., {Wechsler}, R.~H., {Behroozi}, P.~S., \& {Busha},
  M.~T. 2011, \apj, 733, 62

\bibitem[{{McConnachie} {et~al.}(2009){McConnachie}, {Irwin}, {Ibata},
  {Dubinski}, {Widrow}, {Martin}, {C{\^o}t{\'e}}, {Dotter}, {Navarro},
  {Ferguson}, {Puzia}, {Lewis}, {Babul}, {Barmby}, {Bienaym{\'e}}, {Chapman},
  {Cockcroft}, {Collins}, {Fardal}, {Harris}, {Huxor}, {Mackey},
  {Pe{\~n}arrubia}, {Rich}, {Richer}, {Siebert}, {Tanvir}, {Valls-Gabaud}, \&
  {Venn}}]{pandas09}
{McConnachie}, A.~W., {Irwin}, M.~J., {Ibata}, R.~A., {Dubinski}, J., {Widrow},
  L.~M., {Martin}, N.~F., {C{\^o}t{\'e}}, P., {Dotter}, A.~L., {Navarro},
  J.~F., {Ferguson}, A.~M.~N., {Puzia}, T.~H., {Lewis}, G.~F., {Babul}, A.,
  {Barmby}, P., {Bienaym{\'e}}, O., {Chapman}, S.~C., {Cockcroft}, R.,
  {Collins}, M.~L.~M., {Fardal}, M.~A., {Harris}, W.~E., {Huxor}, A., {Mackey},
  A.~D., {Pe{\~n}arrubia}, J., {Rich}, R.~M., {Richer}, H.~B., {Siebert}, A.,
  {Tanvir}, N., {Valls-Gabaud}, D., \& {Venn}, K.~A. 2009, \nat, 461, 66

\bibitem[{{Moore} {et~al.}(1999){Moore}, {Ghigna}, {Governato}, {Lake},
  {Quinn}, {Stadel}, \& {Tozzi}}]{Moore99}
{Moore}, B., {Ghigna}, S., {Governato}, F., {Lake}, G., {Quinn}, T., {Stadel},
  J., \& {Tozzi}, P. 1999, \apjl, 524, L19

\bibitem[{{Peebles} \& {Nusser}(2010)}]{Peebles2010}
{Peebles}, P.~J.~E., \& {Nusser}, A. 2010, \nat, 465, 565

\bibitem[{{Purcell} \& {Zentner}(2012)}]{Purcell2012}
{Purcell}, C.~W., \& {Zentner}, A.~R. 2012, {JCAP}, 12, 7

\bibitem[{{Ribas} {et~al.}(2005){Ribas}, {Jordi}, {Vilardell}, {Fitzpatrick},
  {Hilditch}, \& {Guinan}}]{ribas05}
{Ribas}, I., {Jordi}, C., {Vilardell}, F., {Fitzpatrick}, E.~L., {Hilditch},
  R.~W., \& {Guinan}, E.~F. 2005, \apjl, 635, L37

\bibitem[{{Riebe} {et~al.}(2011){Riebe}, {Partl}, {Enke}, {Forero-Romero},
  {Gottloeber}, {Klypin}, {Lemson}, {Prada}, {Primack}, {Steinmetz}, \&
  {Turchaninov}}]{2011arXiv1109.0003R}
{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottloeber}, S.,
  {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz}, M.,
  \& {Turchaninov}, V. 2011, ArXiv e-prints

\bibitem[{{Teyssier} {et~al.}(2012){Teyssier}, {Johnston}, \&
  {Kuhlen}}]{Teyssier2012}
{Teyssier}, M., {Johnston}, K.~V., \& {Kuhlen}, M. 2012, \mnras, 426, 1808

\bibitem[{{Tikhonov} \& {Klypin}(2009)}]{Anton09}
{Tikhonov}, A.~V., \& {Klypin}, A. 2009, \mnras, 395, 1915

\bibitem[{{Tollerud} {et~al.}(2011){Tollerud}, {Boylan-Kolchin}, {Barton},
  {Bullock}, \& {Trinh}}]{Tollerud2011}
{Tollerud}, E.~J., {Boylan-Kolchin}, M., {Barton}, E.~J., {Bullock}, J.~S., \&
  {Trinh}, C.~Q. 2011, \apj, 738, 102

\bibitem[{{van der Marel} {et~al.}(2012){van der Marel}, {Fardal}, {Besla},
  {Beaton}, {Sohn}, {Anderson}, {Brown}, \& {Guhathakurta}}]{vanderMarel12}
{van der Marel}, R.~P., {Fardal}, M., {Besla}, G., {Beaton}, R.~L., {Sohn},
  S.~T., {Anderson}, J., {Brown}, T., \& {Guhathakurta}, P. 2012, \apj, 753, 8

\bibitem[{{van der Marel} \& {Guhathakurta}(2008)}]{vanderMarel08}
{van der Marel}, R.~P., \& {Guhathakurta}, P. 2008, \apj, 678, 187

\end{thebibliography}

\end{document}

### 2013-01-06 code/generateInfo.py

@@ -1,0 +1,134 @@
+from numpy import *
+def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='test.out', hubble=0.73):
+    """
+    This loads the data from a FOF catalog and a list of pairs IDs. 
+    From this information we calculate the distance between the pairs, 
+    radial and tangential physical velocities, masses, and angular momentum.
+
+    - "dmh_file" FOF file
+    - "pairs_ID_file" Pair indexes file
+    - "output_file" File to store the output
+    """
+    G_GRAV = 4.54E-48 #units of Mpc^3 Msun^-1 s^-2
+    KM_TO_MPC = 3.2E-20
+    E_UNITS = 1E-36
+    L_UNITS = 1.0
+
+    out = open(output_file, 'w')
+    outraw = open(output_file+'.raw', 'w')
+    halos_prop=loadtxt(dmh_filename) 
+    iso_pairs_prop=loadtxt(pairs_ID_filename) 
+    lambda_data=loadtxt(lambda_file)
+
+    outraw.write('# M1 M2 x1 y1 z1 x2 y2 z2 v1_x v1_y v1_z v2_x v2_y v2_z\n')
+    outraw.write('# Msun Msun Mpc Mpc Mpc Mpc Mpc Mpc km/s km/s kms/ km/s km/s km/s\n')
+
+
+    #print iso_pairs_prop, iso_pairs_prop.size
+    if(iso_pairs_prop.size==2):
+        iso_id_1 = empty([1])
+        iso_id_2 = empty([1])
+        iso_id_1[0] = int_(iso_pairs_prop[0])
+        iso_id_2[0] = int_(iso_pairs_prop[1])
+    else:
+        iso_id_1 = int_(iso_pairs_prop[:,0])
+        iso_id_2 = int_(iso_pairs_prop[:,1])
+
+    n_pairs = iso_id_1.size
+    print n_pairs
+    
+    n=0
+
+    for i in range(n_pairs):        
+        #pivot on the less massive halo (i.e. the MW)
+        id_1 = iso_id_2[i]-1
+        id_2 = iso_id_1[i]-1
+
+        props_id_1 = halos_prop[id_1,:]
+        props_id_2 = halos_prop[id_2,:]
+        
+        mass_1 = props_id_1[8]/hubble
+        mass_2 = props_id_2[8]/hubble
+                
[…]

### 2013-01-06 code/makeTempHistogramPlot.py

@@ -1,0 +1,201 @@
+from numpy import *
+import math
+import scipy 
+import numpy as np
+from pylab import *
+import matplotlib.pyplot as pl
+from matplotlib import rc
+from matplotlib.ticker import NullFormatter
+
+def ellipse(ra,rb,ang,xcenter,ycenter):
+    theta_bins = linspace(start = ang, stop = 2.0*pi, num = 100)
+    X = ra*cos(theta_bins) + xcenter
+    Y = rb*sin(theta_bins) + ycenter
+    return X,Y
+
+def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], xlims=-99, ylims =-99 , \
+                              nxbins = 50,nybins=50, bw=0, nbins=100,contours=1,sigma=1,line=1, 
+                          xlabel='A', ylabel='B', mean_obs=0, sigma_obs=0, 
+                          X_field=array([]), Y_field=array([]), Z_field=array([]), levels=array([])):
+#bw = 0 for color, = 1 for black and white
+#line = 0 for no line, =1 for line
+#sigma = 1 for display % below line, =0 for not
+#contours = 1 for display 1,2,3 sigma contours, = 0 for not.
+    
+    
+
+    # Define the x and y data
+    x = xdata
+    y = ydata
+
+# Set up default x and y limits
+    if (xlims == -99): xlims = [min(x),max(x)]
+    if (ylims == -99): ylims = [min(y),max(y)]
+    
+# Set up your x and y labels
+    xlabel = xlabel
+    ylabel = ylabel
+    mtitle = ''
+    
+    # Define the locations for the axes
+    left, width = 0.12, 0.55
+    bottom, height = 0.12, 0.55
+    bottom_h = left_h = left+width+0.02
+    
+    # Set up the geometry of the three plots
+    rect_temperature = [left, bottom, width, height] # dimensions of temp plot
+    rect_histx = [left, bottom_h, width, 0.25] # dimensions of x-histogram
+    rect_histy = [left_h, bottom, 0.25, height] # dimensions of y-histogram
+    
+    # Set up the size of the figure
+    fig = plt.figure(1, figsize=(9.5,9))
+    
+    # Make the three plots
+    axTemperature = plt.axes(rect_temperature) # temperature plot
+    axHistx = plt.axes(rect_histx) # x histogram
[…]

### 2013-01-08 code/generateInfo.py
version exploring the reduced spin parameter for the pair

@@ -94,6 +94,7 @@ def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='tes
         #separation
         norm_r_12 = sqrt(sum(r_12*r_12))
         total_mass = mass_1 + mass_2
+        reduced_mass  = mass_1 * mass_2 / total_mass
         #print "E", E_kin, E_pot, L_12_norm
         #decompose betwee radial and tangetial velocity
         norm_r_12 = sqrt(sum(r_12*r_12))
@@ -112,10 +113,10 @@ def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='tes
 
         if(v_radial_norm<0.0 and norm_r_12 < 1.0):
 #            print v_tan_norm, v_radial_norm, norm_v_12,lambda1_1, lambda2_1, lambda3_1, lambda1_2, lambda2_2, lambda3_2
-            out.write('%f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f\n' % 
+            out.write('%f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f\n' % 
                       (v_radial_norm, v_tan_norm, v_cm_norm, v_1_norm, 
                        v_2_norm, E_kin, E_pot, L_12_norm, lambda1_1, 
-                       lambda2_1, lambda3_1, lambda1_2, lambda2_2, lambda3_2, norm_r_12, total_mass))
+                       lambda2_1, lambda3_1, lambda1_2, lambda2_2, lambda3_2, norm_r_12, total_mass, reduced_mass))
 
             outraw.write('%e %e %f %f %f %f %f %f %f %f %f %f %f %f\n' % 
                       (mass_1, mass_2, r_1[0], r_1[1], r_1[2], r_2[0], r_2[1], r_2[2],

### 2013-01-10 code/makeTempHistogramPlot.py
small modification to makeTempHistogramPlot to add a new option to write the caption for the clues points

@@ -13,7 +13,7 @@ def ellipse(ra,rb,ang,xcenter,ycenter):
     Y = rb*sin(theta_bins) + ycenter
     return X,Y
 
-def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], xlims=-99, ylims =-99 , \
+def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], clues_label='',xlims=-99, ylims =-99 , \
                               nxbins = 50,nybins=50, bw=0, nbins=100,contours=1,sigma=1,line=1, 
                           xlabel='A', ylabel='B', mean_obs=0, sigma_obs=0, 
                           X_field=array([]), Y_field=array([]), Z_field=array([]), levels=array([])):
@@ -88,8 +88,8 @@ def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], xlims=-99, ylims =
     else: contourcolor = 'white'
 
     if clues.size:
-        axTemperature.scatter(clues[:,0], clues[:,1], color='white', edgecolors='black', s=200, label='$\mathrm{Constrained\ Simulations}$')        
-        axTemperature.legend(loc=0, scatterpoints=1)
+        axTemperature.scatter(clues[:,0], clues[:,1], color='white', edgecolors='black', s=200, label=clues_label)        
+        axTemperature.legend(loc=0, scatterpoints=1, fontsize=20)
 
     if X_field.size:
         print 'HOLA'

### 2013-01-10 paper/ms.tex
removed in the manuscript most references to the environment

@@
 
-\subsection{The kinematic environment}
-We implement a large scale environment classification to quantify the dynamical state of the dark matter field.
-A detailed description of the algorithm can be found in \cite{Hoffman12}.
-Here we briefly describe its main physical assumptions.
-
-The classification is based on computing the following velocity shear over a grid:
-
-\begin{equation}
-\Sigma_{ij}=-\frac{1}{2H_{0}}\left(\frac{\partial v_{i}}{\partial r_{j}}+\frac{\partial v_{j}}{\partial r_{i}}\right),
-\end{equation}
-
-where $i,j=x,y,z$, $H_{0}$ is the Hubble constant and $v_{i}, r_{i}$ correspond to the peculiar velocities and comoving positions, respectively.
-The matrix  $\Sigma_{ij}$ is symmetric and real, allowing for its diagonalization with real eigenvalues $\lambda_{1}>\lambda_{2}>\lambda_{3}$.
-A positive value of $\lambda_{i}>0$ implies collapse in comoving coordinates along the direction defined the corresponding eigenvector $u_{i}$.
-This method allows for the classification of each point as a peak, filament, sheet and void, if there are three, two, one and zero eigenvalues larger than a given threshold $\lambda_{\rm th}$, respectively.
-
-Environment classification is an intrinsic multi-scale problem.
-Morphological features can appear or disappear depending on the physical scale that defines the density and velocity field.
-In the case of the LG kinematics there is a physical scale defined by the typical separation between the MW and M31, which is on the order of $\sim 0.5$ \hMpc.
-For this reason, all the results for the environment that we present here are computed over a grid of $256^3$ cells with a spacing of $0.97$\hMpc, where a first CIC interpolation is performed and later a Gaussian smoothing with a variance scale of $0.97$\hMpc is applied.
-In this Letter we are interested in the possible range of $\lambda_{i}$ spanned by the halo pairs that display dynamics similar to those of the observed LG.
-
 \section{Results}
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_narrow_rt.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow.pdf}
 \caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_narrow_EJ.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow.pdf}
 \caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.50\textwidth]{./figures/test_lambda.pdf}
-\caption{Integrated probability distributions for the three eigenvalues in the V-web environment classification algorithms.
-The dark region marks the eigenvalue interval defined by the nine (9) pairs with similar kinematics as observed in the Local Group.}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_narrow.pdf}
+\caption{Spin parameter}
 \label{fig:lambda}
 \end{center}
-\end{figure}
-
-We compute the V-web eigenvalue distribution for all halos more massive than $10^{11}$\hMsun in the Bolshoi Simulation.
-Figure \ref{fig:lambda} illustrates the results of this exercise.
-The solid line shows the integrated distribution for the three eigenvalues.
-The shadow regions are defined by the minimum and maximum of the eigenvalues in the nine (9) pairs in the LG-like sample that show similar kinematics as the observations.
-
-There is a preference for the largest eigenvalue to be positive  and have an upper bound $\lambda_{1}<1$, the second largest eigenvalue is bounded between $-0.1 < \lambda_{2} < 0.3$, while the smallest eigenvalue is always negative with a lower bound $\lambda_{3}>-0.5$.
-Under these conditions the peak and the void environment are completely excluded as environmental hosts of LG-like pairs for reasonable values of the threshold eigenvalue $\lambda_{\rm th}\approx 0.2-0.4$ \citep{Hoffman12,Libeskind12}.
-All pairs with kinematics similar to observations must reside in filaments or sheets.
-The result is similar if we consider instead the $108$ pairs with a weak tangential velocity component $f_{\rm t}<0.32$.
-
-However, the inverse condition is weak.
-Making a selection of pairs by constraining the eigenvalues $\lambda_{i}$ to the interval mentioned above, does not bias significantly the values for the radial and tangential velocities.
-This suggests that residing in filaments/sheets is a condition common to isolated pairs, regardless of their kinematics.
-A detailed study on the influence of environment in LG-like pairs will be presented in a forthcoming publication (Bustamante et al.
-{\it in prep}).
+\end{figure*}
 
@@
 However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in modest concordance with observations.
-This conforms to the fact that the observational values do not seem to be common in a $\Lambda$CDM Universe.
-Additionally we have found that pairs with velocities dominated by the radial component (by the same amount allowed by observations) are preferably to be found in filaments and sheets of dark matter.
-}
+This conforms to the fact that the observational values do not seem to be common in a $\Lambda$CDM Universe.}
 
@@
 \bibliographystyle{apj}
-%\bibliography{references} 
-\begin{thebibliography}{25}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{{Besla} {et~al.}(2007){Besla}, {Kallivayalil}, {Hernquist},
-  {Robertson}, {Cox}, {van der Marel}, \& {Alcock}}]{Besla2007}
-{Besla}, G., {Kallivayalil}, N., {Hernquist}, L., {Robertson}, B., {Cox},
-  T.~J., {van der Marel}, R.~P., \& {Alcock}, C. 2007, \apj, 668, 949
-
-\bibitem[{{Forero-Romero} {et~al.}(2011){Forero-Romero}, {Hoffman}, {Yepes},
-  {Gottl{\"o}ber}, {Piontek}, {Klypin}, \& {Steinmetz}}]{ForeroRomero2011}
-{Forero-Romero}, J.~E., {Hoffman}, Y., {Yepes}, G., {Gottl{\"o}ber}, S.,
-  {Piontek}, R., {Klypin}, A., \& {Steinmetz}, M. 2011, \mnras, 417, 1434
-
-\bibitem[{{Fouquet} {et~al.}(2012){Fouquet}, {Hammer}, {Yang}, {Puech}, \&
-  {Flores}}]{Fouquet2012}
-{Fouquet}, S., {Hammer}, F., {Yang}, Y., {Puech}, M., \& {Flores}, H. 2012,
-  \mnras, 427, 1769
-
-\bibitem[{{Freeman} \& {Bland-Hawthorn}(2002)}]{Freeman2002}
-{Freeman}, K., \& {Bland-Hawthorn}, J. 2002, \araa, 40, 487
-
-\bibitem[{{Gottloeber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
-  {Yepes}}]{clues2010}
-{Gottloeber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, ArXiv e-prints
-
-\bibitem[{{Hoffman} {et~al.}(2012){Hoffman}, {Metuki}, {Yepes},
-  {Gottl{\"o}ber}, {Forero-Romero}, {Libeskind}, \& {Knebe}}]{Hoffman12}
-{Hoffman}, Y., {Metuki}, O., {Yepes}, G., {Gottl{\"o}ber}, S., {Forero-Romero},
-  J.~E., {Libeskind}, N.~I., \& {Knebe}, A. 2012, \mnras, 425, 2049
-
-\bibitem[{{Jarosik} {et~al.}(2011){Jarosik}, {Bennett}, {Dunkley}, {Gold},
-  {Greason}, {Halpern}, {Hill}, {Hinshaw}, {Kogut}, {Komatsu}, {Larson},
-  {Limon}, {Meyer}, {Nolta}, {Odegard}, {Page}, {Smith}, {Spergel}, {Tucker},
-  {Weiland}, {Wollack}, \& {Wright}}]{Jarosik2011}
-{Jarosik}, N., {Bennett}, C.~L., {Dunkley}, J., {Gold}, B., {Greason}, M.~R.,
-  {Halpern}, M., {Hill}, R.~S., {Hinshaw}, G., {Kogut}, A., {Komatsu}, E.,
-  {Larson}, D., {Limon}, M., {Meyer}, S.~S., {Nolta}, M.~R., {Odegard}, N.,
-  {Page}, L., {Smith}, K.~M., {Spergel}, D.~N., {Tucker}, G.~S., {Weiland},
-  J.~L., {Wollack}, E., \& {Wright}, E.~L. 2011, \apjs, 192, 14
-
-\bibitem[{{Karachentsev} {et~al.}(2004){Karachentsev}, {Karachentseva},
-  {Huchtmeier}, \& {Makarov}}]{Karachentsev04}
-{Karachentsev}, I.~D., {Karachentseva}, V.~E., {Huchtmeier}, W.~K., \&
-  {Makarov}, D.~I. 2004, \aj, 127, 2031
-
-\bibitem[{{Kazantzidis} {et~al.}(2008){Kazantzidis}, {Bullock}, {Zentner},
-  {Kravtsov}, \& {Moustakas}}]{Kazantzidis2008}
-{Kazantzidis}, S., {Bullock}, J.~S., {Zentner}, A.~R., {Kravtsov}, A.~V., \&
-  {Moustakas}, L.~A. 2008, \apj, 688, 254
-
-\bibitem[{{Klypin} {et~al.}(1999){Klypin}, {Kravtsov}, {Valenzuela}, \&
-  {Prada}}]{Klypin99}
-{Klypin}, A., {Kravtsov}, A.~V., {Valenzuela}, O., \& {Prada}, F. 1999, \apj,
-  522, 82
-
-\bibitem[{{Klypin} {et~al.}(2011){Klypin}, {Trujillo-Gomez}, \&
-  {Primack}}]{Bolshoi}
-{Klypin}, A.~A., {Trujillo-Gomez}, S., \& {Primack}, J. 2011, \apj, 740, 102
-
-\bibitem[{{Knebe} {et~al.}(2011){Knebe}, {Libeskind}, {Doumler}, {Yepes},
-  {Gottl{\"o}ber}, \& {Hoffman}}]{Knebe2011}
-{Knebe}, A., {Libeskind}, N.~I., {Doumler}, T., {Yepes}, G., {Gottl{\"o}ber},
-  S., \& {Hoffman}, Y. 2011, \mnras, 417, L56
-
-\bibitem[{{Libeskind} {et~al.}(2012){Libeskind}, {Hoffman}, {Forero-Romero},
-  {Gottl{\"o}ber}, {Knebe}, {Steinmetz}, \& {Klypin}}]{Libeskind12}
-{Libeskind}, N.~I., {Hoffman}, Y., {Forero-Romero}, J., {Gottl{\"o}ber}, S.,
-  {Knebe}, A., {Steinmetz}, M., \& {Klypin}, A. 2012, ArXiv e-prints
-
-\bibitem[{{Liu} {et~al.}(2011){Liu}, {Gerke}, {Wechsler}, {Behroozi}, \&
-  {Busha}}]{Liu2011}
-{Liu}, L., {Gerke}, B.~F., {Wechsler}, R.~H., {Behroozi}, P.~S., \& {Busha},
-  M.~T. 2011, \apj, 733, 62
-
-\bibitem[{{McConnachie} {et~al.}(2009){McConnachie}, {Irwin}, {Ibata},
-  {Dubinski}, {Widrow}, {Martin}, {C{\^o}t{\'e}}, {Dotter}, {Navarro},
-  {Ferguson}, {Puzia}, {Lewis}, {Babul}, {Barmby}, {Bienaym{\'e}}, {Chapman},
-  {Cockcroft}, {Collins}, {Fardal}, {Harris}, {Huxor}, {Mackey},
-  {Pe{\~n}arrubia}, {Rich}, {Richer}, {Siebert}, {Tanvir}, {Valls-Gabaud}, \&
-  {Venn}}]{pandas09}
-{McConnachie}, A.~W., {Irwin}, M.~J., {Ibata}, R.~A., {Dubinski}, J., {Widrow},
-  L.~M., {Martin}, N.~F., {C{\^o}t{\'e}}, P., {Dotter}, A.~L., {Navarro},
-  J.~F., {Ferguson}, A.~M.~N., {Puzia}, T.~H., {Lewis}, G.~F., {Babul}, A.,
-  {Barmby}, P., {Bienaym{\'e}}, O., {Chapman}, S.~C., {Cockcroft}, R.,
-  {Collins}, M.~L.~M., {Fardal}, M.~A., {Harris}, W.~E., {Huxor}, A., {Mackey},
-  A.~D., {Pe{\~n}arrubia}, J., {Rich}, R.~M., {Richer}, H.~B., {Siebert}, A.,
-  {Tanvir}, N., {Valls-Gabaud}, D., \& {Venn}, K.~A. 2009, \nat, 461, 66
-
-\bibitem[{{Moore} {et~al.}(1999){Moore}, {Ghigna}, {Governato}, {Lake},
-  {Quinn}, {Stadel}, \& {Tozzi}}]{Moore99}
-{Moore}, B., {Ghigna}, S., {Governato}, F., {Lake}, G., {Quinn}, T., {Stadel},
-  J., \& {Tozzi}, P. 1999, \apjl, 524, L19
-
-\bibitem[{{Peebles} \& {Nusser}(2010)}]{Peebles2010}
-{Peebles}, P.~J.~E., \& {Nusser}, A. 2010, \nat, 465, 565
-
-\bibitem[{{Purcell} \& {Zentner}(2012)}]{Purcell2012}
-{Purcell}, C.~W., \& {Zentner}, A.~R. 2012, {JCAP}, 12, 7
-
-\bibitem[{{Ribas} {et~al.}(2005){Ribas}, {Jordi}, {Vilardell}, {Fitzpatrick},
-  {Hilditch}, \& {Guinan}}]{ribas05}
-{Ribas}, I., {Jordi}, C., {Vilardell}, F., {Fitzpatrick}, E.~L., {Hilditch},
-  R.~W., \& {Guinan}, E.~F. 2005, \apjl, 635, L37
-
-\bibitem[{{Riebe} {et~al.}(2011){Riebe}, {Partl}, {Enke}, {Forero-Romero},
-  {Gottloeber}, {Klypin}, {Lemson}, {Prada}, {Primack}, {Steinmetz}, \&
-  {Turchaninov}}]{2011arXiv1109.0003R}
-{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottloeber}, S.,
-  {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz}, M.,
-  \& {Turchaninov}, V. 2011, ArXiv e-prints
-
-\bibitem[{{Teyssier} {et~al.}(2012){Teyssier}, {Johnston}, \&
-  {Kuhlen}}]{Teyssier2012}
-{Teyssier}, M., {Johnston}, K.~V., \& {Kuhlen}, M. 2012, \mnras, 426, 1808
-
-\bibitem[{{Tikhonov} \& {Klypin}(2009)}]{Anton09}
-{Tikhonov}, A.~V., \& {Klypin}, A. 2009, \mnras, 395, 1915
-
-\bibitem[{{Tollerud} {et~al.}(2011){Tollerud}, {Boylan-Kolchin}, {Barton},
-  {Bullock}, \& {Trinh}}]{Tollerud2011}
-{Tollerud}, E.~J., {Boylan-Kolchin}, M., {Barton}, E.~J., {Bullock}, J.~S., \&
-  {Trinh}, C.~Q. 2011, \apj, 738, 102
-
-\bibitem[{{van der Marel} {et~al.}(2012){van der Marel}, {Fardal}, {Besla},
-  {Beaton}, {Sohn}, {Anderson}, {Brown}, \& {Guhathakurta}}]{vanderMarel12}
-{van der Marel}, R.~P., {Fardal}, M., {Besla}, G., {Beaton}, R.~L., {Sohn},
-  S.~T., {Anderson}, J., {Brown}, T., \& {Guhathakurta}, P. 2012, \apj, 753, 8
-
-\bibitem[{{van der Marel} \& {Guhathakurta}(2008)}]{vanderMarel08}
-{van der Marel}, R.~P., \& {Guhathakurta}, P. 2008, \apj, 678, 187
-
-\end{thebibliography}
+\bibliography{references}
 

### 2013-01-11 paper/ms.tex
new text in the paper. started writing on the dimensionless spin parameter.

@@
 
-\shorttitle{The not so common kinematics of the Local Group}
+\shorttitle{Local Group Kinematics}
 \shortauthors{Forero-Romero et al.}
@@
 \begin{document}
-\title{The not so common kinematics of the Local Group}
+\title{The kinematics of the Local Group in a cosmological context}
 \author{
@@
 Expressing these results in terms of the angular momentum and mechanical energy per unit mass also reflects this disagreement albeit in a less strong fashion due to the larger uncertainties inferred for these physical quantities.
+From these quantities we define a reduced spin parameter and use it to easily characterize the LG with repect to other halo pairs in $\Lambda$CDM.
 We also include in our analysis results from constrained simulations and cosmic web charachterization techniques.
-The results of this study imply that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.}
-
+The results of this study suggest that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.}
 \end{abstract}
@@
 
+\subsection{Reduced Spin Parameter}
+
+In complete analogy to the spin parameter, $\lambda$, used to describe dark matter halos, we define the same parameter for halo pairs.
+This quantity combines into a single scalar the energy and the angular momentum allowing us to make an even more straightforward comparison between the results allowed for observations an the expectations in $\Lambda$CDM.
+The precise definition of the dimensionless $\lambda$ parameter is the following
+
+\begin{equation}
+\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}, 
+\end{equation}
+
+where the $\mu=m_{\rm M31}m_{\rm MW}/M)$ is the reduced mass for the pair.
+
+We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
+The results can be found in Figure \ref{fig:lambda}.
+The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
+
+Both panels tell in a clear way the same story as the $e_{\rm tot}-l_{\rm orb}$ plane.
+
 \subsection{Environment}
@@
 
+The data and codes that allow the results of this paper to be reproduced can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}
+
 \bibliographystyle{apj}

### 2013-01-12 paper/ms.tex
improved discusion on lambda parameter

@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow.pdf}
-\caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
-The left panel corresponds to the general LG-like pair simple, the right panel corresponds to the subsample with cuts on pair separation and total mass, matching the observational constraints.
-The squared panel shows as a shaded histogram the number of pairs on the radial and tangential velocities plane.
-The rectangular panels show the histograms when only one of the velocity components is used to bin the data.
-The half ellipsoid corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}.
-The circles represent the positions of the pairs from the constrained simulations.
-The highest number density in the plot is obtained around $v_{\rm rad,\Lambda CDM} = 65\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms}
-\label{fig:rt}
-\end{center}
-
-\end{figure*}
-
 \subsection{Radial and tangential velocities}
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow.pdf}
+\caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
+The left panel corresponds to the general LG-like pair simple, the right panel corresponds to the subsample with cuts on pair separation and total mass, matching the observational constraints.
+The squared panel shows as a shaded histogram the number of pairs on the radial and tangential velocities plane.
+The rectangular panels show the histograms when only one of the velocity components is used to bin the data.
+The half ellipsoid corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}.
+The circles represent the positions of the pairs from the constrained simulations.
+The highest number density in the plot is obtained around $v_{\rm rad,\Lambda CDM} = 65\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms}
+\label{fig:rt}
+\end{center}
+
+\end{figure*}
+
 \subsection{Reduced Angular Momentum and Energy}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow.pdf}
+\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
+The panel distribution is the same as in Figure \ref{fig:rt}.
+The contour line shows the $1-\sigma$ uncertainty estimated through Monte Carlo techniques from the observational values summarized in Table 1.
+The poorer constraints on this plane are due to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$).
+}
+\label{fig:EJ}
+\end{center}
+\end{figure*}
 
@@
 
-\subsection{Reduced Spin Parameter}
-
-In complete analogy to the spin parameter, $\lambda$, used to describe dark matter halos, we define the same parameter for halo pairs.
-This quantity combines into a single scalar the energy and the angular momentum allowing us to make an even more straightforward comparison between the results allowed for observations an the expectations in $\Lambda$CDM.
-The precise definition of the dimensionless $\lambda$ parameter is the following
-
-\begin{equation}
-\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}, 
-\end{equation}
-
-where the $\mu=m_{\rm M31}m_{\rm MW}/M)$ is the reduced mass for the pair.
-
-We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
-The results can be found in Figure \ref{fig:lambda}.
-The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
-
-Both panels tell in a clear way the same story as the $e_{\rm tot}-l_{\rm orb}$ plane.
-
-\subsection{Environment}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow.pdf}
-\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
-The panel distribution is the same as in Figure \ref{fig:rt}.
-The contour line shows the $1-\sigma$ uncertainty estimated through Monte Carlo techniques from the observational values summarized in Table 1.
-The poorer constraints on this plane are due to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$).
-}
-\label{fig:EJ}
-\end{center}
-\end{figure*}
-
 \begin{figure*}
@@
 \end{figure*}
+
+\subsection{Reduced Spin Parameter}
+
+{\bf In complete analogy to the spin parameter used to describe dark matter halos, we define the same parameter, $\lambda$, for halo pairs.
+This quantity combines into a single scalar the energy and the angular momentum allowing us to make an even more straightforward comparison between the results allowed for observations an the expectations in $\Lambda$CDM.
+The precise definition of the dimensionless $\lambda$ parameter is the following
+}
+
+\begin{equation}
+\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}, 
+\end{equation}
+
+where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.
+
+We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
+The results can be found in Figure \ref{fig:lambda}.
+The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
+
+Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
+The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
+
+In terms of the reduced spin parameter $\lambda$ seems to be lower in the LG than expected from LCDM.
+In the case where the pairs are restricted to have the same total mass and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are by construction in the same range as observations.
+The disagreement must come from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation, but still they are broadly consistent due to the large observational uncertainties.
+
+Figure \ref{fig:lambda} also shows the results from the three constrained simulations. The
 
@@
 
-The data and codes that allow the results of this paper to be reproduced can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}
+The data and codes needed to reproduce the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
 

### 2013-01-12 paper/ms.tex
full draft with lambda discussion before mailing to coauthors

@@
 
-\section{Simulation and Environment}
-\label{sec:methods}
-\subsection{The Bolshoi and Constrained Simulations}
-
 \label{subsec:lg-sample}
@@
 \end{figure*}
+
+\section{Simulation and Environment}
+\label{sec:methods}
+\subsection{The Bolshoi and Constrained Simulations}
 
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_narrow.pdf}
+\caption{{\rm \label{fig:lambda} Normalized histograms of the dimensionless spin parameter $\lambda$ for the pairs in the Bolshoi simulation and its infered values for the LG from the observational constraints from a Monte Carlo simulation.
+The left panel includes all the LG-pairs while the right panel includes a reduced sample of pairs that match the separation and total mass infered from Local Group observations.
+The vertical lines with the white dots represent the values inferred for pairs in the constrained simulations.}}
 \caption{Spin parameter}
@@
 
-\subsection{Reduced Spin Parameter}
+\subsection{Dimensionless Spin Parameter}
 
@@
 
-where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.
+{\bf where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.}
 
-We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
+{\bf We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
 The results can be found in Figure \ref{fig:lambda}.
 The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
+}
 
-Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
+{\bf Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
 The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
+}
 
-In terms of the reduced spin parameter $\lambda$ seems to be lower in the LG than expected from LCDM.
+{\bf In terms of the reduced spin parameter $\lambda$ seems to be lower in the LG than expected from LCDM.
 In the case where the pairs are restricted to have the same total mass and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are by construction in the same range as observations.
-The disagreement must come from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation, but still they are broadly consistent due to the large observational uncertainties.
+The disagreement must come from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation, but still they are broadly consistent due to the large observational uncertainties.}
 
-Figure \ref{fig:lambda} also shows the results from the three constrained simulations. The
+{\bf Left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations.
+In this case all cases seem to be a better match to the observational expectations than to the $\Lambda$CDM pair population.
+In the case of tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\lambda$ value matches with the maximum in the $\lambda$ distribution infered from observations.}
 
@@
 {\bf In the three pairs from constrained simulations we find kinematics dominated by radial velocities.
-However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in modest concordance with observations.
-This conforms to the fact that the observational values do not seem to be common in a $\Lambda$CDM Universe.}
+However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in broad concordance with observations.
+There is only one pair that fullfills all the separation, total mass constraints and matches the most probable value for the dimensionless spin parameter $\lambda$ inferred from observations.
+}
 
-{\bf In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and mechanical energy.
-These values are in tension with the observational results.
-Pairs with highly radial orbits are not common in $\Lambda$CDM.
-This hints that the formation and evolution of the Local Group had special characteristics with respect to the average expectations in $\Lambda$CDM.
-Understanding the origin and implications of this result deserve further study.}
+{\bf The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs.
+Reduced unceratainties in masses, distances and velocities can be included to produce an updated estimate for this parameter.
+This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.}
+
+{\bf In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and total mechanical energy.
+The values for the orbital angular momentum and energy, merged into the $\lambda$ spin parameter, are in mild disagreement with the observational constraints.
+However, there is a strong tension with the precise values for the radial and tangential velocities.
+
+Under the approximation of consevartion of the orbital angular momentum and mechanical energy, we see that an agreement in the total angular momentum and energy and a marked difference with the precise balance between today's radial and tangential velocities could only be explained if the initial conditions for the formation of the Local Group are special in comparison to the initial conditions of any other pair of dark matter halos in the $\Lambda$CDM cosmology.
+
+We consider that reducing the uncertainties in the observations relevant to this study are needed.
+Likewise, additional theoretical work to understand the origin and implications of this result deserve further study.}
 
@@
 
-The data and codes needed to reproduce the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
+The data and source code needed to reproduce the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
 

### 2013-01-21 paper/ms.tex
minor modifications to ms.tex with SG comments inside

@@
 \usepackage {graphicx}
-\bibliographystyle{apj}
+%\bibliographystyle{apj}
 \usepackage{graphics}
@@
 \font\smcap=cmcsc10
-\def\caii{Ca\,{\smcap ii}}
 
@@
 Furthermore, we find that only $\sim8\%$ of the pairs show a ratio between its tangential and radial velocity consistent with observations.
-{\bf This tension increases if we make a narrower selection in the pair sample to match the observational constraints on pair separation and total mass.
+This tension increases if we make a narrower selection in the pair sample to match the observational constraints on pair separation and total mass.
 Expressing these results in terms of the angular momentum and mechanical energy per unit mass also reflects this disagreement albeit in a less strong fashion due to the larger uncertainties inferred for these physical quantities.
-From these quantities we define a reduced spin parameter and use it to easily characterize the LG with repect to other halo pairs in $\Lambda$CDM.
-We also include in our analysis results from constrained simulations and cosmic web charachterization techniques.
-The results of this study suggest that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.}
+From these quantities Peeble's spin parameter is evaluated and is used to characterize the LG with repect to other halo pairs in $\Lambda$CDM.
+We also include in our analysis results from constrained simulations.
+The characterisation results of this study suggest that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.
 \end{abstract}
@@
 This measurement opens up the possibility to study in detail the dynamics of the LG in a cosmological context.
-The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 32$ km s$^{-1}$.
+The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km s$^{-1}$.
 Together with the values of the relative radial velocity of $v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative motion of the MW and Andromeda is consistent with a head-on collision.
@@
 The simulation has a cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$ particles.
-The cosmological parameters used in the simulation are $\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and $n=0.95$, corresponding to the matter density, vacuum energy density, the normalization of the power spectrum, the reduced Hubble constant and the index of the slope in the initial power spectrum.
-This set of parameter is compatible with the analysis  of the seventh year of data from the Wilkinson Microwave Anisotropy Probe (WMAP) \citep{Jarosik2011}.
+The cosmological parameters used in the simulation are $\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and $n=0.95$, corresponding to the matter density, vacuum energy density, the normalization of the power spectrum, the dimensionless Hubble constant and the index of the slope in the initial power spectrum.
+This set of parameters is compatible with the analysis  of the seventh year of data from the Wilkinson Microwave Anisotropy Probe (WMAP) \citep{Jarosik2011}.
 A detailed description of this simulation can be found in \citep{Bolshoi}.
 
-With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun.
+With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun. {\bf  In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm.}
+
 In this paper we use the halos obtained through a FOF algorithm with a linking length $0.17$ times the average interparticle distance.
@@
 The constrained simulations we use in this Letter are part of the Constrained Local UniversE Simulations (CLUES) project whose main objective is to reproduce the large scale structure in the Local Universe as accurately as observations permit.
-The algorithm and observational constraints to construct the initial conditions are described in \citep{clues2010}.
+The algorithm and observational constraints to construct the initial conditions are
+SG  I think the observations are much more accurate than our reconstructions, so it would be better to replace 
+ ``observations permit.''  by ``possible.'', which includes both the limitations of the observations and the reconstraction algorithm.
+ described in \citep{clues2010}.
 We use three dark matter only simulations, each has a cubic volume of $64$\hMpc on a side, with the density field sampled with $1024^3$ particles.
@@
 \item There cannot be halos more massive than $5\times 10^{12}$\hMsun within a radius of $2$\hMpc with respect to every object centre \citep{Karachentsev04,Anton09}.
-\item There cannot be halos more massive than $5\times 10^{13}$\hMsun with  a radius of $3$\hMpc with respect to every object centre \citep{Karachentsev04}.
+\item There cannot be halos more massive than $5\times 10^{13}$\hMsun with  a radius of $3$\hMpc with
+SG    within a radius of
+ respect to every object centre \citep{Karachentsev04}.
 \end{enumerate}
@@
 These pairs fulfill all the above mentioned conditions and additionally are located in a place with the right distances with respect to the Virgo and Coma clusters in the simulation.
-Figure \ref{fig:distros} show the distribution of separations and total pair mass computed from the $1256$ pairs in the Bolshoi simulation.
+Figure \ref{fig:distros} show the
+SG    Coma is outside of the 64 Mpc/h box
+ distribution of separations and total pair mass computed from the $1256$ pairs in the Bolshoi simulation.
 
@@
 The rectangular panels show the histograms when only one of the velocity components is used to bin the data.
-The half ellipsoid corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}.
+The half ellipsoid
+SG ellipse  -- ellipsoid is 3D or higher dim
+ corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}.
 The circles represent the positions of the pairs from the constrained simulations.
@@
 The contour line shows the $1-\sigma$ uncertainty estimated through Monte Carlo techniques from the observational values summarized in Table 1.
-The poorer constraints on this plane are due to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$).
-}
+The poorer constraints on this plane are due
+SG  poorer than what? it is explained in the text, but for clarity I would prefer something like
+The constraints in this plot are less restrictive than in Fig. 2 because we have added ......
+ to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$). }
 \label{fig:EJ}
@@
 
-In the right panel of Figure \ref{fig:EJ} we see that further selection of LG-like pairs to have similar separation and total mass makes them occupy a narrower region.
-This shows that there are preferred values for $e_{\rm tot}-l_{\rm orb}$.
+In the right panel of Figure \ref{fig:EJ} we see that further selection of LG-like pairs to have similar separation
+SG I think better saying more precise, i.e.
+``that further reduction of the LG-like pairs following the procedure decribed for the right panel of Fig.2  makes them occupy a narrower region in the total energy.''
+ and total mass makes them occupy a narrower region. This shows that there are preferred values for $e_{\rm
+SG I don't see this, looks like only e_tot is more constrained but not l_orb.
+ tot}-l_{\rm orb}$.
 Under the assumption of quasi-conserved quantity it means that their formation and evolution conditions must be similar.
@@
 
-{\bf In complete analogy to the spin parameter used to describe dark matter halos, we define the same parameter, $\lambda$, for halo pairs.
+{\bf In complete analogy to the spin parameter used to describe dark matter halos, we define the same
+SG remove ``complete'',  there is an analogy but not a complete or incomplete analogy 
+ parameter, $\lambda$, for halo pairs.
 This quantity combines into a single scalar the energy and the angular momentum allowing us to make an even more straightforward comparison between the results allowed for observations an the expectations in $\Lambda$CDM.
-The precise definition of the dimensionless $\lambda$ parameter is the following
+The precise definition of the
+SG remove ``precise'',  an unprecise definition would be not a definition
+ dimensionless $\lambda$ parameter is the following
 }
@@
 
-{\bf The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs.
-Reduced unceratainties in masses, distances and velocities can be included to produce an updated estimate for this parameter.
-This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.}
+{\bf The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs. Reduced unceratainties in masses, distances and velocities can be included to produce an updated
+SG uncertainties
+ estimate for this parameter. This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.}
 
@@
 
+SG: Please add the sentence form the multidark database page:
+The MultiDark Database used in this paper and the web application providing online access to it were constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064.
+The Bolshoi simulation was run on the NASA's Pleiades supercomputer at the NASA Ames Research Center.
+
 \bibliographystyle{apj}

### 2013-01-21 paper/ms.tex
Half-revised paper draft

@@
 Detailed studies on the Magellanic Clouds dynamics and their possible link to M31 add to the interest of understanding the details of the LG kinematics and dynamics \citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}.
-However, a general concern in the use of the LG as a tool for near-field cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG regarding the properties of interest \citep{Liu2011,ForeroRomero2011,Purcell2012}.
+However, a general concern in the use of the LG as a tool for near-field cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG regarding the properties of interest \citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.
 
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation.pdf}
-\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass.pdf}
+\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
 \caption{\label{fig:distros} Distributions of separations and total halo mass (Milky Way and M31) in the LG-like pairs selected from the Bolshoi Simulation.}
@@
 
-With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun. {\bf  In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm.}
-
-In this paper we use the halos obtained through a FOF algorithm with a linking length $0.17$ times the average interparticle distance.
-We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
+With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun.
+{\bf  In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}.
+The halos are selected to have an overdensity of 200 times the critical overdensity.
+Furthermore, we only include in the analysis haloes whose center is located outside the virial radius of any other halo.} We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
 The database allows us to obtain the comoving positions, peculiar velocities and masses for all the halos in the simulation volume at $z=0$.
 The Hubble flow is taken into account to convert the relative velocities into physical velocities and allow for a comparison with observations.
+We have verified that the main conclusions of this paper hold in the case of halos defined by a FOF algorithm with a linking lentgth 0.17 times the interparticle distance.
 
-The constrained simulations we use in this Letter are part of the Constrained Local UniversE Simulations (CLUES) project whose main objective is to reproduce the large scale structure in the Local Universe as accurately as observations permit.
-The algorithm and observational constraints to construct the initial conditions are
-SG  I think the observations are much more accurate than our reconstructions, so it would be better to replace 
- ``observations permit.''  by ``possible.'', which includes both the limitations of the observations and the reconstraction algorithm.
- described in \citep{clues2010}.
+The constrained simulations we use in this Letter are part of the Constrained Local UniversE Simulations (CLUES) project whose main objective is to reproduce the large scale structure in the Local Universe as accurately as possible.
+The algorithm and observational constraints to construct the initial conditions are  described in \citep{clues2010}.
 We use three dark matter only simulations, each has a cubic volume of $64$\hMpc on a side, with the density field sampled with $1024^3$ particles.
 The cosmological density parameter is $\Omega_m=0.28$, the cosmological constant $\Omega_{\Lambda}=0.72$, the dimensionless Hubble parameter $h=0.73$, the spectral index of the primordial density perturbations $n=0.96$ and the power spectrum normalization $\sigma_{8}=0.817$, also consistent with WMAP 7th year data.
-The halo catalogs are constructed with the same FOF code used on the Bolshoi Simulation.
 
@@
 \item There cannot be halos more massive than $5\times 10^{12}$\hMsun within a radius of $2$\hMpc with respect to every object centre \citep{Karachentsev04,Anton09}.
-\item There cannot be halos more massive than $5\times 10^{13}$\hMsun with  a radius of $3$\hMpc with
-SG    within a radius of
- respect to every object centre \citep{Karachentsev04}.
+\item There cannot be halos more massive than $5\times 10^{13}$\hMsun within  a radius of $3$\hMpc with respect to every object centre \citep{Karachentsev04}.
 \end{enumerate}
 
-This sample in the Bolshoi simulation has $1256$ pairs.
-This is $\sim 30\%$ smaller than the the number of pairs we found in \cite{ForeroRomero2011}, which did not include the condition 5) in the construction of the sample.
+This sample in the Bolshoi simulation has $1923$ pairs.
 Additionally, there is a sample of three (3) pairs constructed from the three constrained realizations.
-These pairs fulfill all the above mentioned conditions and additionally are located in a place with the right distances with respect to the Virgo and Coma clusters in the simulation.
-Figure \ref{fig:distros} show the
-SG    Coma is outside of the 64 Mpc/h box
- distribution of separations and total pair mass computed from the $1256$ pairs in the Bolshoi simulation.
+These pairs fulfill all the above mentioned conditions and additionally are located in a place with the right distances with respect to the Virgo cluster in the simulation.
+Figure \ref{fig:distros} show the  distribution of separations and total pair mass computed from the $1923$ pairs in the Bolshoi simulation.
 
-{\bf Additional constraints will be set on this population in order to match the observational estimates for separation and total mass.
-The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}.}
+Additional constraints will be set on this population in order to match the observational estimates for separation and total mass.
+The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}.
 
@@
 Most of the LG-like pairs (left panel) in the Bolshoi Simulation and the pairs in the constrained simulation have radial and tangential velocities notably different from the observational constraints.
-The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -65\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram.
-There are $45$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations.
-This makes the LG-like pairs consistent with observations $\sim5$ times less likely to be found than a pair with velocities around $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$.
+The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram.
+There are $23$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations.
+This makes the LG-like pairs consistent with observations $\sim3$ times less likely to be found than a pair with velocities around $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$.
 
 From Figure \ref{fig:rt} it is also clear that there is a significant number of pairs with a high tangential-to-radial velocity ratio.
-The peak in the pair number density is located around a region of $f_{\rm t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 1$, while the observations suggest $f_{\rm t}<0.32$.
-In the Bolshoi Simulation, out of $1256$ LG-like pairs we find that $108$ pairs ($\sim 8\%$ of the total sample) have values for $f_{\rm t}$ consistent with that constraint.
+The peak in the pair number density is located around a region of $f_{\rm t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$, while the observations suggest $f_{\rm t}<0.32$.
+In the Bolshoi Simulation, out of $1923$ LG-like pairs we find that $242$ pairs ($\sim 12\%$ of the total sample) have values for $f_{\rm t}$ consistent with that constraint.
 {\bf The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.}
@@
 This change emphasizes the tension found from the full LG-like sample.
-The sample is smaller ($1256$ pairs are reduced to $180$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -36\pm 5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $10$ pairs around that value and zero pairs consistent with observations.
+The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm10$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm10$\kms with $10$ pairs around that value and zero pairs consistent with observations.
 This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to the separation and total mass conditions used to define the LG-like pairs.
-{\bf We note that this tighter selection also excludes 2 of the 3 pairs from the constrained simulations.}
+We note that this tighter selection also excludes 2 of the 3 pairs from the constrained simulations.
+
+NOTE: check below_here
 
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow_BDM.pdf}
 \caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow_BDM.pdf}
 \caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_narrow_BDM.pdf}
 \caption{{\rm \label{fig:lambda} Normalized histograms of the dimensionless spin parameter $\lambda$ for the pairs in the Bolshoi simulation and its infered values for the LG from the observational constraints from a Monte Carlo simulation.

### 2013-01-21 code/makeTempHistogramPlot.py
minor update in maketempHistogram to handle different font size in the legend

@@ -89,7 +89,8 @@ def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], clues_label='',xli
 
     if clues.size:
         axTemperature.scatter(clues[:,0], clues[:,1], color='white', edgecolors='black', s=200, label=clues_label)        
-        axTemperature.legend(loc=0, scatterpoints=1, fontsize=20)
+#        axTemperature.legend(loc=0, scatterpoints=1, fontsize=20)
+        axTemperature.legend(loc=0, scatterpoints=1, prop={'size':20})
 
     if X_field.size:
         print 'HOLA'

### 2013-01-22 paper/ms.tex
includes all corrections from Yehuda and Stefan

@@
 We define the Local Group (LG) to be composed by two dominant separate halos hosting the MW and the M31 galaxy fulfilling certain isolation criteria.
-We find that the most probable values for the tangential and radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-65\pm 5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm 5$\kms.
-Within a similar absolute uncertainty defined by observations the pairs centered around these values are $\sim5$ times more abundant than the pairs in the observational interval.
-Furthermore, we find that only $\sim8\%$ of the pairs show a ratio between its tangential and radial velocity consistent with observations.
+We find that the most probable values for the tangential and radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-60\pm 15$\kms and $v_{\rm tan, \Lambda CDM}=50\pm 5$\kms.
+Within a similar absolute uncertainty defined by observations the pairs centered around these values are $\sim3$ times more abundant than the pairs in the observational interval.
+Furthermore, we find that only $\sim12\%$ of the pairs show a ratio between its tangential and radial velocity consistent with observations.
 This tension increases if we make a narrower selection in the pair sample to match the observational constraints on pair separation and total mass.
@@
 With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun.
-{\bf  In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}.
-The halos are selected to have an overdensity of 200 times the critical overdensity.
-Furthermore, we only include in the analysis haloes whose center is located outside the virial radius of any other halo.} We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
+In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}.
+The halos are selected to have an overdensity of 200 times the critical density.
+Furthermore, we only include in the analysis haloes whose center is located outside the virial radius of any other halo.
+We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
 The database allows us to obtain the comoving positions, peculiar velocities and masses for all the halos in the simulation volume at $z=0$.
-The Hubble flow is taken into account to convert the relative velocities into physical velocities and allow for a comparison with observations.
+The positions and velocities of these haloes correspond to the average values of the $250$ most bound particles.
+The Hubble flow is taken into account to convert the peculiar velocities into physical velocities and allow for a comparison with observations.
 We have verified that the main conclusions of this paper hold in the case of halos defined by a FOF algorithm with a linking lentgth 0.17 times the interparticle distance.
@@
 \subsection{The LG-sample}
-Based on the FOF catalogs in the Bolshoi simulation we construct a halo pairs sample with masses and isolation properties consistent with the inferred properties for the MW and M31.
+Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with masses and isolation properties consistent with the inferred properties for the MW and M31.
 The characteristics we impose to define a LG-like halo pair are the following:
@@
 The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}.
+We refer to this sample as a reduced LG-like sample.
+The additional characteritiscs we impose in addition to the ones listed above are
+
+\begin{enumerate}
+\item The separation betwen the center of mass of the halos is in the range $700-800$kpc \citep{ribas05,vanderMarel08}.
+\item The total mass of the two halos is in the range $1-4\times 10^{12}$\Msun \citep{vanderMarel12}.
+\end{enumerate}
 
 \begin{table}
-\caption{Summary of the kinematic properties for M31 in the galactocentric frame as reported by \citep{vanderMarel12}. Values in parenthesis correspond to vector components}
+\caption{Summary of the kinematic properties for M31 in the galactocentric frame as reported by \citep{vanderMarel12}.
+Values in parenthesis correspond to vector components.
+$\sigma_{\bf x}$ represents the uncertainty on the components of vector ${\bf x}$}
 \begin{center}
@@
 In the Bolshoi Simulation, out of $1923$ LG-like pairs we find that $242$ pairs ($\sim 12\%$ of the total sample) have values for $f_{\rm t}$ consistent with that constraint.
-{\bf The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.}
+The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
 
@@
 This change emphasizes the tension found from the full LG-like sample.
-The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm10$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm10$\kms with $10$ pairs around that value and zero pairs consistent with observations.
+The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ pairs around that value and zero pairs consistent with observations.
 This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to the separation and total mass conditions used to define the LG-like pairs.
 We note that this tighter selection also excludes 2 of the 3 pairs from the constrained simulations.
-
-NOTE: check below_here
+Taking into account the results for the two different kinds of samples, we find that that the prefered values for the tangential and radial velocity components of halo pairs are bound by the values  $v_{\rm rad,\Lambda CDM} = -60\pm 15$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms.
 
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_narrow_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
 \caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
@@
 The rectangular panels show the histograms when only one of the velocity components is used to bin the data.
-The half ellipsoid
-SG ellipse  -- ellipsoid is 3D or higher dim
- corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.3$ km s$^{-1}$ \citep{vanderMarel12}.
+The half ellipse corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.4$ km s$^{-1}$ \citep{vanderMarel12}.
 The circles represent the positions of the pairs from the constrained simulations.
-The highest number density in the plot is obtained around $v_{\rm rad,\Lambda CDM} = 65\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 62\pm5$ \kms}
+The highest number density in the left panel is obtained around $v_{\rm rad,\Lambda CDM} = -70\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, for the right panel it is located around $v_{\rm rad,\Lambda CDM} = -50\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ }
 \label{fig:rt}
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_narrow_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
 \caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
 The panel distribution is the same as in Figure \ref{fig:rt}.
-The contour line shows the $1-\sigma$ uncertainty estimated through Monte Carlo techniques from the observational values summarized in Table 1.
-The poorer constraints on this plane are due
-SG  poorer than what? it is explained in the text, but for clarity I would prefer something like
-The constraints in this plot are less restrictive than in Fig. 2 because we have added ......
- to the added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$). }
+The contour line encloses $68\%$ of the Monte Carlo generated points to estimated the uncertainties from the observational values summarized in Table 1.
+The constraints in this plot are less restrictive than in Fig.\ref{fig:rt} due to added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$).
+}
 \label{fig:EJ}
@@
 
-{\bf This formulation has a clear theoretical advantage if one considers the angular momentum and the mechanical energy as quasi-conserved dynamical quantities.
+This formulation has a clear theoretical advantage if one considers the angular momentum and the mechanical energy as quasi-conserved dynamical quantities.
 This means that, after some formation time, these quantities do not significantly evolve as it is the case for the radial and tangential velocities.
-But there is an observational disadvantage.
-These quantities have to be derived from observations, increasing the uncertainties in their determination.}
+Conversely, there is an observational disadvantage.
+These quantities have to be derived from observations, increasing the uncertainties in their determination.
 
@@
 The closed line is the iso-density contour in the 2D histogram enclosing $68\%$ of the Monte Carlo generated pairs consistent with observations.
-The poor observational constraints on the $e_{\rm tot}-l_{\rm orb}$ plane in comparison to the radial-tangential velocity plane comes from the fact that we are adding up large uncertainties in three different aspects.
+The poor observational constraints on the $e_{\rm tot}-l_{\rm orb}$ plane in comparison to the radial-tangential velocity plane come from the fact that we are adding up large uncertainties in three different aspects.
 First, the tangential velocity has a $100\%$ uncertainty, making the angular momentum compatible with very low and high values.
@@
 Third, the uncertainty in the total mass of the Local Group is close to $20\%$, which has an impact on the potential energy.
+In Figure \ref{fig:EJ} the angular momentum does not reach a zero value at the 1-$\sigma$ as one could expect from Fig.
+\ref{fig:rt}, however it does reach a zero value when the isocontour encloses $80\%$ of the total number of generated points.
+The reason for this small difference is that our Monte Carlo simulation takes as a starting point the components of the 3D velocity.
+In turn, these values have uncertainties derived by MonteCarlo techniques, unlike the observational uncertainties directly inferred from the measured tangential and radial velocity components.
 
-In the right panel of Figure \ref{fig:EJ} we see that further selection of LG-like pairs to have similar separation
-SG I think better saying more precise, i.e.
-``that further reduction of the LG-like pairs following the procedure decribed for the right panel of Fig.2  makes them occupy a narrower region in the total energy.''
- and total mass makes them occupy a narrower region. This shows that there are preferred values for $e_{\rm
-SG I don't see this, looks like only e_tot is more constrained but not l_orb.
- tot}-l_{\rm orb}$.
-Under the assumption of quasi-conserved quantity it means that their formation and evolution conditions must be similar.
+In the right panel of Figure \ref{fig:EJ} we see that further reduction of the LG-like pairs following the procedure decribed for the right panel of Figure \ref{fig:EJ} makes them occupy a narrower region in the total energy and orbital angular momentum plane.
+We also see that $e_{\rm tot}$ becomes more constrained that $l_{\rm orb}$.
+Under the assumption of quasi-conserved quantity a narrower region in the $e_{\rm tot}-l_{\rm orb}$ plan implies that the formation and evolution conditions for the pairs in that region must be similar.
 In this case the agreement with observations is moderate.
-A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would clearly bring the observations in the region of low $l_{\rm orb}$, clearly below the average $\Lambda$CDM expectation.
+A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would clearly bring the observations in the region of low $l_{\rm orb}$, clearly outside the average $\Lambda$CDM expectation.
 
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_narrow_BDM.pdf}
-\caption{{\rm \label{fig:lambda} Normalized histograms of the dimensionless spin parameter $\lambda$ for the pairs in the Bolshoi simulation and its infered values for the LG from the observational constraints from a Monte Carlo simulation.
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM_narrow.pdf}
+\caption{{\rm \label{fig:lambda} Normalized histograms of Peeble's spin parameter $\lambda$ for the pairs in the Bolshoi simulation and its infered values for the LG from the observational constraints from a Monte Carlo simulation.
 The left panel includes all the LG-pairs while the right panel includes a reduced sample of pairs that match the separation and total mass infered from Local Group observations.
 The vertical lines with the white dots represent the values inferred for pairs in the constrained simulations.}}
-\caption{Spin parameter}
 \label{fig:lambda}
@@
 
-{\bf In complete analogy to the spin parameter used to describe dark matter halos, we define the same
-SG remove ``complete'',  there is an analogy but not a complete or incomplete analogy 
- parameter, $\lambda$, for halo pairs.
-This quantity combines into a single scalar the energy and the angular momentum allowing us to make an even more straightforward comparison between the results allowed for observations an the expectations in $\Lambda$CDM.
-The precise definition of the
-SG remove ``precise'',  an unprecise definition would be not a definition
- dimensionless $\lambda$ parameter is the following
-}
+The imensionless spin parameter $\lambda$ \citep{Peebles1971} is used here to characterize the dynamical state of the observed and simulated LGs.
+The parameter measures the dynamical role of the angular momentum in terms of the gravitational attraction and is defined by
 
@@
 
-{\bf where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.}
+where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.
 
-{\bf We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
+We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
 The results can be found in Figure \ref{fig:lambda}.
 The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
-}
+Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
+The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
 
-{\bf Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
-The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
-}
+The spin parameter $\lambda$ seems to be lower in the observed LG than the average expectation from $\Lambda$CDM.
+In observations the mean of the distribution is $\log_{10}(\lambda_{\rm LG})=-1.76$, while $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.51$ for the full LG-like sample and $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.38$ for the reduced sample.
+In the case where the pairs are restricted to have the same total mass and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are is close to the range define by observations, the disagreement in $\lambda$ comes mostly from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation.
 
-{\bf In terms of the reduced spin parameter $\lambda$ seems to be lower in the LG than expected from LCDM.
-In the case where the pairs are restricted to have the same total mass and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are by construction in the same range as observations.
-The disagreement must come from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation, but still they are broadly consistent due to the large observational uncertainties.}
-
-{\bf Left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations.
+The left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations with values $\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$.
 In this case all cases seem to be a better match to the observational expectations than to the $\Lambda$CDM pair population.
-In the case of tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\lambda$ value matches with the maximum in the $\lambda$ distribution infered from observations.}
+In the case of tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\log_{10}\lambda$ value of $1.65$ matches with the observational value $\log_{10}(\lambda_{\rm LG})=-1.76$.
 
@@
 
-{\bf In the three pairs from constrained simulations we find kinematics dominated by radial velocities.
+In the three pairs from constrained simulations we find kinematics dominated by radial velocities.
 However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in broad concordance with observations.
 There is only one pair that fullfills all the separation, total mass constraints and matches the most probable value for the dimensionless spin parameter $\lambda$ inferred from observations.
-}
 
-{\bf The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs. Reduced unceratainties in masses, distances and velocities can be included to produce an updated
-SG uncertainties
- estimate for this parameter. This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.}
+The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs.
+Reduced uncertainties in masses, distances and velocities can be included to produce an updated estimate for this parameter.
+This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.
 
-{\bf In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and total mechanical energy.
+In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and total mechanical energy.
 The values for the orbital angular momentum and energy, merged into the $\lambda$ spin parameter, are in mild disagreement with the observational constraints.
@@
 We consider that reducing the uncertainties in the observations relevant to this study are needed.
-Likewise, additional theoretical work to understand the origin and implications of this result deserve further study.}
+Likewise, additional theoretical work to understand the origin and implications of this result deserve further study.
 
@@
 J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de Investigaciones at Universidad de los Andes through its {\it Fondo de Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber Foundation through its fellowship administered by the International Astronomical Union.
+J.E.F-R also acknowledges early discussions with Alejandro Garc\'ia that motivated this work.
 
-The data and source code needed to reproduce the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
+The data and source code and instructions to replicate the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
+Thanks to the IPython community \citep{IPython}.
+Thanks to Jessica Kirkpatrick for releasing her Python code to make nice plots of 2D histograms.
 
-SG: Please add the sentence form the multidark database page:
 The MultiDark Database used in this paper and the web application providing online access to it were constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064.

### 2013-01-22 code/makeTempHistogramPlot.py
due credit to original author

@@ -13,6 +13,8 @@ def ellipse(ra,rb,ang,xcenter,ycenter):
     Y = rb*sin(theta_bins) + ycenter
     return X,Y
 
+#this function is defined was initially written by Jessica Kirkpatrick here:
+# http://jessresearch.blogspot.com/2012/04/pretty-plots-2d-histogram-with-1d.html
 def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], clues_label='',xlims=-99, ylims =-99 , \
                               nxbins = 50,nybins=50, bw=0, nbins=100,contours=1,sigma=1,line=1, 
                           xlabel='A', ylabel='B', mean_obs=0, sigma_obs=0, 

### 2013-01-30 paper/ms.tex
Yehuda's comments

@@
 \font\smcap=cmcsc10
+\newcommand{\YH}[1]{\textcolor{blue}{\bf YH: #1}}
+\newcommand{\NEW}[1]{\textcolor{blue}{  #1}}
+\newcommand{\SOUT}[1]{\textcolor{red}{\sout{  #1}}}
 
@@
 A new valuable piece of information in this issue is the recent observational determination of the proper-motion measurements of M31, which until recently had been out of reach \citep{vanderMarel12}. 
-This measurement opens up the possibility to study in detail the dynamics of the LG in a cosmological context.
+%YH:This measurement opens up the possibility to study in detail the dynamics of the LG in a cosmological context. 
+%YH
 The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km s$^{-1}$.
@@
 
-This Letter presents such study.
-We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar isolation criteria as inferred in the LG.
+This Ltter presents such study. We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar 
+%YH: isolation criteria 
+characteristics %YH
+as inferred in the LG.
 We quantify these results in terms of the number of pairs with given radial and tangential velocities in the galactocentric rest frame.
@@
 
-In addition we make use of three N-body simulations with constrained initial conditions which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc.
-The special feature of these simulations is that each volume features a pair of halos with the right isolation and relative positioning characteristics to be considered Local Group halos.
-Finally, we provide an interpretation of these results in terms of the large scale structure environment.
-We find what places in the cosmic web favor the presence of dark matter halos with kinematics similar to the LG.
+In addition we make use of three 
+%YH: N-body simulations with constrained initial conditions 
+constrained N-body simulations %YH 
+which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc.
+The special feature of these simulations is that each volume features a pair of halos with the right
+%YH: isolation and relative positioning 
+%YH
+characteristics to be considered 
+%YH: Local Group halos.  
+LG-like objects.%YH
+%YH:   
+%Finally, we provide an interpretation of these results in terms of the large scale structure environment. 
+%We find what places in the cosmic web favor the presence of dark matter halos with kinematics similar to the LG.
 
-This Letter is structured as follows.
-In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs and the method we use to characterize the cosmic web.
+This Letter is structured as follows. In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs. 
+%YH: and the method we use to characterize the cosmic web. 
+%YH
 In Section \ref{sec:results} we present the results for the dynamics in these pairs in terms of the tangential/radial velocities and the orbital angular momentum/mechanical energy.
@@
 \subsection{The LG-sample}
-Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with masses and isolation properties consistent with the inferred properties for the MW and M31.
-The characteristics we impose to define a LG-like halo pair are the following:
+Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with 
+%YH: Jaime you keep using 'masses and isolation properties', but your criteria include also the infall velocities, 
+%%hence they are dynamical in their nature.
+%masses and isolation 
+the dynamical %YH
+properties consistent with 
+%YH:
+%the inferred properties for 
+those of %YH
+the MW and M31. The 
+%YH:
+%characteristics 
+criteria %YH
+we impose to define a LG-like halo pair are the following:
 
@@
 
+%YH: Jaime, aren't you using a Virgo constraint?
+
 This sample in the Bolshoi simulation has $1923$ pairs.
@@
 
-Additional constraints will be set on this population in order to match the observational estimates for separation and total mass.
+%YH:
+%Additional constraints will be set on this population in order to match the observational estimates for separation and total mass. 
+%The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}. 
+%We refer to this sample as a reduced LG-like sample. The additional characteritiscs we impose in addition to the ones listed above are
 The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}. 
-We refer to this sample as a reduced LG-like sample.
-The additional characteritiscs we impose in addition to the ones listed above are
-
+A more restrictive subsample of LG-like objects has been constructed so as to fit obey the observational bounds on the masses and separation of the two main halos. These amount to: %YH
 \begin{enumerate}
-\item The separation betwen the center of mass of the halos is in the range $700-800$kpc \citep{ribas05,vanderMarel08}.
+\item The separation between the center of mass of the halos is in the range $700-800$kpc \citep{ribas05,vanderMarel08}.
 \item The total mass of the two halos is in the range $1-4\times 10^{12}$\Msun \citep{vanderMarel12}.
 \end{enumerate}
+%YH
+Note that only one constrained LG-like object is included in the subsample.%YH
 
@@
 The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram.
-There are $23$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations.
-This makes the LG-like pairs consistent with observations $\sim3$ times less likely to be found than a pair with velocities around $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$.
+%YH:
+%There are $23$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational 
+%constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations. 
+There are only 23 pairs out of the full sample, and only 8 pairs out of the subsample, within the subspace allowed by he Monte Carlo sampling of the observational errors, assuming normal distribution of the estimated errors.
+({\bf Jaime - please be more spacific here. How big is your sample? The $10^6$ that appears down the text?})
+ %YH
+%YH: Jaime, I find the following sentence to be vague and I suggest to remove it.
+%% You may comment about this finding in the Conclusions.
+%This makes the LG-like pairs consistent with observations $\sim3$ times less likely to be found than a pair with velocities around 
+% $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$. %YH
 
@@
 
-The right panel in Figure \ref{fig:rt} shows the results of making narrower selection of the LG-like pairs to separations between $700-800$ kpc and a total mass in the two halos to be in the range $1-4 \times 10^{12}$\Msun.
-This change emphasizes the tension found from the full LG-like sample.
-The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ pairs around that value and zero pairs consistent with observations.
-This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to the separation and total mass conditions used to define the LG-like pairs.
-We note that this tighter selection also excludes 2 of the 3 pairs from the constrained simulations.
-Taking into account the results for the two different kinds of samples, we find that that the prefered values for the tangential and radial velocity components of halo pairs are bound by the values  $v_{\rm rad,\Lambda CDM} = -60\pm 15$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms.
+%YH: Jaime, here you are repeating yourself.
+%The right panel in Figure \ref{fig:rt} shows the results of making narrower selection of the LG-like pairs to separations between $700-800$ kpc and a
+% total mass in the two halos to be in the range $1-4 \times 10^{12}$\Msun. 
+%This change emphasizes the tension found from the full LG-like sample. 
+The difference between the left and right panels of Figure \ref{fig:rt} reflects  the tension found from the full and the sub- sample of the LG-like objects. %YH
+%YH:
+%The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential 
+% velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ pairs around that value and 
+% zero pairs consistent with observations. 
+% This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to 
+% the separation and total mass conditions used to define the LG-like pairs.  We note that this tighter selection also excludes 2 of the 3 pairs from the 
+% constrained simulations. Taking into account the results for the two different kinds of samples, we find that that the prefered values for the tangential 
+%and radial velocity components of halo pairs are bound by the values  $v_{\rm rad,\Lambda CDM} = -60\pm 15$\kms and 
+% $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms. 
+The restrictive selection of objects in the subsample reduces the number of objects from  $1923$   to $158$ and their most probable configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval and none consistent with observations.
+The high eccentric orbit of the observed LG  constitutes an unlikely configurations for the $\Lambda$CDM LG-like objects. This holds for the full and the subsample of objects. 
+{\bf Jaime, can you translate the above statements to a quantitative  liklihood?}
+%YH
 
@@
 
-To gain additional physical insight we consider the two halos as points of mass $m_{\rm M31}$ and $m_{\rm MW}$.
-In the center of mass their orbit has an angular momentum $L=\mu|{{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and mechanical energy $E=\frac{1}{2}\mu |{\bf v}_{\rm M31}|^2 - G\mu M/|{\bf r}_{\rm M31}|$, where the total mass is $M=m_{\rm M31}+m_{\rm MW}$, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the gravitational constant.
-We express the results of the previous section in terms of the reduced angular momentum and energy $l_{orb}=L/\mu$ and $e_{\rm tot}=E/\mu$.
+%YH: I've rewritten this paragraph
+The two-body problem of point-like masses can serve as a proxy for the dynamics of the LG and be used as a tool for 
+studying the LG within the framework of the standard cosmological model.
+Within the model the dynamics of the LG is governed by the (center of mass) 
+reduced angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and reduced energy 
+$e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf r}_{\rm M31}|$, 
+where the total mass is $M=m_{\rm M31}+m_{\rm MW}$
+%YH:
+%, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ %YH
+and $G$ is the gravitational constant. %YH: No need to be so explicit.
+%We express the results of the previous section in terms of the reduced angular momentum and energy $l_{orb}=L/\mu$ and $e_{\rm tot}=E/\mu$. 
+%YH
 
@@
 This means that, after some formation time, these quantities do not significantly evolve as it is the case for the radial and tangential velocities.
-Conversely, there is an observational disadvantage.
-These quantities have to be derived from observations, increasing the uncertainties in their determination.
+%YH: the following statement doesn't make much sense. Everything needs to be derived from observations.
+%Conversely, there is an observational disadvantage. 
+%These quantities have to be derived from observations, increasing the uncertainties in their determination.
+A possible conflict between the likelihood of the LG in the $\Lambda$CDM model as viewed from the radial and tangential velocities as opposed to the angular momentum and energy perspectives would point to a particular phase on the LG two-body orbit.
+%YH
+%YH: merge the two paragraphs.
+%In Table \ref{table:1} we list the M31 position (${\bf r}_{\rm M31}$), M31 velocity (${\bf v}_{\rm M31}$) and total mass ($M$) that we use to estimate the
+% uncertainties in the $e_{\rm tot}-l_{\rm orb}$ plane. To do this we generate $10^6$ pairs with different values for these three variables using a 
+%Gaussian distribution for each vector component with a variance equal to their uncertainties ${\sigma_{\bf r}}_{\rm M31}$ and 
+%${\sigma_{\bf v}}_{\rm M31}$). For each pair we calculate its $l_{\rm orb}$ and $e_{\rm tot}$ values.  
+Given the masses, velocity and position of the observed MW and M31 (TAble \ref{table:1} ) the reduced energy and angular momentum are calculated.
+The Monte Carlo sampling is used to to estimate the uncertainty in these quantities.
+%YH
 
-In Table \ref{table:1} we list the M31 position (${\bf r}_{\rm M31}$), M31 velocity (${\bf v}_{\rm M31}$) and total mass ($M$) that we use to estimate the uncertainties in the $e_{\rm tot}-l_{\rm orb}$ plane.
-To do this we generate $10^6$ pairs with different values for these three variables using a Gaussian distribution for each vector component with a variance equal to their uncertainties ${\sigma_{\bf r}}_{\rm M31}$ and ${\sigma_{\bf v}}_{\rm M31}$)..
-For each pair we calculate its $l_{\rm orb}$ and $e_{\rm tot}$ values.
+%YH:
+%The results are summarized in Figure \ref{fig:EJ} following the same plotting conventions from Figure \ref{fig:rt}. The closed line is the iso-density 
+%contour in the 2D histogram enclosing $68\%$ of the Monte Carlo generated pairs consistent with observations.
+Figure \ref{fig:EJ} presents the reduced  energy and angular momentum distribution of the constrained and unconstrained LG-like  objects of the full and subsample, following the   plotting conventions from Figure \ref{fig:rt}.
+%YH - I think that the following two sentences are redundand. The ApJ reader know how the error budget is calculated.
+ %The poor observational constraints on the $e_{\rm tot}-l_{\ofrm orb}$ plane in comparison to the radial-tangential velocity plane come from the fact 
+ % that we are adding up large uncertainties in three different aspects. First, the tangential velocity has a $100\%$ uncertainty, making the angular 
+ % momentum compatible with very low and high values. Second, the uncertainty on the square of the norm of the velocity has an 
+ % uncertainty of $40\%$, having an impact on the kinetic energy. 
+ %Third, the uncertainty in the total mass of the Local Group is close to $20\%$, which has an impact on the potential energy. 
+% In Figure \ref{fig:EJ} the angular momentum does not reach a zero value at the 1-$\sigma$ as one could expect from Fig. \ref{fig:rt}, 
+%however it does reach a zero value when the isocontour encloses $80\%$ of the total number of generated points. The reason for this small 
+%difference is that our Monte Carlo simulation takes as a starting point the components of the 3D velocity. In turn, these values have uncertainties 
+%derived by MonteCarlo techniques, unlike the observational uncertainties directly inferred from the measured tangential and radial velocity 
+%components.
+The  1-$\sigma$ contour of Figure \ref{fig:EJ} does not cross the zero level, as can be naively expected from Figure \ref{fig:rt}.
+This does happen for the $\approx 80\%$ level contour.
+The small difference between the two presentations of the observational uncertainties stems from the Monte Carol sampling.
+%YH
 
-The results are summarized in Figure \ref{fig:EJ} following the same plotting conventions from Figure \ref{fig:rt}.
-The closed line is the iso-density contour in the 2D histogram enclosing $68\%$ of the Monte Carlo generated pairs consistent with observations.
-The poor observational constraints on the $e_{\rm tot}-l_{\rm orb}$ plane in comparison to the radial-tangential velocity plane come from the fact that we are adding up large uncertainties in three different aspects.
-First, the tangential velocity has a $100\%$ uncertainty, making the angular momentum compatible with very low and high values.
-Second, the uncertainty on the square of the norm of the velocity has an uncertainty of $40\%$, having an impact on the kinetic energy.
-Third, the uncertainty in the total mass of the Local Group is close to $20\%$, which has an impact on the potential energy.
-In Figure \ref{fig:EJ} the angular momentum does not reach a zero value at the 1-$\sigma$ as one could expect from Fig.
-\ref{fig:rt}, however it does reach a zero value when the isocontour encloses $80\%$ of the total number of generated points.
-The reason for this small difference is that our Monte Carlo simulation takes as a starting point the components of the 3D velocity.
-In turn, these values have uncertainties derived by MonteCarlo techniques, unlike the observational uncertainties directly inferred from the measured tangential and radial velocity components.
-
-In the right panel of Figure \ref{fig:EJ} we see that further reduction of the LG-like pairs following the procedure decribed for the right panel of Figure \ref{fig:EJ} makes them occupy a narrower region in the total energy and orbital angular momentum plane.
-We also see that $e_{\rm tot}$ becomes more constrained that $l_{\rm orb}$.
-Under the assumption of quasi-conserved quantity a narrower region in the $e_{\rm tot}-l_{\rm orb}$ plan implies that the formation and evolution conditions for the pairs in that region must be similar.
-In this case the agreement with observations is moderate.
-A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would clearly bring the observations in the region of low $l_{\rm orb}$, clearly outside the average $\Lambda$CDM expectation.
+%YH: The following paragraph is quite vague. Instead of words I suggest to provide numbers for the fraction of LG-like 
+%% simulated objects obeying the 1-sigma observational constraints.
+% In the right panel of Figure \ref{fig:EJ} we see that further reduction of the LG-like pairs following the procedure described for the right panel of 
+%Figure \ref{fig:EJ} makes them occupy a narrower region in the total energy and orbital angular momentum plane. We also see that $e_{\rm tot}$ 
+%becomes more constrained that $l_{\rm orb}$. Under the assumption of quasi-conserved quantity a narrower region in the $e_{\rm tot}-l_{\rm orb}$ 
+%plan implies that the formation and evolution conditions for the pairs in that region must be similar. In this case the agreement with observations is 
+%moderate. A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
+%clearly bring the observations in the region of low $l_{\rm orb}$, clearly outside the average $\Lambda$CDM expectation.
+Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and $l_{\rm orb}$ constraints are less restrictive compared with the radial and tangential velocity constraints.
+For the unconstrained full (sub-) sample XXX\% (YYY\%)  of the LG-like pairs obey the 1-$\sigma$ observational constraints.
+All the constrained LGs of the full and sub- samples are consistent with the 1-$sigma$ observational uncertainty.
+%YH
 
@@
 
-The imensionless spin parameter $\lambda$ \citep{Peebles1971} is used here to characterize the dynamical state of the observed and simulated LGs.
+The 
+%YH:
+%imensionless 
+dimensionless %YH
+spin parameter $\lambda$ \citep{Peebles1971} is used here to characterize the dynamical state of the observed and simulated LGs.
 The parameter measures the dynamical role of the angular momentum in terms of the gravitational attraction and is defined by
@@
 
-We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the MonteCrlo simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$.
-The results can be found in Figure \ref{fig:lambda}.
-The left panel shows the results from the full sample of isolated pairs while the right panel corresponds to the smaller pair sample with restrictions on the separation and total mass.
-Both panels in Figure \ref{fig:lambda} tell in a clearer way the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane.
+We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the 
+%YH:
+%MonteCrlo 
+Monte Carlo%YH
+simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$. 
+%YH:
+%The results can be found in Figure \ref{fig:lambda}. The left panel shows the results from the full sample of isolated pairs while the right panel 
+%corresponds to the smaller pair sample with restrictions on the separation and total mass. Both panels in Figure \ref{fig:lambda} tell in a clearer way 
+%the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane. 
+Figure \ref{fig:lambda} shows the likelihood distribution of the observational errors and the spin parameter of the LG-like pairs of the full (left panel) and the sub- (right panel) samples.
+The value of $\lambda$ of the constrained LGs is shown as well.
+%YH
 The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
 
-The spin parameter $\lambda$ seems to be lower in the observed LG than the average expectation from $\Lambda$CDM.
-In observations the mean of the distribution is $\log_{10}(\lambda_{\rm LG})=-1.76$, while $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.51$ for the full LG-like sample and $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.38$ for the reduced sample.
-In the case where the pairs are restricted to have the same total mass and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are is close to the range define by observations, the disagreement in $\lambda$ comes mostly from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the observations than the $\Lambda$CDM expectation.
-
-The left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations with values $\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$.
-In this case all cases seem to be a better match to the observational expectations than to the $\Lambda$CDM pair population.
-In the case of tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\log_{10}\lambda$ value of $1.65$ matches with the observational value $\log_{10}(\lambda_{\rm LG})=-1.76$.
+%YH:
+%The spin parameter $\lambda$ seems to be lower in the observed LG than the average expectation from $\Lambda$CDM. In observations the mean 
+%of the distribution is $\log_{10}(\lambda_{\rm LG})=-1.76$, while $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.51$ for the full LG-like sample 
+%and $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.38$ for the reduced sample. In the case where the pairs are restricted to have the same total mass 
+%and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are is close to the range define by observations, 
+%the disagreement in $\lambda$ comes mostly from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the %observations than the $\Lambda$CDM expectation.
+The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
+Inspection of Figure \ref{fig:lambda} shows that the $\lambda$ distribution is close to a lognormal one and thus resembles the distribution of the spin parameter of DM halos in simulation ({\bf Jaime - please provide a reference}) .
+The estimated value of $\log_{10}\lambda$ of the observed LG is $-1.76$, compared with the mean value of 
+ $\log_{10}\lambda$ of $-1.51$ for the full sample and $-1.38$ for the subsample. The values of the spin parameter of the 3 constrained LGs are
+%
+%
+%The left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations with values 
+$\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging to the sole constrained LG of the subsample. 
+%In this case all cases seem to be a better match to the observational expectations than to the $\Lambda$CDM pair population. In the case of 
+%tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\log_{10}\lambda$ value of 
+%$1.65$ matches with the observational value $\log_{10}(\lambda_{\rm LG})=-1.76$.
+It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology.
+This implies that the observed LG doesn't conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
+%YH
 
@@
 In the simulation we select a sample of halo pairs in the mass range $5\times 10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation conditions of the Local Group from other massive structures.
-While the observations show that M31 moves towards us with a highly radial velocity, the simulation shows that the most common configuration at $z=0$ has values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm5$ \kms.
+While the observations show that M31 moves towards us
+%YH:
+%with a highly radial velocity
+on a highly eccentric orbit,%YH
+ the simulation shows that the most common configuration at $z=0$ has values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm5$ \kms.
 
@@
 
-The MultiDark Database used in this paper and the web application providing online access to it were constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064.
+The MultiDark Database used in this paper 
+%YH:
+%and the web application providing online access to it were 
+was %YH
+constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064.
 The Bolshoi simulation was run on the NASA's Pleiades supercomputer at the NASA Ames Research Center.

### 2013-01-30 paper/ms.tex
responded to yehuda's comments

@@
 A new valuable piece of information in this issue is the recent observational determination of the proper-motion measurements of M31, which until recently had been out of reach \citep{vanderMarel12}.
-%YH:This measurement opens up the possibility to study in detail the dynamics of the LG in a cosmological context. 
-%YH
 The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km s$^{-1}$.
 Together with the values of the relative radial velocity of $v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative motion of the MW and Andromeda is consistent with a head-on collision.
-With this information it is possible to ask how common is this kinematic configuration in a $\Lambda$CDM Universe.
+With this information it is possible to quantify how common is this kinematic configuration in a $\Lambda$CDM Universe.
 
-This Ltter presents such study. We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar 
-%YH: isolation criteria 
-characteristics %YH
-as inferred in the LG.
+This Ltter presents such study.
+We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar  characteristics as inferred in the LG.
 We quantify these results in terms of the number of pairs with given radial and tangential velocities in the galactocentric rest frame.
@@
 
-In addition we make use of three 
-%YH: N-body simulations with constrained initial conditions 
-constrained N-body simulations %YH 
-which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc.
-The special feature of these simulations is that each volume features a pair of halos with the right
-%YH: isolation and relative positioning 
-%YH
-characteristics to be considered 
-%YH: Local Group halos.  
-LG-like objects.%YH
-%YH:   
-%Finally, we provide an interpretation of these results in terms of the large scale structure environment. 
-%We find what places in the cosmic web favor the presence of dark matter halos with kinematics similar to the LG.
+In addition we make use of three  constrained N-body simulations which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc.
+The special feature of these simulations is that each volume features a pair of halos with the right characteristics to be considered LG-like objects.
 
-This Letter is structured as follows. In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs. 
-%YH: and the method we use to characterize the cosmic web. 
-%YH
+This Letter is structured as follows.
+In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs.
 In Section \ref{sec:results} we present the results for the dynamics in these pairs in terms of the tangential/radial velocities and the orbital angular momentum/mechanical energy.
-In the same section we investigate what kind of large scale environment can foster the production of pairs with dynamics similar to those observed in the LG.
+In the same section we sumarize these dynamical results in terms of the dimensionless spin parameter of the pairs.
 Finally, in the last section we comment and conclude about the implications of these results in the context of the $\Lambda$CDM model.
@@
 \subsection{The LG-sample}
-Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with 
-%YH: Jaime you keep using 'masses and isolation properties', but your criteria include also the infall velocities, 
-%%hence they are dynamical in their nature.
-%masses and isolation 
-the dynamical %YH
-properties consistent with 
-%YH:
-%the inferred properties for 
-those of %YH
-the MW and M31. The 
-%YH:
-%characteristics 
-criteria %YH
-we impose to define a LG-like halo pair are the following:
+Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with the dynamical  properties consistent with  those of  the MW and M31.
+The  criteria  we impose to define a LG-like halo pair are the following:
 
@@
 %YH: Jaime, aren't you using a Virgo constraint?
+%J.: What is a Virgo constraint in terms of Bolshoi? I am using the same constraints that we used in the old DMassembly paper.
 
@@
 
-%YH:
-%Additional constraints will be set on this population in order to match the observational estimates for separation and total mass. 
-%The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}. 
-%We refer to this sample as a reduced LG-like sample. The additional characteritiscs we impose in addition to the ones listed above are
 The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}. 
-A more restrictive subsample of LG-like objects has been constructed so as to fit obey the observational bounds on the masses and separation of the two main halos. These amount to: %YH
+A more restrictive subsample of LG-like objects has been constructed so as to fit obey the observational bounds on the masses and separation of the two main halos. These amount to: 
 \begin{enumerate}
@@
 \end{enumerate}
-%YH
-Note that only one constrained LG-like object is included in the subsample.%YH
+
+Note that only one constrained LG-like object is included in the subsample.
 
@@
 $r_{\rm M31}$ &(kpc) & $770\pm 40$\\
+$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
+$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
 $M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\\hline
+
 \end{tabular}\\
@@
 The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram.
-%YH:
-%There are $23$ pairs in the region centered around those values within an interval width comparable with the uncertainties in the observational 
-%constraints ($\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations. 
-There are only 23 pairs out of the full sample, and only 8 pairs out of the subsample, within the subspace allowed by he Monte Carlo sampling of the observational errors, assuming normal distribution of the estimated errors.
-({\bf Jaime - please be more spacific here. How big is your sample? The $10^6$ that appears down the text?})
- %YH
-%YH: Jaime, I find the following sentence to be vague and I suggest to remove it.
-%% You may comment about this finding in the Conclusions.
-%This makes the LG-like pairs consistent with observations $\sim3$ times less likely to be found than a pair with velocities around 
-% $v_{\rm rad,\Lambda CDM}-v_{\rm tan,\Lambda CDM}$. %YH
+There are $23$ pairs out of the full LG-like pairs sample centered in those values within the same range of absolute uncertainties allowed by observations (i.e.
+$\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations.
+
+The left panel in the same Figure \ref{fig:rt} shows the results for a more restrictive selection of objects in the LG-like sample, reducing the number of objects from  $1923$   to $158$ and their most probable configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval and none consistent with observations.
+In this case none of the pairs are found in the interval allowed by observations.
+The high eccentric orbit of the observed LG  constitutes an unlikely configurations for the $\Lambda$CDM LG-like objects.
+This holds for the full and the subsample of objects.
+
+({\bf Jaime - please be more spacific here.
+How big is your sample?
+The $10^6$ that appears down the text?.
+Yehuda- I corrected the whole paragraph.
+The results here have nothing to do with the montecarlo.})
 
@@
 The peak in the pair number density is located around a region of $f_{\rm t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$, while the observations suggest $f_{\rm t}<0.32$.
-In the Bolshoi Simulation, out of $1923$ LG-like pairs we find that $242$ pairs ($\sim 12\%$ of the total sample) have values for $f_{\rm t}$ consistent with that constraint.
+In the Bolshoi Simulation, out of $1923$ LG-like pairs we find that $242$ pairs ($\sim 12\%$ of the sample) have values for $f_{\rm t}$ consistent with that constraint, for the more restrictive sample there are $13$ pairs within that constraint ($\sim 8\%$ of the sample).
 The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
 
-%YH: Jaime, here you are repeating yourself.
-%The right panel in Figure \ref{fig:rt} shows the results of making narrower selection of the LG-like pairs to separations between $700-800$ kpc and a
-% total mass in the two halos to be in the range $1-4 \times 10^{12}$\Msun. 
-%This change emphasizes the tension found from the full LG-like sample. 
-The difference between the left and right panels of Figure \ref{fig:rt} reflects  the tension found from the full and the sub- sample of the LG-like objects. %YH
-%YH:
-%The sample is smaller ($1923$ pairs are reduced to $158$) and the values corresponding to the highest density region in the radial-tangential 
-% velocity plane are  $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ pairs around that value and 
-% zero pairs consistent with observations. 
-% This shows that the atypical LG dynamics in $\Lambda$CDM and the low fraction of radially dominated velocities are robust results with respect to 
-% the separation and total mass conditions used to define the LG-like pairs.  We note that this tighter selection also excludes 2 of the 3 pairs from the 
-% constrained simulations. Taking into account the results for the two different kinds of samples, we find that that the prefered values for the tangential 
-%and radial velocity components of halo pairs are bound by the values  $v_{\rm rad,\Lambda CDM} = -60\pm 15$\kms and 
-% $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms. 
-The restrictive selection of objects in the subsample reduces the number of objects from  $1923$   to $158$ and their most probable configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval and none consistent with observations.
-The high eccentric orbit of the observed LG  constitutes an unlikely configurations for the $\Lambda$CDM LG-like objects. This holds for the full and the subsample of objects. 
-{\bf Jaime, can you translate the above statements to a quantitative  liklihood?}
-%YH
+The difference between the left and right panels of Figure \ref{fig:rt} reflects  the tension found from the full and the sub- sample of the LG-like objects.
+{\bf Yehuda, What tension???
+The tension comes between theory and observation, not between the two theoretical results.}%YH
 
@@
 
-%YH: I've rewritten this paragraph
-The two-body problem of point-like masses can serve as a proxy for the dynamics of the LG and be used as a tool for 
-studying the LG within the framework of the standard cosmological model.
-Within the model the dynamics of the LG is governed by the (center of mass) 
-reduced angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and reduced energy 
-$e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf r}_{\rm M31}|$, 
-where the total mass is $M=m_{\rm M31}+m_{\rm MW}$
-%YH:
-%, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ %YH
-and $G$ is the gravitational constant. %YH: No need to be so explicit.
-%We express the results of the previous section in terms of the reduced angular momentum and energy $l_{orb}=L/\mu$ and $e_{\rm tot}=E/\mu$. 
-%YH
+The two-body problem of point-like masses can serve as a proxy for the dynamics of the LG and be used as a tool for  studying the LG within the framework of the standard cosmological model.
+Within the model the dynamics of the LG is governed by the (center of mass)  reduced angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and reduced energy  $e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the gravitational constant.
 
@@
 %YH: the following statement doesn't make much sense. Everything needs to be derived from observations.
-%Conversely, there is an observational disadvantage. 
-%These quantities have to be derived from observations, increasing the uncertainties in their determination.
+%JF: I keep because it is important to mention that the observational uncertainties will be larger.
+Conversely, there is an observational disadvantage. These quantities are derived from very different kinds of observations, increasing the uncertainties in their final determination.
+
 A possible conflict between the likelihood of the LG in the $\Lambda$CDM model as viewed from the radial and tangential velocities as opposed to the angular momentum and energy perspectives would point to a particular phase on the LG two-body orbit.
-%YH
-%YH: merge the two paragraphs.
-%In Table \ref{table:1} we list the M31 position (${\bf r}_{\rm M31}$), M31 velocity (${\bf v}_{\rm M31}$) and total mass ($M$) that we use to estimate the
-% uncertainties in the $e_{\rm tot}-l_{\rm orb}$ plane. To do this we generate $10^6$ pairs with different values for these three variables using a 
-%Gaussian distribution for each vector component with a variance equal to their uncertainties ${\sigma_{\bf r}}_{\rm M31}$ and 
-%${\sigma_{\bf v}}_{\rm M31}$). For each pair we calculate its $l_{\rm orb}$ and $e_{\rm tot}$ values.  
 Given the masses, velocity and position of the observed MW and M31 (TAble \ref{table:1} ) the reduced energy and angular momentum are calculated.
@@
 
-%YH:
-%The results are summarized in Figure \ref{fig:EJ} following the same plotting conventions from Figure \ref{fig:rt}. The closed line is the iso-density 
-%contour in the 2D histogram enclosing $68\%$ of the Monte Carlo generated pairs consistent with observations.
 Figure \ref{fig:EJ} presents the reduced  energy and angular momentum distribution of the constrained and unconstrained LG-like  objects of the full and subsample, following the   plotting conventions from Figure \ref{fig:rt}.
-%YH - I think that the following two sentences are redundand. The ApJ reader know how the error budget is calculated.
- %The poor observational constraints on the $e_{\rm tot}-l_{\ofrm orb}$ plane in comparison to the radial-tangential velocity plane come from the fact 
- % that we are adding up large uncertainties in three different aspects. First, the tangential velocity has a $100\%$ uncertainty, making the angular 
- % momentum compatible with very low and high values. Second, the uncertainty on the square of the norm of the velocity has an 
- % uncertainty of $40\%$, having an impact on the kinetic energy. 
- %Third, the uncertainty in the total mass of the Local Group is close to $20\%$, which has an impact on the potential energy. 
-% In Figure \ref{fig:EJ} the angular momentum does not reach a zero value at the 1-$\sigma$ as one could expect from Fig. \ref{fig:rt}, 
-%however it does reach a zero value when the isocontour encloses $80\%$ of the total number of generated points. The reason for this small 
-%difference is that our Monte Carlo simulation takes as a starting point the components of the 3D velocity. In turn, these values have uncertainties 
-%derived by MonteCarlo techniques, unlike the observational uncertainties directly inferred from the measured tangential and radial velocity 
-%components.
 The  1-$\sigma$ contour of Figure \ref{fig:EJ} does not cross the zero level, as can be naively expected from Figure \ref{fig:rt}.
@@
 The small difference between the two presentations of the observational uncertainties stems from the Monte Carol sampling.
-%YH
 
-%YH: The following paragraph is quite vague. Instead of words I suggest to provide numbers for the fraction of LG-like 
-%% simulated objects obeying the 1-sigma observational constraints.
-% In the right panel of Figure \ref{fig:EJ} we see that further reduction of the LG-like pairs following the procedure described for the right panel of 
-%Figure \ref{fig:EJ} makes them occupy a narrower region in the total energy and orbital angular momentum plane. We also see that $e_{\rm tot}$ 
-%becomes more constrained that $l_{\rm orb}$. Under the assumption of quasi-conserved quantity a narrower region in the $e_{\rm tot}-l_{\rm orb}$ 
-%plan implies that the formation and evolution conditions for the pairs in that region must be similar. In this case the agreement with observations is 
-%moderate. A reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
-%clearly bring the observations in the region of low $l_{\rm orb}$, clearly outside the average $\Lambda$CDM expectation.
 Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and $l_{\rm orb}$ constraints are less restrictive compared with the radial and tangential velocity constraints.
-For the unconstrained full (sub-) sample XXX\% (YYY\%)  of the LG-like pairs obey the 1-$\sigma$ observational constraints.
-All the constrained LGs of the full and sub- samples are consistent with the 1-$sigma$ observational uncertainty.
-%YH
+For the unconstrained full (sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$ observational constraints.
+All the constrained LGs of the full
+%JF
+%and sub- 
+%Not correct, there are 2 constrained pairs that cannot even make it into the sub samples.
+samples are consistent with the 1-$sigma$ observational uncertainty.
+%JF
+%this point is important:
+However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
+bring these pairs and the $\Lambda$CDM expectation outside the 1-$\sigma$ observational uncertainty.
 
@@
 
-The 
-%YH:
-%imensionless 
-dimensionless %YH
+The dimensionless 
 spin parameter $\lambda$ \citep{Peebles1971} is used here to characterize the dynamical state of the observed and simulated LGs.
@@
 We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the 
-%YH:
-%MonteCrlo 
-Monte Carlo%YH
+Monte Carlo
 simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$. 
-%YH:
-%The results can be found in Figure \ref{fig:lambda}. The left panel shows the results from the full sample of isolated pairs while the right panel 
-%corresponds to the smaller pair sample with restrictions on the separation and total mass. Both panels in Figure \ref{fig:lambda} tell in a clearer way 
-%the story sketched in Figure \ref{fig:EJ} on the $e_{\rm tot}-l_{\rm orb}$ plane. 
 Figure \ref{fig:lambda} shows the likelihood distribution of the observational errors and the spin parameter of the LG-like pairs of the full (left panel) and the sub- (right panel) samples.
 The value of $\lambda$ of the constrained LGs is shown as well.
-%YH
+
 The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
 
-%YH:
-%The spin parameter $\lambda$ seems to be lower in the observed LG than the average expectation from $\Lambda$CDM. In observations the mean 
-%of the distribution is $\log_{10}(\lambda_{\rm LG})=-1.76$, while $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.51$ for the full LG-like sample 
-%and $\log_{10}(\lambda_{\rm \Lambda CDM})=-1.38$ for the reduced sample. In the case where the pairs are restricted to have the same total mass 
-%and separation as the observational constraints, the reduced energy $e_{\rm tot }$ in the simulation are is close to the range define by observations, 
-%the disagreement in $\lambda$ comes mostly from the values in the reduced orbital angular momentum $j_{\rm orb}$ which is lower in the %observations than the $\Lambda$CDM expectation.
 The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
@@
  $\log_{10}\lambda$ of $-1.51$ for the full sample and $-1.38$ for the subsample. The values of the spin parameter of the 3 constrained LGs are
+
+$\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging to the sole constrained LG of the subsample.
+
 % 
-%
-%The left panel in Figure \ref{fig:lambda} shows the results from the three constrained simulations with values 
-$\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging to the sole constrained LG of the subsample. 
-%In this case all cases seem to be a better match to the observational expectations than to the $\Lambda$CDM pair population. In the case of 
-%tigher restrictions on the separation and total mass only one pair in the constrained simulations is retained, its $\log_{10}\lambda$ value of 
-%$1.65$ matches with the observational value $\log_{10}(\lambda_{\rm LG})=-1.76$.
 It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
-This implies that the observed LG doesn't conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
+%JF. Yehuda, I don't understaind this sentence. 
+This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
 %YH
@@
 While the observations show that M31 moves towards us
-%YH:
-%with a highly radial velocity
 on a highly eccentric orbit,%YH

### 2013-02-08 paper/ms.tex
added line breaks

@@
 
-Recent observations constrained the tangential velocity of M31 with respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm 4.4$\kms \citep{vanderMarel12}.
-In this study we use a large volume high resolution N-body cosmological simulation to statistically study this kinematics in the context of the $\Lambda$CDM cosmology.
-We define the Local Group (LG) to be composed by two dominant separate halos hosting the MW and the M31 galaxy fulfilling certain isolation criteria.
-We find that the most probable values for the tangential and radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-60\pm 15$\kms and $v_{\rm tan, \Lambda CDM}=50\pm 5$\kms.
-Within a similar absolute uncertainty defined by observations the pairs centered around these values are $\sim3$ times more abundant than the pairs in the observational interval.
-Furthermore, we find that only $\sim12\%$ of the pairs show a ratio between its tangential and radial velocity consistent with observations.
-This tension increases if we make a narrower selection in the pair sample to match the observational constraints on pair separation and total mass.
-Expressing these results in terms of the angular momentum and mechanical energy per unit mass also reflects this disagreement albeit in a less strong fashion due to the larger uncertainties inferred for these physical quantities.
-From these quantities Peeble's spin parameter is evaluated and is used to characterize the LG with repect to other halo pairs in $\Lambda$CDM.
-We also include in our analysis results from constrained simulations.
-The characterisation results of this study suggest that the formation and evolution of the LG differ from the average $\Lambda$CDM halo pair.
+Recent observations constrained the tangential velocity of M31 with
+respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
+the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
+4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
+high resolution N-body cosmological simulation to statistically study
+this kinematics in the context of the $\Lambda$CDM cosmology. We
+define the Local Group (LG) to be composed by two dominant separate
+halos hosting the MW and the M31 galaxy fulfilling certain isolation
+criteria. We find that the most probable values for the tangential and
+radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-60\pm
+15$\kms and $v_{\rm tan, \Lambda CDM}=50\pm 5$\kms. Within a similar
+absolute uncertainty defined by observations the pairs centered around
+these values are $\sim3$ times more abundant than the pairs in the
+observational interval. Furthermore, we find that only $\sim12\%$ of
+the pairs show a ratio between its tangential and radial velocity
+consistent with observations. This tension increases if we make a
+narrower selection in the pair sample to match the observational
+constraints on pair separation and total mass. Expressing these
+results in terms of the angular momentum and mechanical energy per
+unit mass also reflects this disagreement albeit in a less strong
+fashion due to the larger uncertainties inferred for these physical
+quantities. From these quantities Peeble's spin parameter is evaluated
+and is used to characterize the LG with repect to other halo pairs in
+$\Lambda$CDM. We also include in our analysis results from constrained
+simulations. The characterisation results of this study suggest that
+the formation and evolution of the LG differ from the average
+$\Lambda$CDM halo pair. 
 \end{abstract}
@@
 
-The Milky Way (MW) and Andromeda galaxy (M31) are the dominant galaxies in the Local Group (LG).
-Astronomical observations of their mass distribution impose constraints on the standard cosmological model.
-The satellite overabundance problem \citep{Klypin99,Moore99}, tidal disruption features \citep{pandas09} and the disk dominated morphology \citep{Kazantzidis2008} are examples on of how LG studies are linked to the cosmological context.
-Detailed studies on the Magellanic Clouds dynamics and their possible link to M31 add to the interest of understanding the details of the LG kinematics and dynamics \citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}.
-However, a general concern in the use of the LG as a tool for near-field cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG regarding the properties of interest \citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.
+The Milky Way (MW) and Andromeda galaxy (M31) are the dominant
+galaxies in the Local Group (LG). Astronomical observations of their
+mass distribution impose constraints on the standard cosmological
+model. The satellite overabundance problem \citep{Klypin99,Moore99},
+tidal disruption features \citep{pandas09} and the disk dominated
+morphology \citep{Kazantzidis2008} are examples on of how LG studies
+are linked to the cosmological context. Detailed studies on the
+Magellanic Clouds dynamics and their possible link to M31 add to the
+interest of understanding the details of the LG kinematics and
+dynamics
+\citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}. However,
+a general concern in the use of the LG as a tool for near-field
+cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG
+regarding the properties of interest
+\citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.
 
-A new valuable piece of information in this issue is the recent observational determination of the proper-motion measurements of M31, which until recently had been out of reach \citep{vanderMarel12}.
-The reported measurements set an upper bound for the tangential velocity of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km s$^{-1}$.
-Together with the values of the relative radial velocity of $v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative motion of the MW and Andromeda is consistent with a head-on collision.
-With this information it is possible to quantify how common is this kinematic configuration in a $\Lambda$CDM Universe.
+A new valuable piece of information in this issue is the recent
+observational determination of the proper-motion measurements of M31,
+which until recently had been out of reach \citep{vanderMarel12}.  The
+reported measurements set an upper bound for the tangential velocity
+of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km
+s$^{-1}$. Together with the values of the relative radial velocity of
+$v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative
+motion of the MW and Andromeda is consistent with a head-on
+collision. With this information it is possible to quantify how common
+is this kinematic configuration in a $\Lambda$CDM Universe.
 
-This Ltter presents such study.
-We use a large volume, high resolution dark matter only N-body simulation in the concordance $\Lambda$CDM cosmology to find a set of halo pairs with similar  characteristics as inferred in the LG.
-We quantify these results in terms of the number of pairs with given radial and tangential velocities in the galactocentric rest frame.
-We also find the pairs that are consistent with a head on collision in terms of the ratio of the radial to tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.3$ and present these results in terms of the reduced angular momentum and mechanical energy.
+This Ltter presents such study. We use a large volume, high resolution
+dark matter only N-body simulation in the concordance $\Lambda$CDM
+cosmology to find a set of halo pairs with similar  characteristics as
+inferred in the LG. We quantify these results in terms of the number
+of pairs with given radial and tangential velocities in the
+galactocentric rest frame. We also find the pairs that are consistent
+with a head on collision in terms of the ratio of the radial to
+tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.3$ and
+present these results in terms of the reduced angular momentum and
+mechanical energy.
 
-In addition we make use of three  constrained N-body simulations which are constructed to reproduce the observed large scale structure of the Local Universe on scales of a few tens of Mpc.
-The special feature of these simulations is that each volume features a pair of halos with the right characteristics to be considered LG-like objects.
+In addition we make use of three  constrained N-body simulations which
+are constructed to reproduce the observed large scale structure of the
+Local Universe on scales of a few tens of Mpc. The special feature of
+these simulations is that each volume features a pair of halos with
+the right characteristics to be considered LG-like objects.
 
-This Letter is structured as follows.
-In the next section we present the N-body simulations, the criteria we use to select LG-like halo pairs.
-In Section \ref{sec:results} we present the results for the dynamics in these pairs in terms of the tangential/radial velocities and the orbital angular momentum/mechanical energy.
-In the same section we sumarize these dynamical results in terms of the dimensionless spin parameter of the pairs.
-Finally, in the last section we comment and conclude about the implications of these results in the context of the $\Lambda$CDM model.
+This Letter is structured as follows. In the next section we present
+the N-body simulations, the criteria we use to select LG-like halo
+pairs. In Section \ref{sec:results} we present the results for the
+dynamics in these pairs in terms of the tangential/radial velocities
+and the orbital angular momentum/mechanical energy. In the same
+section we sumarize these dynamical results in terms of the
+dimensionless spin parameter of the pairs. Finally, in the last
+section we comment and conclude about the implications of these
+results in the context of the $\Lambda$CDM model.
 
@@
 \includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
-\caption{\label{fig:distros} Distributions of separations and total halo mass (Milky Way and M31) in the LG-like pairs selected from the Bolshoi Simulation.}
+\caption{\label{fig:distros} Distributions of separations and total
+  halo mass (Milky Way and M31) in the LG-like pairs selected from the
+  Bolshoi Simulation.} 
 \end{center}
@@
 
-The Bolshoi simulation follows the non-linear evolution of the dark matter density field using N-body techniques.
-The simulation has a cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$ particles.
-The cosmological parameters used in the simulation are $\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and $n=0.95$, corresponding to the matter density, vacuum energy density, the normalization of the power spectrum, the dimensionless Hubble constant and the index of the slope in the initial power spectrum.
-This set of parameters is compatible with the analysis  of the seventh year of data from the Wilkinson Microwave Anisotropy Probe (WMAP) \citep{Jarosik2011}.
-A detailed description of this simulation can be found in \citep{Bolshoi}.
+The Bolshoi simulation follows the non-linear evolution of the dark
+matter density field using N-body techniques. The simulation has a
+cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$
+particles. The cosmological parameters used in the simulation are
+$\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and
+$n=0.95$, corresponding to the matter density, vacuum energy density,
+the normalization of the power spectrum, the dimensionless Hubble
+constant and the index of the slope in the initial power
+spectrum. This set of parameters is compatible with the analysis  of
+the seventh year of data from the Wilkinson Microwave Anisotropy Probe
+(WMAP) \citep{Jarosik2011}. A detailed description of this simulation
+can be found in \citep{Bolshoi}.
 
-With these parameters the mass per particle is $m_{p}=1.4\times 10^{8}$\hMsun.
-In this paper we use algorithms obtained through the Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}.
-The halos are selected to have an overdensity of 200 times the critical density.
-Furthermore, we only include in the analysis haloes whose center is located outside the virial radius of any other halo.
-We have obtained the data through the public available Multidark database \footnote{{\tt http://www.multidark.org/MultiDark/}} \citep{2011arXiv1109.0003R}.
-The database allows us to obtain the comoving positions, peculiar velocities and masses for all the halos in the simulation volume at $z=0$.
-The positions and velocities of these haloes correspond to the average values of the $250$ most bound particles.
-The Hubble flow is taken into account to convert the peculiar velocities into physical velocities and allow for a comparison with observations.
-We have verified that the main conclusions of this paper hold in the case of halos defined by a FOF algorithm with a linking lentgth 0.17 times the interparticle distance.
+With these parameters the mass per particle is $m_{p}=1.4\times
+10^{8}$\hMsun. In this paper we use algorithms obtained through the
+Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}. The halos are
+selected to have an overdensity of 200 times the critical
+density. Furthermore, we only include in the analysis haloes whose
+center is located outside the virial radius of any other halo.  We
+have obtained the data through the public available Multidark
+database \footnote{{\tt http://www.multidark.org/MultiDark/}}
+\citep{2011arXiv1109.0003R}. The database allows us to obtain the
+comoving positions, peculiar velocities and masses for all the halos
+in the simulation volume at $z=0$. The positions and velocities of
+these haloes correspond to the average values of the $250$ most bound
+particles. The Hubble flow is taken into account to convert the
+peculiar velocities into physical velocities and allow for a
+comparison with observations. We have verified that the main
+conclusions of this paper hold in the case of halos defined by a FOF
+algorithm with a linking lentgth 0.17 times the interparticle
+distance.
 
-The constrained simulations we use in this Letter are part of the Constrained Local UniversE Simulations (CLUES) project whose main objective is to reproduce the large scale structure in the Local Universe as accurately as possible.
-The algorithm and observational constraints to construct the initial conditions are  described in \citep{clues2010}.
-We use three dark matter only simulations, each has a cubic volume of $64$\hMpc on a side, with the density field sampled with $1024^3$ particles.
-The cosmological density parameter is $\Omega_m=0.28$, the cosmological constant $\Omega_{\Lambda}=0.72$, the dimensionless Hubble parameter $h=0.73$, the spectral index of the primordial density perturbations $n=0.96$ and the power spectrum normalization $\sigma_{8}=0.817$, also consistent with WMAP 7th year data.
+The constrained simulations we use in this Letter are part of the
+Constrained Local UniversE Simulations (CLUES) project whose main
+objective is to reproduce the large scale structure in the Local
+Universe as accurately as possible. The algorithm and observational
+constraints to construct the initial conditions are  described in
+\citep{clues2010}.  We use three dark matter only simulations, each
+has a cubic volume of $64$\hMpc on a side, with the density field
+sampled with $1024^3$ particles. The cosmological density parameter is
+$\Omega_m=0.28$, the cosmological constant $\Omega_{\Lambda}=0.72$,
+the dimensionless Hubble parameter $h=0.73$, the spectral index of the
+primordial density perturbations $n=0.96$ and the power spectrum
+normalization $\sigma_{8}=0.817$, also consistent with WMAP 7th year
+data.
 
 \subsection{The LG-sample}
-Based on the BDM catalogs in the Bolshoi simulation we construct a halo pair sample with the dynamical  properties consistent with  those of  the MW and M31.
-The  criteria  we impose to define a LG-like halo pair are the following:
+Based on the BDM catalogs in the Bolshoi simulation we construct a
+halo pair sample with the dynamical  properties consistent with  those
+of  the MW and M31. The  criteria  we impose to define a LG-like halo
+pair are the following:
 
 \begin{enumerate}
-\item Each halo has a mass in the range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$.
-\item With respect to each halo, there cannot be any other halo within the mass range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$ closer than its partner.
-It means that there cannot be ambiguity on the identity of the pair members.
-\item The relative radial velocity between the two halos is negative \citep{vanderMarel12}.
-\item The distance between the center of mass of the halos must be less than $0.7$\hMpc \citep{ribas05,vanderMarel08}.
-\item There cannot be halos more massive than $5\times 10^{12}$\hMsun within a radius of $2$\hMpc with respect to every object centre \citep{Karachentsev04,Anton09}.
-\item There cannot be halos more massive than $5\times 10^{13}$\hMsun within  a radius of $3$\hMpc with respect to every object centre \citep{Karachentsev04}.
+\item Each halo has a mass in the range $5\times10^{11}\hMsun
+  <M_{h}<5\times 10^{12}\hMsun$. 
+\item With respect to each halo, there cannot be any other halo within
+  the mass range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$
+  closer than its partner. It means that there cannot be ambiguity on
+  the identity of the pair members. 
+\item The relative radial velocity between the two halos is negative
+  \citep{vanderMarel12}. 
+\item The distance between the center of mass of the halos must be
+  less than $0.7$\hMpc \citep{ribas05,vanderMarel08}. 
+\item There cannot be halos more massive than $5\times 10^{12}$\hMsun
+  within a radius of $2$\hMpc with respect to every object centre
+  \citep{Karachentsev04,Anton09}. 
+\item There cannot be halos more massive than $5\times 10^{13}$\hMsun
+  within  a radius of $3$\hMpc with respect to every object centre
+  \citep{Karachentsev04}. 
 \end{enumerate}
@@
 
-This sample in the Bolshoi simulation has $1923$ pairs.
-Additionally, there is a sample of three (3) pairs constructed from the three constrained realizations.
-These pairs fulfill all the above mentioned conditions and additionally are located in a place with the right distances with respect to the Virgo cluster in the simulation.
-Figure \ref{fig:distros} show the  distribution of separations and total pair mass computed from the $1923$ pairs in the Bolshoi simulation.
+This sample in the Bolshoi simulation has $1923$ pairs. Additionally,
+there is a sample of three (3) pairs constructed from the three
+constrained realizations. These pairs fulfill all the above mentioned
+conditions and additionally are located in a place with the right
+distances with respect to the Virgo cluster in the simulation. Figure
+\ref{fig:distros} show the  distribution of separations and total pair
+mass computed from the $1923$ pairs in the Bolshoi simulation.
 
-The full observational characteristics that we take in this Letter for the MW-M31 pair are listed in Table \ref{table:1}. 
-A more restrictive subsample of LG-like objects has been constructed so as to fit obey the observational bounds on the masses and separation of the two main halos. These amount to: 
+The full observational characteristics that we take in this Letter for
+the MW-M31 pair are listed in Table \ref{table:1}.  
+A more restrictive subsample of LG-like objects has been constructed
+so as to fit obey the observational bounds on the masses and
+separation of the two main halos. These amount to:   
 \begin{enumerate}
@@
 \vspace{1mm}
-Note: the observational uncertainties in the position vector correspond to a $5\%$ in each component consistent with the uncertainties in the distance \citep[see references in][]{vanderMarel08}
+Note: the observational uncertainties in the position vector
+correspond to a $5\%$ in each component consistent with the
+uncertainties in the distance \citep[see references
+  in][]{vanderMarel08} 
 \end{center}
@@
 
-Figure \ref{fig:rt} summarizes the central finding of this Letter.
-Most of the LG-like pairs (left panel) in the Bolshoi Simulation and the pairs in the constrained simulation have radial and tangential velocities notably different from the observational constraints.
-The most probable velocities in the simulation are around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect the minimum grid size needed to obtain robust statistics from the 2D histogram.
-There are $23$ pairs out of the full LG-like pairs sample centered in those values within the same range of absolute uncertainties allowed by observations (i.e.
-$\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs in the interval allowed by observations.
+Figure \ref{fig:rt} summarizes the central finding of this
+Letter. Most of the LG-like pairs (left panel) in the Bolshoi
+Simulation and the pairs in the constrained simulation have radial and
+tangential velocities notably different from the observational
+constraints.  The most probable velocities in the simulation are
+around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda
+  CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect
+the minimum grid size needed to obtain robust statistics from the 2D
+histogram. There are $23$ pairs out of the full LG-like pairs sample
+centered in those values within the same range of absolute
+uncertainties allowed by observations (i.e. $\sigma_{\rm tan}=17$\kms
+and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs
+in the interval allowed by observations.
 
-The left panel in the same Figure \ref{fig:rt} shows the results for a more restrictive selection of objects in the LG-like sample, reducing the number of objects from  $1923$   to $158$ and their most probable configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval and none consistent with observations.
-In this case none of the pairs are found in the interval allowed by observations.
-The high eccentric orbit of the observed LG  constitutes an unlikely configurations for the $\Lambda$CDM LG-like objects.
-This holds for the full and the subsample of objects.
+The left panel in the same Figure \ref{fig:rt} shows the results for a
+more restrictive selection of objects in the LG-like sample, reducing
+the number of objects from  $1923$   to $158$ and their most probable
+configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm
+  tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval
+and none consistent with observations. In this case none of the pairs
+are found in the interval allowed by observations. The high eccentric
+orbit of the observed LG  constitutes an unlikely configurations for
+the $\Lambda$CDM LG-like objects. This holds for the full and the
+subsample of objects.
 
-({\bf Jaime - please be more spacific here.
-How big is your sample?
-The $10^6$ that appears down the text?.
-Yehuda- I corrected the whole paragraph.
-The results here have nothing to do with the montecarlo.})
+From Figure \ref{fig:rt} it is also clear that there is a significant
+number of pairs with a high tangential-to-radial velocity ratio. The
+peak in the pair number density is located around a region of $f_{\rm
+  t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$,
+while the observations suggest $f_{\rm t}<0.32$. In the Bolshoi
+Simulation, out of $1923$ LG-like pairs we find that $242$ pairs
+($\sim 12\%$ of the sample) have values for $f_{\rm t}$ consistent
+with that constraint, for the more restrictive sample there are $13$
+pairs within that constraint ($\sim 8\%$ of the sample). The three
+pairs from the constrained realizations are also outside the region of
+the most probable values in a $\Lambda$CDM cosmology with
+tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
 
-From Figure \ref{fig:rt} it is also clear that there is a significant number of pairs with a high tangential-to-radial velocity ratio.
-The peak in the pair number density is located around a region of $f_{\rm t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$, while the observations suggest $f_{\rm t}<0.32$.
-In the Bolshoi Simulation, out of $1923$ LG-like pairs we find that $242$ pairs ($\sim 12\%$ of the sample) have values for $f_{\rm t}$ consistent with that constraint, for the more restrictive sample there are $13$ pairs within that constraint ($\sim 8\%$ of the sample).
-The three pairs from the constrained realizations are also outside the region of the most probable values in a $\Lambda$CDM cosmology with tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
-
-The difference between the left and right panels of Figure \ref{fig:rt} reflects  the tension found from the full and the sub- sample of the LG-like objects.
-{\bf Yehuda, What tension???
-The tension comes between theory and observation, not between the two theoretical results.}%YH
+The difference between the left and right panels of Figure
+\ref{fig:rt} reflects  the tension found from the full and the sub-
+sample of the LG-like objects. {\bf Yehuda, What tension??? The
+  tension comes between theory and observation, not between the two
+  theoretical results.}
 
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
-\caption{Histograms of the radial and tangential velocities for LG-like halo pairs in the Bolshoi simulation.
-The left panel corresponds to the general LG-like pair simple, the right panel corresponds to the subsample with cuts on pair separation and total mass, matching the observational constraints.
-The squared panel shows as a shaded histogram the number of pairs on the radial and tangential velocities plane.
-The rectangular panels show the histograms when only one of the velocity components is used to bin the data.
-The half ellipse corresponds to the observed observational constraints for M31 in the Galactocentric rest frame: $v_{\rm rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.4$ km s$^{-1}$ \citep{vanderMarel12}.
-The circles represent the positions of the pairs from the constrained simulations.
-The highest number density in the left panel is obtained around $v_{\rm rad,\Lambda CDM} = -70\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, for the right panel it is located around $v_{\rm rad,\Lambda CDM} = -50\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ }
+\caption{Histograms of the radial and tangential velocities for
+  LG-like halo pairs in the Bolshoi simulation. The left panel
+  corresponds to the general LG-like pair simple, the right panel
+  corresponds to the subsample with cuts on pair separation and total
+  mass, matching the observational constraints. The squared panel
+  shows as a shaded histogram the number of pairs on the radial and
+  tangential velocities plane. The rectangular panels show the
+  histograms when only one of the velocity components is used to bin
+  the data. The half ellipse corresponds to the observed observational
+  constraints for M31 in the Galactocentric rest frame: $v_{\rm
+    rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.4$ km s$^{-1}$
+  \citep{vanderMarel12}. The circles represent the positions of the
+  pairs from the constrained simulations. The highest number density
+  in the left panel is obtained around $v_{\rm rad,\Lambda CDM} =
+  -70\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, for the
+  right panel it is located around $v_{\rm rad,\Lambda CDM} = -50\pm5$
+  \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ } 
 \label{fig:rt}
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
-\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$) and mechanical energy ($e_{\rm tot}$) per unit of reduced mass calculated considering the halos as point masses.
-The panel distribution is the same as in Figure \ref{fig:rt}.
-The contour line encloses $68\%$ of the Monte Carlo generated points to estimated the uncertainties from the observational values summarized in Table 1.
-The constraints in this plot are less restrictive than in Fig.\ref{fig:rt} due to added uncertainties on the tangential velocity ($100\%$), the square of the norm of the velocity ($40\%$) and the total mass of the two halos ($20\%$).
-}
+\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$)
+  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
+  calculated considering the halos as point masses. The panel
+  distribution is the same as in Figure \ref{fig:rt}. The contour line
+  encloses $68\%$ of the Monte Carlo generated points to estimated the
+  uncertainties from the observational values summarized in Table
+  1. The constraints in this plot are less restrictive than in
+  Fig.\ref{fig:rt} due to added uncertainties on the tangential
+  velocity ($100\%$), the square of the norm of the velocity ($40\%$)
+  and the total mass of the two halos ($20\%$). } 
 \label{fig:EJ}
@@
 
-The two-body problem of point-like masses can serve as a proxy for the dynamics of the LG and be used as a tool for  studying the LG within the framework of the standard cosmological model.
-Within the model the dynamics of the LG is governed by the (center of mass)  reduced angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and reduced energy  $e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$, the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the gravitational constant.
+The two-body problem of point-like masses can serve as a proxy for the
+dynamics of the LG and be used as a tool for  studying the LG within
+the framework of the standard cosmological model. Within the model the
+dynamics of the LG is governed by the (center of mass)  reduced
+angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
+reduced energy  $e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
+  r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$,
+the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the
+gravitational constant.
 
-This formulation has a clear theoretical advantage if one considers the angular momentum and the mechanical energy as quasi-conserved dynamical quantities.
-This means that, after some formation time, these quantities do not significantly evolve as it is the case for the radial and tangential velocities.
-%YH: the following statement doesn't make much sense. Everything needs to be derived from observations.
-%JF: I keep because it is important to mention that the observational uncertainties will be larger.
-Conversely, there is an observational disadvantage. These quantities are derived from very different kinds of observations, increasing the uncertainties in their final determination.
+This formulation has a clear theoretical advantage if one considers
+the angular momentum and the mechanical energy as quasi-conserved
+dynamical quantities. This means that, after some formation time,
+these quantities do not significantly evolve as it is the case for the
+radial and tangential velocities.
 
-A possible conflict between the likelihood of the LG in the $\Lambda$CDM model as viewed from the radial and tangential velocities as opposed to the angular momentum and energy perspectives would point to a particular phase on the LG two-body orbit.
-Given the masses, velocity and position of the observed MW and M31 (TAble \ref{table:1} ) the reduced energy and angular momentum are calculated.
-The Monte Carlo sampling is used to to estimate the uncertainty in these quantities.
-%YH
+Conversely, there is an observational disadvantage. These quantities
+are derived from very different kinds of observations, increasing the
+uncertainties in their final determination.
 
-Figure \ref{fig:EJ} presents the reduced  energy and angular momentum distribution of the constrained and unconstrained LG-like  objects of the full and subsample, following the   plotting conventions from Figure \ref{fig:rt}.
-The  1-$\sigma$ contour of Figure \ref{fig:EJ} does not cross the zero level, as can be naively expected from Figure \ref{fig:rt}.
-This does happen for the $\approx 80\%$ level contour.
-The small difference between the two presentations of the observational uncertainties stems from the Monte Carol sampling.
+A possible conflict between the likelihood of the LG in the
+$\Lambda$CDM model as viewed from the radial and tangential velocities
+as opposed to the angular momentum and energy perspectives would point
+to a particular phase on the LG two-body orbit. Given the masses,
+velocity and position of the observed MW and M31 (TAble \ref{table:1}
+) the reduced energy and angular momentum are calculated. The Monte
+Carlo sampling is used to to estimate the uncertainty in these
+quantities.
 
-Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and $l_{\rm orb}$ constraints are less restrictive compared with the radial and tangential velocity constraints.
-For the unconstrained full (sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$ observational constraints.
-All the constrained LGs of the full
-%JF
-%and sub- 
-%Not correct, there are 2 constrained pairs that cannot even make it into the sub samples.
+Figure \ref{fig:EJ} presents the reduced  energy and angular momentum
+distribution of the constrained and unconstrained LG-like  objects of
+the full and subsample, following the   plotting conventions from
+Figure \ref{fig:rt}.  The  1-$\sigma$ contour of Figure \ref{fig:EJ}
+does not cross the zero level, as can be naively expected from Figure
+\ref{fig:rt}. This does happen for the $\approx 80\%$ level
+contour. The small difference between the two presentations of the
+observational uncertainties stems from the Monte Carol sampling.
+
+Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
+$l_{\rm orb}$ constraints are less restrictive compared with the
+radial and tangential velocity constraints. For the unconstrained full
+(sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$
+observational constraints. All the constrained LGs of the full
 samples are consistent with the 1-$sigma$ observational uncertainty.
-%JF
-%this point is important:
+
 However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
-bring these pairs and the $\Lambda$CDM expectation outside the 1-$\sigma$ observational uncertainty.
+bring these pairs and the $\Lambda$CDM expectation outside the
+1-$\sigma$ observational uncertainty.
 
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM_narrow.pdf}
-\caption{{\rm \label{fig:lambda} Normalized histograms of Peeble's spin parameter $\lambda$ for the pairs in the Bolshoi simulation and its infered values for the LG from the observational constraints from a Monte Carlo simulation.
-The left panel includes all the LG-pairs while the right panel includes a reduced sample of pairs that match the separation and total mass infered from Local Group observations.
-The vertical lines with the white dots represent the values inferred for pairs in the constrained simulations.}}
+\caption{{\rm \label{fig:lambda} Normalized histograms of Peeble's
+    spin parameter $\lambda$ for the pairs in the Bolshoi simulation
+    and its infered values for the LG from the observational
+    constraints from a Monte Carlo simulation. The left panel includes
+    all the LG-pairs while the right panel includes a reduced sample
+    of pairs that match the separation and total mass infered from
+    Local Group observations. The vertical lines with the white dots
+    represent the values inferred for pairs in the constrained
+    simulations.}} 
 \label{fig:lambda}
@@
 The dimensionless 
-spin parameter $\lambda$ \citep{Peebles1971} is used here to characterize the dynamical state of the observed and simulated LGs.
-The parameter measures the dynamical role of the angular momentum in terms of the gravitational attraction and is defined by
+spin parameter $\lambda$ \citep{Peebles1971} is used here to
+characterize the dynamical state of the observed and simulated
+LGs. The parameter measures the dynamical role of the angular momentum
+in terms of the gravitational attraction and is defined by
 
@@
 
-We compare the distribution for  $\lambda$ obtained from the pair population in the Bolshoi simulation and the distribution from the 
-Monte Carlo
+We compare the distribution for  $\lambda$ obtained from the pair
+population in the Bolshoi simulation and the distribution from the  Monte Carlo
 simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$. 
-Figure \ref{fig:lambda} shows the likelihood distribution of the observational errors and the spin parameter of the LG-like pairs of the full (left panel) and the sub- (right panel) samples.
-The value of $\lambda$ of the constrained LGs is shown as well.
+Figure \ref{fig:lambda} shows the likelihood distribution of the
+observational errors and the spin parameter of the LG-like pairs of
+the full (left panel) and the sub- (right panel) samples.  The value
+of $\lambda$ of the constrained LGs is shown as well.
 
-The expected configuration in energy and angular momentum of the LG is slightly inconsistent with the expectation from halo pairs in $\Lambda CDM$.
+The expected configuration in energy and angular momentum of the LG is
+slightly inconsistent with the expectation from halo pairs in $\Lambda
+CDM$.
 
 The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
-Inspection of Figure \ref{fig:lambda} shows that the $\lambda$ distribution is close to a lognormal one and thus resembles the distribution of the spin parameter of DM halos in simulation ({\bf Jaime - please provide a reference}) .
-The estimated value of $\log_{10}\lambda$ of the observed LG is $-1.76$, compared with the mean value of 
- $\log_{10}\lambda$ of $-1.51$ for the full sample and $-1.38$ for the subsample. The values of the spin parameter of the 3 constrained LGs are
+Inspection of Figure \ref{fig:lambda} shows that the $\lambda$
+distribution is close to a lognormal one and thus resembles the
+distribution of the spin parameter of DM halos in simulation ({\bf
+  Jaime - please provide a reference}) .  
+The estimated value of $\log_{10}\lambda$ of the observed LG is
+$-1.76$, compared with the mean value of $\log_{10}\lambda$ of $-1.51$
+for the full sample and $-1.38$ for the subsample. The values of the
+spin parameter of the 3 constrained LGs are $\log_{10}(\lambda_{\rm
+  CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging
+to the sole constrained LG of the subsample.
 
-$\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging to the sole constrained LG of the subsample.
-
-% 
 It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
 %JF. Yehuda, I don't understaind this sentence. 
-This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
+%This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
 %YH
@@
 \section{Conclusions}
-We have presented a comparison between the observed kinematics for the M31 in the galactocentric restframe and the expectations for a large N-body cosmological simulation in the $\Lambda$CDM cosmology.
-In the simulation we select a sample of halo pairs in the mass range $5\times 10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation conditions of the Local Group from other massive structures.
-While the observations show that M31 moves towards us
-on a highly eccentric orbit,%YH
- the simulation shows that the most common configuration at $z=0$ has values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm5$ \kms.
+We have presented a comparison between the observed kinematics for the
+M31 in the galactocentric restframe and the expectations for a large
+N-body cosmological simulation in the $\Lambda$CDM cosmology. In the
+simulation we select a sample of halo pairs in the mass range $5\times
+10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation
+conditions of the Local Group from other massive structures. While the
+observations show that M31 moves towards us  on a highly eccentric orbit,
+ the simulation shows that the most common configuration at $z=0$ has
+ values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan,
+   \Lambda CDM}=62\pm5$ \kms.
 
-Using the same absolute values for the uncertainty in the observed velocity components, we find that halos within the preferred $\Lambda$CDM values are five times more common than pairs compatible with the observational constraint.
-The qualitative nature of these results is still valid after a narrower selection on separation and total pair mass.
-Additionally, pairs with a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar to observations) represent $8\%$ of the total sample of LG-like pairs.
-Making an tighter selection to match the observational constraints on the separation and total mass results in zero pairs compatible with observations.
+Using the same absolute values for the uncertainty in the observed
+velocity components, we find that halos within the preferred
+$\Lambda$CDM values are five times more common than pairs compatible
+with the observational constraint.  The qualitative nature of these
+results is still valid after a narrower selection on separation and
+total pair mass. Additionally, pairs with a fraction of tangential to
+radial velocity $f_{\rm t}<0.32$ (similar to observations) represent
+$8\%$ of the total sample of LG-like pairs. Making an tighter
+selection to match the observational constraints on the separation and
+total mass results in zero pairs compatible with observations.
 
-Approximating the LG as two point masses we express the above mentioned results in terms of the orbital angular momentum $l_{\rm orb}$ and the mechanical energy $e_{\rm tot}$ per unit of reduced mass.
-We find that the uncertainties in the tangential velocity, the square of the norm of the velocity and the total mass in the LG produce poorer constraints on the number of simulated pairs that are consistent with the observations.
-Nevertheless, in the case of the LG-pair sample that also fulfills the separation and total mass criteria there is a slight tension between simulation and observation.
-A reduction by a factor of $2$ in the observational uncertainty on the radial velocity would clarify this issue.
+Approximating the LG as two point masses we express the above
+mentioned results in terms of the orbital angular momentum $l_{\rm
+  orb}$ and the mechanical energy $e_{\rm tot}$ per unit of reduced
+mass. We find that the uncertainties in the tangential velocity, the
+square of the norm of the velocity and the total mass in the LG
+produce poorer constraints on the number of simulated pairs that are
+consistent with the observations. Nevertheless, in the case of the
+LG-pair sample that also fulfills the separation and total mass
+criteria there is a slight tension between simulation and
+observation. A reduction by a factor of $2$ in the observational
+uncertainty on the radial velocity would clarify this issue.
 
-In the three pairs from constrained simulations we find kinematics dominated by radial velocities.
-However their velocity components differ from the observational constraints and their mechanical energy and orbital angular momentum are in broad concordance with observations.
-There is only one pair that fullfills all the separation, total mass constraints and matches the most probable value for the dimensionless spin parameter $\lambda$ inferred from observations.
+In the three pairs from constrained simulations we find kinematics
+dominated by radial velocities. However their velocity components
+differ from the observational constraints and their mechanical energy
+and orbital angular momentum are in broad concordance with
+observations. There is only one pair that fullfills all the
+separation, total mass constraints and matches the most probable value
+for the dimensionless spin parameter $\lambda$ inferred from
+observations.
 
-The $\lambda$ spin parameter is itself a useful instrument to gauge the dynamical state of pairs.
-Reduced uncertainties in masses, distances and velocities can be included to produce an updated estimate for this parameter.
-This will immedialtly improve the statistical comparison between our Local Group and the expectations from $\Lambda$CDM.
+The $\lambda$ spin parameter is itself a useful instrument to gauge
+the dynamical state of pairs. Reduced uncertainties in masses,
+distances and velocities can be included to produce an updated
+estimate for this parameter. This will immedialtly improve the
+statistical comparison between our Local Group and the expectations
+from $\Lambda$CDM.
 
-In this Letter we have shown that LG-like pairs in $\Lambda$CDM show prefered values for their relative velocities, angular momentum and total mechanical energy.
-The values for the orbital angular momentum and energy, merged into the $\lambda$ spin parameter, are in mild disagreement with the observational constraints.
-However, there is a strong tension with the precise values for the radial and tangential velocities.
+In this Letter we have shown that LG-like pairs in $\Lambda$CDM show
+prefered values for their relative velocities, angular momentum and
+total mechanical energy. The values for the orbital angular momentum
+and energy, merged into the $\lambda$ spin parameter, are in mild
+disagreement with the observational constraints. However, there is a
+strong tension with the precise values for the radial and tangential
+velocities.
 
-Under the approximation of consevartion of the orbital angular momentum and mechanical energy, we see that an agreement in the total angular momentum and energy and a marked difference with the precise balance between today's radial and tangential velocities could only be explained if the initial conditions for the formation of the Local Group are special in comparison to the initial conditions of any other pair of dark matter halos in the $\Lambda$CDM cosmology.
+Under the approximation of consevartion of the orbital angular
+momentum and mechanical energy, we see that an agreement in the total
+angular momentum and energy and a marked difference with the precise
+balance between today's radial and tangential velocities could only be
+explained if the initial conditions for the formation of the Local
+Group are special in comparison to the initial conditions of any other
+pair of dark matter halos in the $\Lambda$CDM cosmology.
 
-We consider that reducing the uncertainties in the observations relevant to this study are needed.
-Likewise, additional theoretical work to understand the origin and implications of this result deserve further study.
+We consider that reducing the uncertainties in the observations
+relevant to this study are needed. Likewise, additional theoretical
+work to understand the origin and implications of this result deserve
+further study.
 
@@
 \section*{Acknowledgments}  
-J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de Investigaciones at Universidad de los Andes through its {\it Fondo de Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber Foundation through its fellowship administered by the International Astronomical Union.
-J.E.F-R also acknowledges early discussions with Alejandro Garc\'ia that motivated this work.
+J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de
+Investigaciones at Universidad de los Andes through its {\it Fondo de
+  Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber
+Foundation through its fellowship administered by the International
+Astronomical Union. J.E.F-R also acknowledges early discussions with
+Alejandro Garc\'ia that motivated this work.
 
-The data and source code and instructions to replicate the results of this paper can be found as a github repository {\texttt https://github.com/forero/LG\_Kinematics/}.
-Thanks to the IPython community \citep{IPython}.
-Thanks to Jessica Kirkpatrick for releasing her Python code to make nice plots of 2D histograms.
+The data and source code and instructions to replicate the results of
+this paper can be found as a github repository {\texttt
+  https://github.com/forero/LG\_Kinematics/}. Thanks to the IPython
+community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing
+her Python code to make nice plots of 2D histograms.
 
-The MultiDark Database used in this paper 
-%YH:
-%and the web application providing online access to it were 
-was %YH
-constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064.
-The Bolshoi simulation was run on the NASA's Pleiades supercomputer at the NASA Ames Research Center.
+The MultiDark Database used in this paper  was constructed as part of
+the activities of the German Astrophysical Virtual Observatory as
+result of a collaboration between the Leibniz-Institute for
+Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider
+Project CSD2009-00064. The Bolshoi simulation was run on the NASA's
+Pleiades supercomputer at the NASA Ames Research Center.
 

### 2013-02-08 paper/ms.tex
including new tables with a summary

@@
 \begin{abstract}
-
 Recent observations constrained the tangential velocity of M31 with
@@
 4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
-high resolution N-body cosmological simulation to statistically study
-this kinematics in the context of the $\Lambda$CDM cosmology. We
-define the Local Group (LG) to be composed by two dominant separate
-halos hosting the MW and the M31 galaxy fulfilling certain isolation
-criteria. We find that the most probable values for the tangential and
-radial velocities in these pairs are $v_{\rm rad, \Lambda CDM}=-60\pm
-15$\kms and $v_{\rm tan, \Lambda CDM}=50\pm 5$\kms. Within a similar
-absolute uncertainty defined by observations the pairs centered around
-these values are $\sim3$ times more abundant than the pairs in the
-observational interval. Furthermore, we find that only $\sim12\%$ of
-the pairs show a ratio between its tangential and radial velocity
-consistent with observations. This tension increases if we make a
-narrower selection in the pair sample to match the observational
-constraints on pair separation and total mass. Expressing these
-results in terms of the angular momentum and mechanical energy per
-unit mass also reflects this disagreement albeit in a less strong
-fashion due to the larger uncertainties inferred for these physical
-quantities. From these quantities Peeble's spin parameter is evaluated
-and is used to characterize the LG with repect to other halo pairs in
-$\Lambda$CDM. We also include in our analysis results from constrained
-simulations. The characterisation results of this study suggest that
-the formation and evolution of the LG differ from the average
-$\Lambda$CDM halo pair. 
+high resolution N-body cosmologicalsimulation (BOLSHOI) to
+statistically study this kinematics in the context of the $\Lambda$CDM
+cosmology and three constrained simulations designed to reproduce the
+Local Group (LG) and the local cosmic neighborhood. A LG is defined
+here as  composed by two dominant separate halos hosting the MW and
+the M31 galaxy fulfilling certain isolation criteria. An ensemble  of
+all BOLSHOI pairs fulfilling this selection is constructed, and also a
+subsample of it which obeys  more stringent observational constraints
+on the mass and separation of of the MW and M31 (hereafter the full-
+and sub- samples). The comparison of the ensembles of simulated pairs
+with the observed LG has been done with respect to the observed radial
+and tangential,  the reduced orbital energy ($e$) and angular momentum
+($l$) of the LG, and with respect to the dimensionless spin parameter,
+$\lambda$. Our main results are: 
+%
+%
+% Jaime here I suggest to give in a very brief way the numbers of Table 2 that I have suggested.
+%
+The three constrained LGs are more consistent with the unconstrained
+sample of LGs rather then with the observed LG.
+
+It follows that the quasi-conserved quantities quantities that
+characterize the orbit of the LG, i.e. $e$, $l$ and $\lambda$, do not
+challenge the standard $\Lambda$CDM model, but that model is in
+tension with regard to the particular phase on the orbit of the LG,
+namely the actual values of the radial and tangential velocities. This
+might hint to a problem of the $\Lambda$ CDM model to reproduce the
+observed LG.
+
 \end{abstract}
@@
 
-Note that only one constrained LG-like object is included in the subsample.
+Including these conditions the full pair sample is reduced from $1923$
+to $158$ pairs. Note that only one constrained LG-like object is
+included in the subsample.
 
 \begin{table}
-\caption{Summary of the kinematic properties for M31 in the galactocentric frame as reported by \citep{vanderMarel12}.
-Values in parenthesis correspond to vector components.
-$\sigma_{\bf x}$ represents the uncertainty on the components of vector ${\bf x}$}
+\caption{Summary of observational constraints. The kinematic
+  properties for M31 are reported in the galactocentric
+  \citep{vanderMarel12}. Values in parenthesis correspond to vector
+  components. $\sigma_{\bf x}$ represents the uncertainty on the
+  components of vector ${\bf x}$. The values for the individual halo
+masses are consistent with the priors used by \cite{vanderMarel12}.}
 \begin{center}
-\begin{tabular}{ccc}\hline\hline
+\begin{tabular}{ccc}\hline
+$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
+$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
+$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
 ${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
@@
 $\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
-$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
-$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
-$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
 $M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
 $M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
-$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\\hline
+$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\
+$\log_{10}\lambda$& & $-1.72\pm 0.07$ \\\hline
+\end{tabular}\\
+\end{center}
+\vspace{1mm}
+Note: the observational uncertainties in the position vector
+correspond to a $5\%$ in each component consistent with the
+uncertainties in the distance \citep[see references
+  in][]{vanderMarel08}.
+Note: the value for $\log_{10}\lambda$ is obtained in this Letter
+from a Monte Carlo simulation as described in Section XX.
+\label{table:1}
+\end{table}
 
-\end{tabular}\\
+\begin{table}
+\caption{Summary of results from $\Lambda$CDM.}
+\begin{center}
+\begin{tabular}{cccc}\hline
+ &  & Full Sample & Reduced Sample\\
+$v_{\rm M31,rad}$ &(\kms) & $-70\pm 10$ & $-90\pm 10$\\
+$v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
+$\log_{10}\lambda$& & $-1.47\pm 0.13$& \\ 
+\end{tabular}
+\end{center}
 \vspace{1mm}
@@
   in][]{vanderMarel08}
-\end{center}
-\label{table:1}
+
+\label{table:2}
 \end{table}
@@
 \label{sec:results}
+
+\begin{table*}
+\caption{Summary of the comparison against $\Lambda$CDM.}
+\begin{center}
+\begin{tabular}{ccccc}\hline
+Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
+consistent& (\%) Pairs consistent\\ 
+& with observations & with observations & with LCDM & with LCDM\\ 
+ & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
+$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
+$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
+$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
+$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\
+\end{tabular}
+\end{center}
+\vspace{1mm}
+Note: the observational uncertainties in the position vector
+correspond to a $5\%$ in each component consistent with the
+uncertainties in the distance \citep[see references
+  in][]{vanderMarel08}
+
+\label{table:2}
+\end{table*}
 

### 2013-02-08 code/makeTempHistogramPlot.py
irrelevant edits

@@ -70,7 +70,7 @@ def MakeTempHistogramPlot(xdata,ydata,filename=None,clues=[], clues_label='',xli
 
     # Make the 'main' temperature plot
     xbins = linspace(start = xmin, stop = xmax, num = nxbins)
-    ybins = linspace(start = ymin, stop = ymax, num = nybins)
+    ybins = linspace(start = ymin, stop = ymax, num = nybins) 
     xcenter = (xbins[0:-1]+xbins[1:])/2.0
     ycenter = (ybins[0:-1]+ybins[1:])/2.0
     aspectratio = 1.0*(xmax - xmin)/(1.0*ymax - ymin)

### 2013-02-11 code/generateInfo.py
took out the eigenvalue information and files

@@ -1,5 +1,5 @@
 from numpy import *
-def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='test.out', hubble=0.73):
+def generate_info(dmh_filename, pairs_ID_filename, output_file='test.out', hubble=0.73):
     """
     This loads the data from a FOF catalog and a list of pairs IDs. 
     From this information we calculate the distance between the pairs, 
@@ -18,7 +18,7 @@ def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='tes
     outraw = open(output_file+'.raw', 'w')
     halos_prop=loadtxt(dmh_filename) 
     iso_pairs_prop=loadtxt(pairs_ID_filename) 
-    lambda_data=loadtxt(lambda_file)
+
 
     outraw.write('# M1 M2 x1 y1 z1 x2 y2 z2 v1_x v1_y v1_z v2_x v2_y v2_z\n')
     outraw.write('# Msun Msun Mpc Mpc Mpc Mpc Mpc Mpc km/s km/s kms/ km/s km/s km/s\n')
@@ -62,14 +62,6 @@ def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='tes
         r_12 = r_2-r_1
         v_12 = v_2-v_1
         
-        #eigenvalues
-        lambda1_1 = lambda_data[id_1-1,0]
-        lambda2_1 = lambda_data[id_1-1,1]
-        lambda3_1 = lambda_data[id_1-1,2]
-
-        lambda1_2 = lambda_data[id_2-1,0]
-        lambda2_2 = lambda_data[id_2-1,1]
-        lambda3_2 = lambda_data[id_2-1,2]
 
         #relative velocity including Hubble flow
         hubble_flow = hubble*100*r_12
@@ -112,11 +104,10 @@ def generate_info(dmh_filename, pairs_ID_filename, lambda_file, output_file='tes
 
 
         if(v_radial_norm<0.0 and norm_r_12 < 1.0):
-#            print v_tan_norm, v_radial_norm, norm_v_12,lambda1_1, lambda2_1, lambda3_1, lambda1_2, lambda2_2, lambda3_2
-            out.write('%f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f %f\n' % 
+            out.write('%f %f %f %f %f %f %f %f %f %f %f \n' % 
[…]

### 2013-02-11 paper/ms.tex
updated version

@@
 subsample of it which obeys  more stringent observational constraints
-on the mass and separation of of the MW and M31 (hereafter the full-
-and sub- samples). The comparison of the ensembles of simulated pairs
-with the observed LG has been done with respect to the observed radial
-and tangential,  the reduced orbital energy ($e$) and angular momentum
-($l$) of the LG, and with respect to the dimensionless spin parameter,
-$\lambda$. Our main results are: 
-%
-%
-% Jaime here I suggest to give in a very brief way the numbers of Table 2 that I have suggested.
-%
-The three constrained LGs are more consistent with the unconstrained
-sample of LGs rather then with the observed LG.
-
-It follows that the quasi-conserved quantities quantities that
+on the mass and separation of of the MW and M31. The comparison of the
+ensembles of simulated pairs with the observed LG has been done with
+respect to the observed radial and tangential,  the reduced orbital
+energy ($e$) and angular momentum ($l$) of the LG, and with respect to
+the dimensionless spin parameter, $\lambda$. Our main results are: (i)
+the prefered radial and tangential velocities for pairs in $\Lambda$CDM
+are $v_{\rm r}=-80\pm 20$, $v_{\rm t}=50\pm 10$, (ii) pairs around
+that region are $3$ to $13$ times more common than pairs within the
+observational values, (iii) $15\%$ to $24\%$ of pairs in $\Lambda$CDM have
+energy and angular momentum consistent with observations while (iv) $9\%$
+to $13\%$ of pairs in the same sample show similar values to the inferred dimensionless spin
+parameter. It follows that the quasi-conserved quantities quantities that
 characterize the orbit of the LG, i.e. $e$, $l$ and $\lambda$, do not
@@
 namely the actual values of the radial and tangential velocities. This
-might hint to a problem of the $\Lambda$ CDM model to reproduce the
+might hint to a problem of the $\Lambda$CDM model to reproduce the
 observed LG.
@@
 $v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
-$\log_{10}\lambda$& & $-1.47\pm 0.13$& \\ 
+$\log_{10}\lambda$& & $-1.47\pm 0.13$& $-1.34\pm 0.12$\\\hline
 \end{tabular}
@@
 
-\begin{table*}
-\caption{Summary of the comparison against $\Lambda$CDM.}
-\begin{center}
-\begin{tabular}{ccccc}\hline
-Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
-consistent& (\%) Pairs consistent\\ 
-& with observations & with observations & with LCDM & with LCDM\\ 
- & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
-$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
-$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
-$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
-$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\
-\end{tabular}
-\end{center}
-\vspace{1mm}
-Note: the observational uncertainties in the position vector
-correspond to a $5\%$ in each component consistent with the
-uncertainties in the distance \citep[see references
-  in][]{vanderMarel08}
-
-\label{table:2}
-\end{table*}
-
 \subsection{Radial and tangential velocities}
@@
 Figure \ref{fig:rt} summarizes the central finding of this
-Letter. Most of the LG-like pairs (left panel) in the Bolshoi
-Simulation and the pairs in the constrained simulation have radial and
-tangential velocities notably different from the observational
-constraints.  The most probable velocities in the simulation are
-around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda
-  CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect
-the minimum grid size needed to obtain robust statistics from the 2D
-histogram. There are $23$ pairs out of the full LG-like pairs sample
-centered in those values within the same range of absolute
-uncertainties allowed by observations (i.e. $\sigma_{\rm tan}=17$\kms
-and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs
-in the interval allowed by observations.
+Letter. Most of the pairs in the two samples constructed from the
+Bolshoi Simulation have radial and tangential velocities notably
+different from the observational constraints.
 
-The left panel in the same Figure \ref{fig:rt} shows the results for a
-more restrictive selection of objects in the LG-like sample, reducing
-the number of objects from  $1923$   to $158$ and their most probable
-configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm
-  tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval
-and none consistent with observations. In this case none of the pairs
-are found in the interval allowed by observations. The high eccentric
-orbit of the observed LG  constitutes an unlikely configurations for
-the $\Lambda$CDM LG-like objects. This holds for the full and the
-subsample of objects.
+The most probable radial and tangential velocities in $\Lambda$CDM are
+summarized in Table \ref{table:2}. For the full sample we have $v_{\rm
+  rad,\Lambda CDM} = -70\pm10$\kms and $v_{\rm tan,\Lambda CDM} =
+50\pm20$ \kms. For the reduced sample $v_{\rm
+  rad,\Lambda CDM} = -90\pm10$\kms and $v_{\rm tan,\Lambda CDM} =
+50\pm20$ \kms, where the uncertainties in these values reflect
+the minimum grid size needed to obtain robust statistics for the 2D
+histogram.
+
+The number of pairs withing the range of uncertainty of observations
+are listed in Table \ref{table:3} (columns 2 and 3). In the same table
+(columns 4 and 5) we summarize the number of pairs around the
+$\Lambda$CDM values within the same range of absolute observational
+uncertainty (i.e. $\sigma_{\rm tan}=17$\kms and $\sigma_{\rm
+  rad}=4$\kms.)
+
+From these results we infer that the pairs around the preferred phase
+space for $\Lambda$CDM are at least $13$ times more common than pairs
+with the observed velocities for the Local group. We highlight that
+this is a lower bound given that in the reduced sample none of the
+pairs are found in the interval allowed by observations. The high
+eccentric orbit of the observed LG  constitutes an unlikely
+configurations for the $\Lambda$CDM LG-like objects. This holds for
+the full and the subsample of objects.
 
@@
   t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$,
-while the observations suggest $f_{\rm t}<0.32$. In the Bolshoi
-Simulation, out of $1923$ LG-like pairs we find that $242$ pairs
-($\sim 12\%$ of the sample) have values for $f_{\rm t}$ consistent
-with that constraint, for the more restrictive sample there are $13$
-pairs within that constraint ($\sim 8\%$ of the sample). The three
+while the observations suggest $f_{\rm t}<0.32$. As summarized in Table
+\ref{table:3} we find that only between $8\%$ to $12\%$ of the pairs are
+consistent with the observational constraint. The three
 pairs from the constrained realizations are also outside the region of
@@
 tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
-
-The difference between the left and right panels of Figure
-\ref{fig:rt} reflects  the tension found from the full and the sub-
-sample of the LG-like objects. {\bf Yehuda, What tension??? The
-  tension comes between theory and observation, not between the two
-  theoretical results.}
 
@@
 \caption{Histograms of the radial and tangential velocities for
-  LG-like halo pairs in the Bolshoi simulation. The left panel
-  corresponds to the general LG-like pair simple, the right panel
-  corresponds to the subsample with cuts on pair separation and total
-  mass, matching the observational constraints. The squared panel
-  shows as a shaded histogram the number of pairs on the radial and
-  tangential velocities plane. The rectangular panels show the
+  LG-like halo pairs in the Bolshoi simulation. The left (right) panel
+  correspond to the full (reduced) pair sample.
+The squared panel shows as a shaded histogram the number of pairs on
+the radial and tangential velocities plane. The rectangular panels show the
   histograms when only one of the velocity components is used to bin
   the data. The half ellipse corresponds to the observed observational
-  constraints for M31 in the Galactocentric rest frame: $v_{\rm
-    rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.4$ km s$^{-1}$
-  \citep{vanderMarel12}. The circles represent the positions of the
-  pairs from the constrained simulations. The highest number density
-  in the left panel is obtained around $v_{\rm rad,\Lambda CDM} =
-  -70\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, for the
-  right panel it is located around $v_{\rm rad,\Lambda CDM} = -50\pm5$
-  \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ } 
+  and the circles represent the positions of the
+  pairs from the constrained simulations. The locii for the highest
+  density regions are listed in Table \ref{table:2}.} 
 \label{fig:rt}
@@
 \end{figure*}
+
+\begin{table*}
+\caption{Summary of the comparison of the observational results
+  against $\Lambda$CDM.}
+\begin{center}
+\begin{tabular}{ccccc}\hline
+Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
+consistent& (\%) Pairs consistent\\ 
+property & with observations & with observations & with $\Lambda$CDM & with $\Lambda$CDM\\ 
+ & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
+$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
+$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
+$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
+$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
+\end{tabular}
+\end{center}
+\vspace{1mm}
+Note: In columns 4 and 5 the number of pairs around the prefered region
+in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
+Fig. \ref{figure:rt} and summarized in Table \ref{table:2}) are
+calculated within the same observational absolute uncertainty.
+\label{table:3}
+\end{table*}
 
@@
 
-Conversely, there is an observational disadvantage. These quantities
-are derived from very different kinds of observations, increasing the
-uncertainties in their final determination.
+However, this formulation has an observational disadvantage. The
+reduced energy and angular momentum are derived from very different
+kinds of observations, increasing the uncertainties in their final
+determination.
 
@@
 to a particular phase on the LG two-body orbit. Given the masses,
-velocity and position of the observed MW and M31 (TAble \ref{table:1}
+velocity and position of the observed MW and M31 (Table \ref{table:1}
 ) the reduced energy and angular momentum are calculated. The Monte
@@
 does not cross the zero level, as can be naively expected from Figure
-\ref{fig:rt}. This does happen for the $\approx 80\%$ level
-contour. The small difference between the two presentations of the
-observational uncertainties stems from the Monte Carol sampling.
+\ref{fig:rt}. This happens for the $\approx 80\%$ level
+contour. The small difference between the presentations of the
+observational uncertainties in Figure \ref{fig:rt} versus Figure
+\ref{fig:EJ} stems from the Monte Carlo sampling.
 
-Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
+Visual inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
 $l_{\rm orb}$ constraints are less restrictive compared with the
-radial and tangential velocity constraints. For the unconstrained full
-(sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$
-observational constraints. All the constrained LGs of the full
-samples are consistent with the 1-$sigma$ observational uncertainty.
+radial and tangential velocity constraints. As summarize in Table
+\ref{table:3}, for the unconstrained full (sub-) sample 15\% (24\%)
+of the LG-like pairs obey the 1-$\sigma$ observational
+constraints.
 
-However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
-bring these pairs and the $\Lambda$CDM expectation outside the
-1-$\sigma$ observational uncertainty.
+All the constrained LGs of the full samples are
+consistent with the 1-$\sigma$ observationaluncertainty.  However a
+reduction by a factor of $2$ in the uncertainty of the tangential
+velocity showing that the tangential velocity is below $17$\kms would
+bring these pairs and the $\Lambda$CDM expectation outside the 1-$\sigma$ observational uncertainty.
 
@@
     and its infered values for the LG from the observational
-    constraints from a Monte Carlo simulation. The left panel includes
-    all the LG-pairs while the right panel includes a reduced sample
-    of pairs that match the separation and total mass infered from
-    Local Group observations. The vertical lines with the white dots
-    represent the values inferred for pairs in the constrained
+    constraints from a Monte Carlo simulation. The left (right) panel
+    corresponds to the full (reduced) sample. The vertical lines with
+    the white dots   represent the values inferred for pairs in the
+    constrained 
     simulations.}} 
@@
 square of the norm of the velocity and the total mass in the LG
-produce poorer constraints on the number of simulated pairs that are
+are less constraining on the number of simulated pairs that are
 consistent with the observations. Nevertheless, in the case of the
@@
 Group are special in comparison to the initial conditions of any other
-pair of dark matter halos in the $\Lambda$CDM cosmology.
-
-We consider that reducing the uncertainties in the observations
-relevant to this study are needed. Likewise, additional theoretical
-work to understand the origin and implications of this result deserve
-further study.
+pair of dark matter halos in the $\Lambda$CDM cosmology. This opens a
+new window into the question of how unique, if at all, is the LG in a
+cosmological context. This will continue to be studies within the
+CLUES framework.
 
@@
 \section*{Acknowledgments}  
-J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de
+JEF-R acknowledges financial support from the Vicerrector\'{\i}a de
 Investigaciones at Universidad de los Andes through its {\it Fondo de
@@
 Foundation through its fellowship administered by the International
-Astronomical Union. J.E.F-R also acknowledges early discussions with
-Alejandro Garc\'ia that motivated this work.
+Astronomical Union. JEF-R also acknowledges early discussions with
+Alejandro Garc\'ia that motivated this work. YH has been supported by
+the ISF (1013/12)
 

### 2013-02-13 paper/ms.tex
brand new version after a final revision round

@@
 4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
-high resolution N-body cosmologicalsimulation (BOLSHOI) to
+high resolution N-body cosmological simulation (Bolshoi) to
 statistically study this kinematics in the context of the $\Lambda$CDM
-cosmology and three constrained simulations designed to reproduce the
-Local Group (LG) and the local cosmic neighborhood. A LG is defined
-here as  composed by two dominant separate halos hosting the MW and
-the M31 galaxy fulfilling certain isolation criteria. An ensemble  of
-all BOLSHOI pairs fulfilling this selection is constructed, and also a
-subsample of it which obeys  more stringent observational constraints
-on the mass and separation of of the MW and M31. The comparison of the
+cosmology together with three constrained simulations designed to reproduce the
+Local Group (LG) and the local cosmic neighborhood. The comparison of the
 ensembles of simulated pairs with the observed LG has been done with
-respect to the observed radial and tangential,  the reduced orbital
-energy ($e$) and angular momentum ($l$) of the LG, and with respect to
-the dimensionless spin parameter, $\lambda$. Our main results are: (i)
-the prefered radial and tangential velocities for pairs in $\Lambda$CDM
-are $v_{\rm r}=-80\pm 20$, $v_{\rm t}=50\pm 10$, (ii) pairs around
-that region are $3$ to $13$ times more common than pairs within the
-observational values, (iii) $15\%$ to $24\%$ of pairs in $\Lambda$CDM have
-energy and angular momentum consistent with observations while (iv) $9\%$
-to $13\%$ of pairs in the same sample show similar values to the inferred dimensionless spin
-parameter. It follows that the quasi-conserved quantities quantities that
-characterize the orbit of the LG, i.e. $e$, $l$ and $\lambda$, do not
-challenge the standard $\Lambda$CDM model, but that model is in
-tension with regard to the particular phase on the orbit of the LG,
-namely the actual values of the radial and tangential velocities. This
-might hint to a problem of the $\Lambda$CDM model to reproduce the
-observed LG.
+respect to the observed radial and tangential velocities, the reduced orbital
+energy ($e_{\rm tot}$) and angular momentum ($l_{\rm orb}$) of the LG,
+and the dimensionless spin parameter, $\lambda$. Our
+main results are: (i) the prefered radial and tangential velocities
+for pairs in $\Lambda$CDM are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm
+  t}=50\pm 10$\kms, (ii) pairs around that region are $3$ to $13$
+times more common than pairs within the observational values, (iii)
+$15\%$ to $24\%$ of pairs in $\Lambda$CDM have energy and angular
+momentum consistent with observations while (iv) $9\%$ to $13\%$ of
+pairs in the same sample show similar values in the inferred
+dimensionless spin parameter. It follows that the quasi-conserved
+quantities quantities that characterize the orbit of the LG,
+i.e. $e_{\rm tot}$,
+$l_{\rm orb}$ and $\lambda$, do not challenge the standard $\Lambda$CDM model,
+but the model is in tension with regard to the actual values of the radial and
+tangential velocities. This might hint to a problem of the
+$\Lambda$CDM model to reproduce the observed LG.
 
@@
 tidal disruption features \citep{pandas09} and the disk dominated
-morphology \citep{Kazantzidis2008} are examples on of how LG studies
+morphology \citep{Kazantzidis2008} are examples on how LG studies
 are linked to the cosmological context. Detailed studies on the
@@
 with a head on collision in terms of the ratio of the radial to
-tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.3$ and
-present these results in terms of the reduced angular momentum and
-mechanical energy.
+tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.32$ and
+present these results in terms of the reduced angular momentum,
+mechanical energy and dimensionless spin parameter.
 
@@
 
-\label{subsec:lg-sample}
-\begin{figure*}
-\begin{center}
-\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
-\caption{\label{fig:distros} Distributions of separations and total
-  halo mass (Milky Way and M31) in the LG-like pairs selected from the
-  Bolshoi Simulation.} 
-\end{center}
-\end{figure*}
+%\label{subsec:lg-sample}
+%\begin{figure*}
+%\begin{center}
+%\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation_BDM.pdf}
+%\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
+%\caption{\label{fig:distros} Distributions of separations and total
+%  halo mass (Milky Way and M31) in the LG-like pairs selected from the
+%  Bolshoi Simulation.} 
+%\end{center}
+%\end{figure*}
 
@@
 With these parameters the mass per particle is $m_{p}=1.4\times
-10^{8}$\hMsun. In this paper we use algorithms obtained through the
+10^{8}$\hMsun. In this paper we use halos obtained through the
 Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}. The halos are
@@
 \begin{enumerate}
-\item Each halo has a mass in the range $5\times10^{11}\hMsun
-  <M_{h}<5\times 10^{12}\hMsun$. 
+\item Each halo has a mass in the range $7\times10^{11}\Msun
+  <M_{h}<7\times 10^{12}\Msun$. 
 \item With respect to each halo, there cannot be any other halo within
-  the mass range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$
+  the mass range $7\times10^{11}\Msun <M_{h}<7\times 10^{12}\Msun$
   closer than its partner. It means that there cannot be ambiguity on
@@
 \item The distance between the center of mass of the halos must be
-  less than $0.7$\hMpc \citep{ribas05,vanderMarel08}. 
-\item There cannot be halos more massive than $5\times 10^{12}$\hMsun
-  within a radius of $2$\hMpc with respect to every object centre
+  less than $1.0$Mpc \citep{ribas05,vanderMarel08}. 
+\item There cannot be halos more massive than $7\times 10^{12}$\Msun
+  within a radius of $3$Mpc with respect to every object centre
   \citep{Karachentsev04,Anton09}. 
-\item There cannot be halos more massive than $5\times 10^{13}$\hMsun
-  within  a radius of $3$\hMpc with respect to every object centre
+\item There cannot be halos more massive than $7\times 10^{13}$\Msun
+  within  a radius of $4$Mpc with respect to every object centre
   \citep{Karachentsev04}. 
@@
 conditions and additionally are located in a place with the right
-distances with respect to the Virgo cluster in the simulation. Figure
-\ref{fig:distros} show the  distribution of separations and total pair
-mass computed from the $1923$ pairs in the Bolshoi simulation.
+distances with respect to the Virgo cluster in the simulation. 
+%Figure
+%\ref{fig:distros} show the  distribution of separations and total pair
+%mass computed from the $1923$ pairs in the Bolshoi
+%simulation. 
+Throughout this Letter werefer to this sample as the full
+sample.
 
@@
 the MW-M31 pair are listed in Table \ref{table:1}.  
-A more restrictive subsample of LG-like objects has been constructed
+A more reduced sample from the full sample has been constructed
 so as to fit obey the observational bounds on the masses and
@@
 \begin{enumerate}
-\item The separation between the center of mass of the halos is in the range $700-800$kpc \citep{ribas05,vanderMarel08}.
-\item The total mass of the two halos is in the range $1-4\times 10^{12}$\Msun \citep{vanderMarel12}.
+\item The separation between the center of mass of the halos is in the
+  range $700-800$kpc \citep{ribas05,vanderMarel08}. 
+\item The total mass of the two halos is in the range $1-4\times
+  10^{12}$\Msun \citep{vanderMarel12}. 
 \end{enumerate}
 
-Including these conditions the full pair sample is reduced from $1923$
-to $158$ pairs. Note that only one constrained LG-like object is
-included in the subsample.
+Including these conditions the full sample is reduced from $1923$
+to $158$ pairs. We refer to this sample as the reduced sample. Note
+that only one constrained LG-like object is included in this subset.
 
 \begin{table}
-\caption{Summary of observational constraints. The kinematic
-  properties for M31 are reported in the galactocentric
-  \citep{vanderMarel12}. Values in parenthesis correspond to vector
-  components. $\sigma_{\bf x}$ represents the uncertainty on the
-  components of vector ${\bf x}$. The values for the individual halo
-masses are consistent with the priors used by \cite{vanderMarel12}.}
+\caption{Summary of observational constraints.}
 \begin{center}
@@
 \vspace{1mm}
-Note: the observational uncertainties in the position vector
+Notes:\\
+\begin{enumerate}
+\item The kinematic
+  properties for M31 are reported in the galactocentric
+  \citep{vanderMarel12}.
+
+\item Values in parenthesis correspond to vector
+  components. $\sigma_{\bf x}$ represents the uncertainty on the
+  components of vector ${\bf x}$. 
+\item The values for the individual halo
+masses are consistent with the priors used by \cite{vanderMarel12}.
+
+\item The observational uncertainties in the position vector
 correspond to a $5\%$ in each component consistent with the
@@
   in][]{vanderMarel08}.
-Note: the value for $\log_{10}\lambda$ is obtained in this Letter
-from a Monte Carlo simulation as described in Section XX.
+
+\item The value for $\log_{10}\lambda$ is obtained in this Letter
+from a Monte Carlo simulation as described in Section
+\ref{subsection:e-l}. 
+\end{enumerate}
 \label{table:1}
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
+\caption{Upper panels: Histograms of the radial and tangential
+  velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
+  panels: Histograms of the orbital angular momentum ($l_{\rm orb}$)
+  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
+  calculated considering the halos as point masses. The left (right) panels
+  correspond to the full (reduced) pair sample. The locii for the highest
+  density regions are in the $v_{\rm r}-v_{\rm t}$ plane are listed in
+  Table \ref{table:2}. 
+  The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
+  the observed observational and the circles represent the positions
+  of the pairs from the constrained simulations. The contour line in
+  the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the Monte
+  Carlo generated points to estimated the uncertainties from the
+  observational values summarized in Table  1.}
+\label{fig:rt}
+\end{center}
+
+\end{figure*}
+
+\begin{table*}
+\caption{Summary of the comparison of the observational results
+  against $\Lambda$CDM.}
+\begin{center}
+\begin{tabular}{ccccc}\hline
+Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
+with highest & (\%) with highest\\ 
+property & with observations & with observations & likelihood in
+$\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
+ & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
+$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & ($<$0.6\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
+$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
+$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
+$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
+\end{tabular}
+\end{center}
+\vspace{1mm}
+Note: In columns 4 and 5 the number of pairs around the prefered region
+in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
+Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
+calculated within the same observational absolute uncertainty.
+\label{table:3}
+\end{table*}
+
 \subsection{Radial and tangential velocities}
@@
 
-The number of pairs withing the range of uncertainty of observations
-are listed in Table \ref{table:3} (columns 2 and 3). In the same table
-(columns 4 and 5) we summarize the number of pairs around the
-$\Lambda$CDM values within the same range of absolute observational
-uncertainty (i.e. $\sigma_{\rm tan}=17$\kms and $\sigma_{\rm
-  rad}=4$\kms.)
+The number of pairs compatible with LG observations are listed in
+Table \ref{table:3} (columns 2 and 3). In the same table (columns 4
+and 5) we summarize the number of pairs around the $\Lambda$CDM values
+within the same range of absolute observational uncertainty
+(i.e. $\sigma_{\rm tan}=17$\kms and $\sigma_{\rm rad}=4$\kms.)
 
@@
 configurations for the $\Lambda$CDM LG-like objects. This holds for
-the full and the subsample of objects.
+the full and the reduced sample of objects.
 
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
-\caption{Histograms of the radial and tangential velocities for
-  LG-like halo pairs in the Bolshoi simulation. The left (right) panel
-  correspond to the full (reduced) pair sample.
-The squared panel shows as a shaded histogram the number of pairs on
-the radial and tangential velocities plane. The rectangular panels show the
-  histograms when only one of the velocity components is used to bin
-  the data. The half ellipse corresponds to the observed observational
-  and the circles represent the positions of the
-  pairs from the constrained simulations. The locii for the highest
-  density regions are listed in Table \ref{table:2}.} 
-\label{fig:rt}
-\end{center}
+\subsection{Reduced Angular Momentum and Energy}
+\label{subsection:e-l}
 
-\end{figure*}
-
-\begin{table*}
-\caption{Summary of the comparison of the observational results
-  against $\Lambda$CDM.}
-\begin{center}
-\begin{tabular}{ccccc}\hline
-Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
-consistent& (\%) Pairs consistent\\ 
-property & with observations & with observations & with $\Lambda$CDM & with $\Lambda$CDM\\ 
- & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
-$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
-$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
-$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
-$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\\hline
-\end{tabular}
-\end{center}
-\vspace{1mm}
-Note: In columns 4 and 5 the number of pairs around the prefered region
-in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
-Fig. \ref{figure:rt} and summarized in Table \ref{table:2}) are
-calculated within the same observational absolute uncertainty.
-\label{table:3}
-\end{table*}
-
-\subsection{Reduced Angular Momentum and Energy}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
-\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$)
-  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
-  calculated considering the halos as point masses. The panel
-  distribution is the same as in Figure \ref{fig:rt}. The contour line
-  encloses $68\%$ of the Monte Carlo generated points to estimated the
-  uncertainties from the observational values summarized in Table
-  1. The constraints in this plot are less restrictive than in
-  Fig.\ref{fig:rt} due to added uncertainties on the tangential
-  velocity ($100\%$), the square of the norm of the velocity ($40\%$)
-  and the total mass of the two halos ($20\%$). } 
-\label{fig:EJ}
-\end{center}
-\end{figure*}
+%\begin{figure*}
+%\begin{center}
+%\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
+%\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
+%\caption{
+%Histograms of the orbital angular momentum ($l_{\rm orb}$)
+%  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
+%  calculated considering the halos as point masses. The panel
+%  distribution is the same as in Figure \ref{fig:rt}. The contour line
+%  encloses $68\%$ of the Monte Carlo generated points to estimated the
+%  uncertainties from the observational values summarized in Table
+%  1. The constraints in this plot are less restrictive than in
+%  Fig.\ref{fig:rt} due to added uncertainties on the tangential
+%  velocity ($100\%$), the square of the norm of the velocity ($40\%$)
+%  and the total mass of the two halos ($20\%$). } 
+%\label{fig:EJ}
+%\end{center}
+%\end{figure*}
 
@@
 dynamics of the LG is governed by the (center of mass)  reduced
-angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
-reduced energy  $e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
+angular momentum $l_{\rm orb}= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
+reduced energy  $e_{\rm tot}=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
   r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$,
@@
 dynamical quantities. This means that, after some formation time,
-these quantities do not significantly evolve as it is the case for the
-radial and tangential velocities.
+these quantities do not significantly evolve.
 
@@
 
-A possible conflict between the likelihood of the LG in the
-$\Lambda$CDM model as viewed from the radial and tangential velocities
-as opposed to the angular momentum and energy perspectives would point
-to a particular phase on the LG two-body orbit. Given the masses,
-velocity and position of the observed MW and M31 (Table \ref{table:1}
-) the reduced energy and angular momentum are calculated. The Monte
-Carlo sampling is used to to estimate the uncertainty in these
-quantities.
-
-Figure \ref{fig:EJ} presents the reduced  energy and angular momentum
-distribution of the constrained and unconstrained LG-like  objects of
-the full and subsample, following the   plotting conventions from
-Figure \ref{fig:rt}.  The  1-$\sigma$ contour of Figure \ref{fig:EJ}
+We use Monte Carlo sampling to estimate the reduced energy and angular
+momentum from the observed properties in the LG listed in
+Table\ref{table:1}. Figure \ref{fig:rt} also presents the results in the
+plane $e_{\rm tot}-l_{\rm orb}$ following the   plotting conventions from
+Figure \ref{fig:rt}. \footnote{The  1-$\sigma$ contour in the lower
+  panels of Figure \ref{fig:rt}
 does not cross the zero level, as can be naively expected from Figure
 \ref{fig:rt}. This happens for the $\approx 80\%$ level
-contour. The small difference between the presentations of the
-observational uncertainties in Figure \ref{fig:rt} versus Figure
-\ref{fig:EJ} stems from the Monte Carlo sampling.
+contour.}  As expected, the $e_{\rm tot}$  and $l_{\rm orb}$
+constraints are less restrictive compared with the radial and
+tangential velocity constraints; 15\% (24\%) of pairs in the full
+(reduced) sample obey the 1-$\sigma$ observational constraints.
 
-Visual inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
-$l_{\rm orb}$ constraints are less restrictive compared with the
-radial and tangential velocity constraints. As summarize in Table
-\ref{table:3}, for the unconstrained full (sub-) sample 15\% (24\%)
-of the LG-like pairs obey the 1-$\sigma$ observational
-constraints.
-
-All the constrained LGs of the full samples are
+All the constrained LGs of the full and reduced sample are
 consistent with the 1-$\sigma$ observationaluncertainty.  However a
@@
 velocity showing that the tangential velocity is below $17$\kms would
-bring these pairs and the $\Lambda$CDM expectation outside the 1-$\sigma$ observational uncertainty.
+bring these pairs and the $\Lambda$CDM expectation outside the
+1-$\sigma$ observational uncertainty.
 
@@
 observational errors and the spin parameter of the LG-like pairs of
-the full (left panel) and the sub- (right panel) samples.  The value
+the full (left panel) and the reduced (right panel) samples.  The value
 of $\lambda$ of the constrained LGs is shown as well.
@@
 The expected configuration in energy and angular momentum of the LG is
-slightly inconsistent with the expectation from halo pairs in $\Lambda
-CDM$.
-
+slightly inconsistent with the expectation from halo pairs in $\Lambda$CDM.
 The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value.
@@
 distribution is close to a lognormal one and thus resembles the
-distribution of the spin parameter of DM halos in simulation ({\bf
-  Jaime - please provide a reference}) .  
-The estimated value of $\log_{10}\lambda$ of the observed LG is
-$-1.76$, compared with the mean value of $\log_{10}\lambda$ of $-1.51$
-for the full sample and $-1.38$ for the subsample. The values of the
-spin parameter of the 3 constrained LGs are $\log_{10}(\lambda_{\rm
-  CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging
-to the sole constrained LG of the subsample.
+distribution of the spin parameter of DM halos in simulation
+\citep{Bett2007}.
 
-It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
-%JF. Yehuda, I don't understaind this sentence. 
-%This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
-%YH
+The estimated value of $\log_{10}\lambda$ of the
+observed LG is $-1.72\pm 0.07$, lower than the mean value of
+$\log_{10}\lambda$ of $-1.47\pm0.13$ ($-1.34\pm 0.12$) for the full
+(reduced) sample. The values of the spin parameter of the 3
+constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
+and $-1.65$, with the latter value belongingto the sole constrained LG
+of the subsample.
 
@@
 N-body cosmological simulation in the $\Lambda$CDM cosmology. In the
-simulation we select a sample of halo pairs in the mass range $5\times
-10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation
-conditions of the Local Group from other massive structures. While the
-observations show that M31 moves towards us  on a highly eccentric orbit,
+simulation we select a large sample of pairs that fulfill isolation
+criteria for the LG. We select a subsabmple that which obeys more
+stringent observational constraints on the mass and separation of the
+MW and M31.
+
+While the observations show that M31 moves towards us  on a highly eccentric orbit,
  the simulation shows that the most common configuration at $z=0$ has
- values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan,
-   \Lambda CDM}=62\pm5$ \kms.
+ values $v_{\rm rad, \Lambda CDM}=-80\pm20$\kms and $v_{\rm tan,
+   \Lambda CDM}=50\pm10$ \kms, where the errorbars are estimated to
+ include the results from the two pair samples.
 
 Using the same absolute values for the uncertainty in the observed
-velocity components, we find that halos within the preferred
-$\Lambda$CDM values are five times more common than pairs compatible
-with the observational constraint.  The qualitative nature of these
-results is still valid after a narrower selection on separation and
-total pair mass. Additionally, pairs with a fraction of tangential to
-radial velocity $f_{\rm t}<0.32$ (similar to observations) represent
-$8\%$ of the total sample of LG-like pairs. Making an tighter
-selection to match the observational constraints on the separation and
-total mass results in zero pairs compatible with observations.
+velocity components, we find that none of the halos in the reduced sample
+are compatible with observations. This makes pairs with the preferred
+$\Lambda$CDM values at least $13$ times more common than pairs
+compatible with the observational constraint. Additionally, pairs with
+a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar
+to observations) represent $8\%$ of the same sample.
 
@@
 consistent with the observations. Nevertheless, in the case of the
-LG-pair sample that also fulfills the separation and total mass
-criteria there is a slight tension between simulation and
+reduced sample there is a slight tension between simulation and
 observation. A reduction by a factor of $2$ in the observational
 uncertainty on the radial velocity would clarify this issue.
+
+We also use the $\lambda$ spin parameter to gauge
+the dynamical state of pairs. The values for the orbital angular momentum
+and energy, merged into the $\lambda$ spin parameter, are in mild
+disagreement with the observational constraints.
 
@@
 
-The $\lambda$ spin parameter is itself a useful instrument to gauge
-the dynamical state of pairs. Reduced uncertainties in masses,
-distances and velocities can be included to produce an updated
-estimate for this parameter. This will immedialtly improve the
-statistical comparison between our Local Group and the expectations
-from $\Lambda$CDM.
-
-In this Letter we have shown that LG-like pairs in $\Lambda$CDM show
-prefered values for their relative velocities, angular momentum and
-total mechanical energy. The values for the orbital angular momentum
-and energy, merged into the $\lambda$ spin parameter, are in mild
-disagreement with the observational constraints. However, there is a
-strong tension with the precise values for the radial and tangential
-velocities.
-
-Under the approximation of consevartion of the orbital angular
-momentum and mechanical energy, we see that an agreement in the total
-angular momentum and energy and a marked difference with the precise
-balance between today's radial and tangential velocities could only be
-explained if the initial conditions for the formation of the Local
-Group are special in comparison to the initial conditions of any other
-pair of dark matter halos in the $\Lambda$CDM cosmology. This opens a
-new window into the question of how unique, if at all, is the LG in a
-cosmological context. This will continue to be studies within the
-CLUES framework.
+Summarizing, we see we see a broad agreement in the total angular
+momentum and energy and a marked difference with the precise balance
+between today's radial and tangential velocities  Under the
+approximation of consevartion of the orbital angular momentum and
+mechanical energy, this could only be explained if the initial
+conditions for the formation of the Local Group are special in
+comparison to the initial conditions of any other pair of dark matter
+halos in the $\Lambda$CDM cosmology. This opens a new window into the
+question of how unique, if at all, is the LG in a cosmological
+context. This will continue to be studies within the CLUES framework.
 
@@
 \bibliographystyle{apj}
-\bibliography{references}
+%\bibliography{references} 
+\begin{thebibliography}{28}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Besla} {et~al.}(2007){Besla}, {Kallivayalil}, {Hernquist},
+  {Robertson}, {Cox}, {van der Marel}, \& {Alcock}}]{Besla2007}
+{Besla}, G., {Kallivayalil}, N., {Hernquist}, L., {Robertson}, B., {Cox},
+  T.~J., {van der Marel}, R.~P., \& {Alcock}, C. 2007, \apj, 668, 949
+
+\bibitem[{{Bett} {et~al.}(2007){Bett}, {Eke}, {Frenk}, {Jenkins}, {Helly}, \&
+  {Navarro}}]{Bett2007}
+{Bett}, P., {Eke}, V., {Frenk}, C.~S., {Jenkins}, A., {Helly}, J., \&
+  {Navarro}, J. 2007, \mnras, 376, 215
+
+\bibitem[{{Busha} {et~al.}(2011){Busha}, {Marshall}, {Wechsler}, {Klypin}, \&
+  {Primack}}]{Busha2011}
+{Busha}, M.~T., {Marshall}, P.~J., {Wechsler}, R.~H., {Klypin}, A., \&
+  {Primack}, J. 2011, \apj, 743, 40
+
+\bibitem[{{Forero-Romero} {et~al.}(2011){Forero-Romero}, {Hoffman}, {Yepes},
+  {Gottl{\"o}ber}, {Piontek}, {Klypin}, \& {Steinmetz}}]{ForeroRomero2011}
+{Forero-Romero}, J.~E., {Hoffman}, Y., {Yepes}, G., {Gottl{\"o}ber}, S.,
+  {Piontek}, R., {Klypin}, A., \& {Steinmetz}, M. 2011, \mnras, 417, 1434
+
+\bibitem[{{Fouquet} {et~al.}(2012){Fouquet}, {Hammer}, {Yang}, {Puech}, \&
+  {Flores}}]{Fouquet2012}
+{Fouquet}, S., {Hammer}, F., {Yang}, Y., {Puech}, M., \& {Flores}, H. 2012,
+  \mnras, 427, 1769
+
+\bibitem[{{Freeman} \& {Bland-Hawthorn}(2002)}]{Freeman2002}
+{Freeman}, K., \& {Bland-Hawthorn}, J. 2002, \araa, 40, 487
+
+\bibitem[{{Gottloeber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
+  {Yepes}}]{clues2010}
+{Gottloeber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, ArXiv e-prints
+
+\bibitem[{{Jarosik} {et~al.}(2011){Jarosik}, {Bennett}, {Dunkley}, {Gold},
+  {Greason}, {Halpern}, {Hill}, {Hinshaw}, {Kogut}, {Komatsu}, {Larson},
+  {Limon}, {Meyer}, {Nolta}, {Odegard}, {Page}, {Smith}, {Spergel}, {Tucker},
+  {Weiland}, {Wollack}, \& {Wright}}]{Jarosik2011}
+{Jarosik}, N., {Bennett}, C.~L., {Dunkley}, J., {Gold}, B., {Greason}, M.~R.,
+  {Halpern}, M., {Hill}, R.~S., {Hinshaw}, G., {Kogut}, A., {Komatsu}, E.,
+  {Larson}, D., {Limon}, M., {Meyer}, S.~S., {Nolta}, M.~R., {Odegard}, N.,
+  {Page}, L., {Smith}, K.~M., {Spergel}, D.~N., {Tucker}, G.~S., {Weiland},
+  J.~L., {Wollack}, E., \& {Wright}, E.~L. 2011, \apjs, 192, 14
+
+\bibitem[{{Karachentsev} {et~al.}(2004){Karachentsev}, {Karachentseva},
+  {Huchtmeier}, \& {Makarov}}]{Karachentsev04}
+{Karachentsev}, I.~D., {Karachentseva}, V.~E., {Huchtmeier}, W.~K., \&
+  {Makarov}, D.~I. 2004, \aj, 127, 2031
+
+\bibitem[{{Kazantzidis} {et~al.}(2008){Kazantzidis}, {Bullock}, {Zentner},
+  {Kravtsov}, \& {Moustakas}}]{Kazantzidis2008}
+{Kazantzidis}, S., {Bullock}, J.~S., {Zentner}, A.~R., {Kravtsov}, A.~V., \&
+  {Moustakas}, L.~A. 2008, \apj, 688, 254
+
+\bibitem[{{Klypin} {et~al.}(1999{\natexlab{a}}){Klypin}, {Gottl{\"o}ber},
+  {Kravtsov}, \& {Khokhlov}}]{KlypinBDM}
+{Klypin}, A., {Gottl{\"o}ber}, S., {Kravtsov}, A.~V., \& {Khokhlov}, A.~M.
+  1999{\natexlab{a}}, \apj, 516, 530
+
+\bibitem[{{Klypin} {et~al.}(1999{\natexlab{b}}){Klypin}, {Kravtsov},
+  {Valenzuela}, \& {Prada}}]{Klypin99}
+{Klypin}, A., {Kravtsov}, A.~V., {Valenzuela}, O., \& {Prada}, F.
+  1999{\natexlab{b}}, \apj, 522, 82
+
+\bibitem[{{Klypin} {et~al.}(2011){Klypin}, {Trujillo-Gomez}, \&
+  {Primack}}]{Bolshoi}
+{Klypin}, A.~A., {Trujillo-Gomez}, S., \& {Primack}, J. 2011, \apj, 740, 102
+
+\bibitem[{{Knebe} {et~al.}(2011){Knebe}, {Libeskind}, {Doumler}, {Yepes},
+  {Gottl{\"o}ber}, \& {Hoffman}}]{Knebe2011}
+{Knebe}, A., {Libeskind}, N.~I., {Doumler}, T., {Yepes}, G., {Gottl{\"o}ber},
+  S., \& {Hoffman}, Y. 2011, \mnras, 417, L56
+
+\bibitem[{{Liu} {et~al.}(2011){Liu}, {Gerke}, {Wechsler}, {Behroozi}, \&
+  {Busha}}]{Liu2011}
+{Liu}, L., {Gerke}, B.~F., {Wechsler}, R.~H., {Behroozi}, P.~S., \& {Busha},
+  M.~T. 2011, \apj, 733, 62
+
+\bibitem[{{McConnachie} {et~al.}(2009){McConnachie}, {Irwin}, {Ibata},
+  {Dubinski}, {Widrow}, {Martin}, {C{\^o}t{\'e}}, {Dotter}, {Navarro},
+  {Ferguson}, {Puzia}, {Lewis}, {Babul}, {Barmby}, {Bienaym{\'e}}, {Chapman},
+  {Cockcroft}, {Collins}, {Fardal}, {Harris}, {Huxor}, {Mackey},
+  {Pe{\~n}arrubia}, {Rich}, {Richer}, {Siebert}, {Tanvir}, {Valls-Gabaud}, \&
+  {Venn}}]{pandas09}
+{McConnachie}, A.~W., {Irwin}, M.~J., {Ibata}, R.~A., {Dubinski}, J., {Widrow},
+  L.~M., {Martin}, N.~F., {C{\^o}t{\'e}}, P., {Dotter}, A.~L., {Navarro},
+  J.~F., {Ferguson}, A.~M.~N., {Puzia}, T.~H., {Lewis}, G.~F., {Babul}, A.,
+  {Barmby}, P., {Bienaym{\'e}}, O., {Chapman}, S.~C., {Cockcroft}, R.,
+  {Collins}, M.~L.~M., {Fardal}, M.~A., {Harris}, W.~E., {Huxor}, A., {Mackey},
+  A.~D., {Pe{\~n}arrubia}, J., {Rich}, R.~M., {Richer}, H.~B., {Siebert}, A.,
+  {Tanvir}, N., {Valls-Gabaud}, D., \& {Venn}, K.~A. 2009, \nat, 461, 66
+
+\bibitem[{{Moore} {et~al.}(1999){Moore}, {Ghigna}, {Governato}, {Lake},
+  {Quinn}, {Stadel}, \& {Tozzi}}]{Moore99}
+{Moore}, B., {Ghigna}, S., {Governato}, F., {Lake}, G., {Quinn}, T., {Stadel},
+  J., \& {Tozzi}, P. 1999, \apjl, 524, L19
+
+\bibitem[{{Peebles}(1971)}]{Peebles1971}
+{Peebles}, P.~J.~E. 1971, \aap, 11, 377
+
+\bibitem[{{Peebles} \& {Nusser}(2010)}]{Peebles2010}
+{Peebles}, P.~J.~E., \& {Nusser}, A. 2010, \nat, 465, 565
+
+\bibitem[{P\'erez \& Granger(2007)}]{IPython}
+P\'erez, F., \& Granger, B.~E. 2007, {C}omput. {S}ci. {E}ng., 9, 21
+
+\bibitem[{{Purcell} \& {Zentner}(2012)}]{Purcell2012}
+{Purcell}, C.~W., \& {Zentner}, A.~R. 2012, {JCAP}, 12, 7
+
+\bibitem[{{Ribas} {et~al.}(2005){Ribas}, {Jordi}, {Vilardell}, {Fitzpatrick},
+  {Hilditch}, \& {Guinan}}]{ribas05}
+{Ribas}, I., {Jordi}, C., {Vilardell}, F., {Fitzpatrick}, E.~L., {Hilditch},
+  R.~W., \& {Guinan}, E.~F. 2005, \apjl, 635, L37
+
+\bibitem[{{Riebe} {et~al.}(2011){Riebe}, {Partl}, {Enke}, {Forero-Romero},
+  {Gottloeber}, {Klypin}, {Lemson}, {Prada}, {Primack}, {Steinmetz}, \&
+  {Turchaninov}}]{2011arXiv1109.0003R}
+{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottloeber}, S.,
+  {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz}, M.,
+  \& {Turchaninov}, V. 2011, ArXiv e-prints
+
+\bibitem[{{Teyssier} {et~al.}(2012){Teyssier}, {Johnston}, \&
+  {Kuhlen}}]{Teyssier2012}
+{Teyssier}, M., {Johnston}, K.~V., \& {Kuhlen}, M. 2012, \mnras, 426, 1808
+
+\bibitem[{{Tikhonov} \& {Klypin}(2009)}]{Anton09}
+{Tikhonov}, A.~V., \& {Klypin}, A. 2009, \mnras, 395, 1915
+
+\bibitem[{{Tollerud} {et~al.}(2011){Tollerud}, {Boylan-Kolchin}, {Barton},
+  {Bullock}, \& {Trinh}}]{Tollerud2011}
+{Tollerud}, E.~J., {Boylan-Kolchin}, M., {Barton}, E.~J., {Bullock}, J.~S., \&
+  {Trinh}, C.~Q. 2011, \apj, 738, 102
+
+\bibitem[{{van der Marel} {et~al.}(2012){van der Marel}, {Fardal}, {Besla},
+  {Beaton}, {Sohn}, {Anderson}, {Brown}, \& {Guhathakurta}}]{vanderMarel12}
+{van der Marel}, R.~P., {Fardal}, M., {Besla}, G., {Beaton}, R.~L., {Sohn},
+  S.~T., {Anderson}, J., {Brown}, T., \& {Guhathakurta}, P. 2012, \apj, 753, 8
+
+\bibitem[{{van der Marel} \& {Guhathakurta}(2008)}]{vanderMarel08}
+{van der Marel}, R.~P., \& {Guhathakurta}, P. 2008, \apj, 678, 187
+
+\end{thebibliography}
 

### 2013-02-15 paper/ms.tex
includes SG modifications

@@
 and the dimensionless spin parameter, $\lambda$. Our
-main results are: (i) the prefered radial and tangential velocities
+main results are: (i) the preferred radial and tangential velocities
 for pairs in $\Lambda$CDM are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm
@@
 
-This Ltter presents such study. We use a large volume, high resolution
+This Letter presents such study. We use a large volume, high resolution
 dark matter only N-body simulation in the concordance $\Lambda$CDM
@@
 and the orbital angular momentum/mechanical energy. In the same
-section we sumarize these dynamical results in terms of the
+section we summarize these dynamical results in terms of the
 dimensionless spin parameter of the pairs. Finally, in the last
@@
 matter density field using N-body techniques. The simulation has a
-cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$
+cubic comoving volume of $(250\hMpc )^3$, sampled with $2048^{3}$
 particles. The cosmological parameters used in the simulation are
@@
 (WMAP) \citep{Jarosik2011}. A detailed description of this simulation
-can be found in \citep{Bolshoi}.
+can be found in \cite{Bolshoi}.
 
@@
 conclusions of this paper hold in the case of halos defined by a FOF
-algorithm with a linking lentgth 0.17 times the interparticle
+algorithm with a linking length 0.17 times the interparticle
 distance.
@@
 The constrained simulations we use in this Letter are part of the
-Constrained Local UniversE Simulations (CLUES) project whose main
+Constrained Local UniversE Simulations (CLUES)  \footnote{{\tt http://www.clues-project.org}} project whose main
 objective is to reproduce the large scale structure in the Local
@@
 constraints to construct the initial conditions are  described in
-\citep{clues2010}.  We use three dark matter only simulations, each
+\cite{clues2010}.  We use three dark matter only simulations, each
 has a cubic volume of $64$\hMpc on a side, with the density field
@@
 %simulation. 
-Throughout this Letter werefer to this sample as the full
+Throughout this Letter we refer to this sample as the full
 sample.
@@
 \item The kinematic
-  properties for M31 are reported in the galactocentric
+  properties for M31 are reported in the galactocentric restframe 
   \citep{vanderMarel12}.
@@
 \vspace{1mm}
-Note: In columns 4 and 5 the number of pairs around the prefered region
+Note: In columns 4 and 5 the number of pairs around the preferred region
 in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
@@
 this is a lower bound given that in the reduced sample none of the
-pairs are found in the interval allowed by observations. The high
+pairs is found in the interval allowed by observations. The high
 eccentric orbit of the observed LG  constitutes an unlikely
@@
 All the constrained LGs of the full and reduced sample are
-consistent with the 1-$\sigma$ observationaluncertainty.  However a
+consistent with the 1-$\sigma$ observational uncertainty.  However a
 reduction by a factor of $2$ in the uncertainty of the tangential
@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM_narrow.pdf}
-\caption{{\rm \label{fig:lambda} Normalized histograms of Peeble's
+\caption{{\rm \label{fig:lambda} Normalized histograms of Peebles'
     spin parameter $\lambda$ for the pairs in the Bolshoi simulation
-    and its infered values for the LG from the observational
+    and its inferred values for the LG from the observational
     constraints from a Monte Carlo simulation. The left (right) panel
@@
 constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
-and $-1.65$, with the latter value belongingto the sole constrained LG
+and $-1.65$, with the latter value belonging belonging toto the sole constrained LG
 of the subsample.
@@
 simulation we select a large sample of pairs that fulfill isolation
-criteria for the LG. We select a subsabmple that which obeys more
+criteria for the LG. We select a subsample which obeys more
 stringent observational constraints on the mass and separation of the
@@
 velocity components, we find that none of the halos in the reduced sample
-are compatible with observations. This makes pairs with the preferred
+is compatible with observations. This makes pairs with the preferred
 $\Lambda$CDM values at least $13$ times more common than pairs
@@
 and orbital angular momentum are in broad concordance with
-observations. There is only one pair that fullfills all the
+observations. There is only one pair that fulfills all the
 separation, total mass constraints and matches the most probable value
@@
 between today's radial and tangential velocities  Under the
-approximation of consevartion of the orbital angular momentum and
+approximation of conservation of the orbital angular momentum and
 mechanical energy, this could only be explained if the initial
@@
 question of how unique, if at all, is the LG in a cosmological
-context. This will continue to be studies within the CLUES framework.
+context. This will continue to be studied within the CLUES framework.
 
@@
 Project CSD2009-00064. The Bolshoi simulation was run on the NASA's
-Pleiades supercomputer at the NASA Ames Research Center.
+Pleiades supercomputer at the NASA Ames Research Center. The CLUES simulation
+was run at LRZ Munich.
 
@@
 
-\bibitem[{{Gottloeber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
+\bibitem[{{Gottl{\"o}ber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
   {Yepes}}]{clues2010}
-{Gottloeber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, ArXiv e-prints
+{Gottl{\"o}ber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, 
+Proceedings of "High Performance Computing in Science and Engineering'',  
+Springer, Eds.: Wagner, S.; Steinmetz, M.; Bode, A.; M\"uller, M.M.
 
@@
   {Turchaninov}}]{2011arXiv1109.0003R}
-{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottloeber}, S.,
+{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottl{\"o}ber}, S.,
   {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz}, M.,

### 2013-02-15 paper/ms.tex
new ackwnoldgements
version submitted to apjl

@@
 4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
-high resolution N-body cosmological simulation (Bolshoi) to
-statistically study this kinematics in the context of the $\Lambda$CDM
-cosmology together with three constrained simulations designed to reproduce the
-Local Group (LG) and the local cosmic neighborhood. The comparison of the
-ensembles of simulated pairs with the observed LG has been done with
-respect to the observed radial and tangential velocities, the reduced orbital
-energy ($e_{\rm tot}$) and angular momentum ($l_{\rm orb}$) of the LG,
-and the dimensionless spin parameter, $\lambda$. Our
-main results are: (i) the preferred radial and tangential velocities
-for pairs in $\Lambda$CDM are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm
-  t}=50\pm 10$\kms, (ii) pairs around that region are $3$ to $13$
-times more common than pairs within the observational values, (iii)
-$15\%$ to $24\%$ of pairs in $\Lambda$CDM have energy and angular
-momentum consistent with observations while (iv) $9\%$ to $13\%$ of
-pairs in the same sample show similar values in the inferred
-dimensionless spin parameter. It follows that the quasi-conserved
-quantities quantities that characterize the orbit of the LG,
-i.e. $e_{\rm tot}$,
-$l_{\rm orb}$ and $\lambda$, do not challenge the standard $\Lambda$CDM model,
-but the model is in tension with regard to the actual values of the radial and
-tangential velocities. This might hint to a problem of the
-$\Lambda$CDM model to reproduce the observed LG.
+high resolution N-body cosmological simulation (Bolshoi) together with three
+constrained simulations to statistically study this kinematics in the
+context of the $\Lambda$CDM. The comparison of the ensembles of
+simulated pairs with the observed LG has been done with respect to the
+observed radial and tangential velocities, the reduced orbital energy
+($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
+dimensionless spin parameter, $\lambda$. Our main results are: (i) the
+preferred radial and tangential velocities for pairs in $\Lambda$CDM
+are $v_{\rm r}=-80\pm 20$\kms, $v_{\rm   t}=50\pm 10$\kms, (ii) pairs
+around that region are $3$ to $13$ times more common than pairs within
+the observational values, (iii) $15\%$ to $24\%$ of LG-like pairs in
+$\Lambda$CDM have energy and angular momentum consistent with
+observations while (iv) $9\%$ to $13\%$ of pairs in the same sample
+show similar values in the inferred dimensionless spin parameter. It
+follows that within current observational uncertainties the
+quasi-conserved quantities that characterize the orbit of
+the LG, i.e. $e_{\rm tot}$, $l_{\rm   orb}$ and $\lambda$, do not
+challenge the standard $\Lambda$CDM model, but the model is in tension
+with regard to the actual values of the radial and tangential
+velocities. This might hint to a problem of the $\Lambda$CDM model to
+reproduce the observed LG.
 
@@
 
-%\label{subsec:lg-sample}
-%\begin{figure*}
-%\begin{center}
-%\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation_BDM.pdf}
-%\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
-%\caption{\label{fig:distros} Distributions of separations and total
-%  halo mass (Milky Way and M31) in the LG-like pairs selected from the
-%  Bolshoi Simulation.} 
-%\end{center}
-%\end{figure*}
-
 \section{Simulation and Environment}
@@
 selected to have an overdensity of 200 times the critical
-density. Furthermore, we only include in the analysis haloes whose
+density. Furthermore, we only include in the analysis halos whose
 center is located outside the virial radius of any other halo.  We
@@
 
-\subsection{The LG-sample}
+\subsection{Two samples of LG-like pairs}
 Based on the BDM catalogs in the Bolshoi simulation we construct a
@@
 
-%YH: Jaime, aren't you using a Virgo constraint?
-%J.: What is a Virgo constraint in terms of Bolshoi? I am using the same constraints that we used in the old DMassembly paper.
-
+Throughout this Letter we refer to this sample as the full
+sample. 
 This sample in the Bolshoi simulation has $1923$ pairs. Additionally,
@@
 distances with respect to the Virgo cluster in the simulation.
-%Figure
-%\ref{fig:distros} show the  distribution of separations and total pair
-%mass computed from the $1923$ pairs in the Bolshoi
-%simulation. 
-Throughout this Letter we refer to this sample as the full
-sample.
 
@@
 \begin{table}
-\caption{Summary of observational constraints.}
+\caption{Summary of kinematic observational constraints.}
 \begin{center}
@@
 \begin{table}
-\caption{Summary of results from $\Lambda$CDM.}
+\caption{Summary of results from the $\Lambda$CDM Bolshoi simulation}
 \begin{center}
@@
 \vspace{1mm}
-Note: the observational uncertainties in the position vector
-correspond to a $5\%$ in each component consistent with the
-uncertainties in the distance \citep[see references
-  in][]{vanderMarel08}
-
+Note: the velocity uncertainties correspond to the minimal scale required to
+obtain robust statistics in the upper panels of Figure \ref{fig:rt}
 \label{table:2}
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1a.eps}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1b.eps}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1c.eps}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1d.eps}
 \caption{Upper panels: Histograms of the radial and tangential
@@
 \begin{tabular}{ccccc}\hline
-Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
+Physical & (\%) Pairs consistent & (\%) Pairs consistent & (\%) Pairs
 with highest & (\%) with highest\\ 
@@
 consistent with the observational constraint. The three
-pairs from the constrained realizations are also outside the region of
-the most probable values in a $\Lambda$CDM cosmology with
-tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.
+pairs from the constrained realizations are also show higher $f_{\rm
+  t}$ ratios than observations: $f_{\rm t}= 0.35, 0.45, 0.73$.
 
@@
 \label{subsection:e-l}
-
-%\begin{figure*}
-%\begin{center}
-%\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
-%\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
-%\caption{
-%Histograms of the orbital angular momentum ($l_{\rm orb}$)
-%  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
-%  calculated considering the halos as point masses. The panel
-%  distribution is the same as in Figure \ref{fig:rt}. The contour line
-%  encloses $68\%$ of the Monte Carlo generated points to estimated the
-%  uncertainties from the observational values summarized in Table
-%  1. The constraints in this plot are less restrictive than in
-%  Fig.\ref{fig:rt} due to added uncertainties on the tangential
-%  velocity ($100\%$), the square of the norm of the velocity ($40\%$)
-%  and the total mass of the two halos ($20\%$). } 
-%\label{fig:EJ}
-%\end{center}
-%\end{figure*}
 
@@
 (reduced) sample obey the 1-$\sigma$ observational constraints.
+We note that for the reduced sample the preferred region is outside the
+1-$\sigma$ observational contours.
 
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM.pdf}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM_narrow.pdf}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_2a.eps}
+\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_2b.eps}
 \caption{{\rm \label{fig:lambda} Normalized histograms of Peebles'
@@
 \begin{equation}
-\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}, 
+\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}.
 \end{equation}
-
-where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.
 
@@
 
-The expected configuration in energy and angular momentum of the LG is
-slightly inconsistent with the expectation from halo pairs in $\Lambda$CDM.
-The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value.
-Inspection of Figure \ref{fig:lambda} shows that the $\lambda$
-distribution is close to a lognormal one and thus resembles the
-distribution of the spin parameter of DM halos in simulation
-\citep{Bett2007}.
+The expected observational value for $\lambda$ is lightly inconsistent with
+the statistics derived from halo pairs in $\Lambda$CDM. The spin
+parameter of the simulated LG-like objects is skewed towards higher
+values compared with the observational estimation.  Inspection
+of Figure \ref{fig:lambda} also shows that the $\lambda$ distribution is
+close to a log-normal one and thus resembles the distribution of the
+spin parameter of DM halos in simulation \citep{Bett2007}.
 
@@
 constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
-and $-1.65$, with the latter value belonging belonging toto the sole constrained LG
-of the subsample.
+and $-1.65$, with the latter value belonging belonging toto the sole
+constrained LG of the reduced sample.
 
@@
 We have presented a comparison between the observed kinematics for the
-M31 in the galactocentric restframe and the expectations for a large
+M31 in the galactocentric rest-frame and the expectations for a large
 N-body cosmological simulation in the $\Lambda$CDM cosmology. In the
 simulation we select a large sample of pairs that fulfill isolation
-criteria for the LG. We select a subsample which obeys more
+criteria for the LG. We select a sub-sample which obeys more
 stringent observational constraints on the mass and separation of the
@@
  values $v_{\rm rad, \Lambda CDM}=-80\pm20$\kms and $v_{\rm tan,
-   \Lambda CDM}=50\pm10$ \kms, where the errorbars are estimated to
+   \Lambda CDM}=50\pm10$ \kms, where the error bars are estimated to
  include the results from the two pair samples.
@@
 a fraction of tangential to radial velocity $f_{\rm t}<0.32$ (similar
-to observations) represent $8\%$ of the same sample.
+to observations) represent $8\%$ to $12\%$ of the pairs.
 
@@
 
-Summarizing, we see we see a broad agreement in the total angular
+Summarizing, we see a broad agreement in the total angular
 momentum and energy and a marked difference with the precise balance
-between today's radial and tangential velocities  Under the
+between today's radial and tangential velocities. Under the
 approximation of conservation of the orbital angular momentum and
@@
 comparison to the initial conditions of any other pair of dark matter
-halos in the $\Lambda$CDM cosmology. This opens a new window into the
+halos in the $\Lambda$CDM cosmology. This could be related to a
+particular place of the Local Group in the Cosmic Web
+\citep{Forero-Romero09,Hoffman12}, opening a new window into the
 question of how unique, if at all, is the LG in a cosmological
@@
 
+GY  acknowledges support from MINECO through research grants AYA 2009
+13875 -C03-02 , AYA 2012 31101  and Consolider Syec CSD 2007 0050.
+Support from Comunidad. De Madrid  through ASTROMADRID research grant
+is also acknowledge
+
 The data and source code and instructions to replicate the results of
-this paper can be found as a github repository {\texttt
-  https://github.com/forero/LG\_Kinematics/}. Thanks to the IPython
+this paper can be found as a github repository {\texttt{https://github.com/forero/LG\_Kinematics/}}. Thanks to the IPython
 community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing
@@
 \bibliographystyle{apj}
-%\bibliography{references} 
-\begin{thebibliography}{28}
+\begin{thebibliography}{30}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 
+\bibitem[{{Forero-Romero} {et~al.}(2009){Forero-Romero}, {Hoffman},
+  {Gottl{\"o}ber}, {Klypin}, \& {Yepes}}]{Forero-Romero09}
+{Forero-Romero}, J.~E., {Hoffman}, Y., {Gottl{\"o}ber}, S., {Klypin}, A., \&
+  {Yepes}, G. 2009, \mnras, 396, 1815
+
 \bibitem[{{Forero-Romero} {et~al.}(2011){Forero-Romero}, {Hoffman}, {Yepes},
@@
 
-\bibitem[{{Gottl{\"o}ber} {et~al.}(2010){Gottloeber}, {Hoffman}, \&
+\bibitem[{{Gottl{\"o}ber} {et~al.}(2010){Gottl{\"o}ber}, {Hoffman}, \&
   {Yepes}}]{clues2010}
-{Gottl{\"o}ber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, 
-Proceedings of "High Performance Computing in Science and Engineering'',  
-Springer, Eds.: Wagner, S.; Steinmetz, M.; Bode, A.; M\"uller, M.M.
+{Gottl{\"o}ber}, S., {Hoffman}, Y., \& {Yepes}, G. 2010, ArXiv e-prints
+
+\bibitem[{{Hoffman} {et~al.}(2012){Hoffman}, {Metuki}, {Yepes},
+  {Gottl{\"o}ber}, {Forero-Romero}, {Libeskind}, \& {Knebe}}]{Hoffman12}
+{Hoffman}, Y., {Metuki}, O., {Yepes}, G., {Gottl{\"o}ber}, S., {Forero-Romero},
+  J.~E., {Libeskind}, N.~I., \& {Knebe}, A. 2012, \mnras, 425, 2049
 
@@
   {Turchaninov}}]{2011arXiv1109.0003R}
-{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottl{\"o}ber}, S.,
+{Riebe}, K., {Partl}, A.~M., {Enke}, H., {Forero-Romero}, J., {Gottloeber}, S.,
   {Klypin}, A., {Lemson}, G., {Prada}, F., {Primack}, J.~R., {Steinmetz}, M.,

### 2013-02-15 code/make2DHistogram.py
Changed the 2D histograms fo follow apj requierements

@@ -1,0 +1,185 @@
+from numpy import *
+import math
+import scipy 
+import numpy as np
+from pylab import *
+import matplotlib.pyplot as pl
+from matplotlib import rc
+from matplotlib.ticker import NullFormatter
+
+def ellipse(ra,rb,ang,xcenter,ycenter):
+    theta_bins = linspace(start = ang, stop = 2.0*pi, num = 100)
+    X = ra*cos(theta_bins) + xcenter
+    Y = rb*sin(theta_bins) + ycenter
+    return X,Y
+
+#this function is defined was initially written by Jessica Kirkpatrick here:
+# http://jessresearch.blogspot.com/2012/04/pretty-plots-2d-histogram-with-1d.html
+def MakeSimple2DPlot(xdata,ydata,filename=None,clues=[], clues_label='',xlims=-99, ylims =-99 , \
+                              nxbins = 50,nybins=50, bw=0, nbins=100,contours=1,sigma=1,line=1, 
+                          xlabel='A', ylabel='B', mean_obs=0, sigma_obs=0, 
+                          X_field=array([]), Y_field=array([]), Z_field=array([]), levels=array([])):
+#bw = 0 for color, = 1 for black and white
+#line = 0 for no line, =1 for line
+#sigma = 1 for display % below line, =0 for not
+#contours = 1 for display 1,2,3 sigma contours, = 0 for not.
+    
+    
+
+    # Define the x and y data
+    x = xdata
+    y = ydata
+
+# Set up default x and y limits
+    if (xlims == -99): xlims = [min(x),max(x)]
+    if (ylims == -99): ylims = [min(y),max(y)]
+    
+# Set up your x and y labels
+    xlabel = xlabel
+    ylabel = ylabel
+    mtitle = ''
+    
+    # Define the locations for the axes
+    left, width = 0.12, 0.80
+    bottom, height = 0.12, 0.80
+    bottom_h = left_h = left+width+0.02
+    
+    # Set up the geometry of the three plots
+    rect_temperature = [left, bottom, width, height] # dimensions of temp plot
+#    rect_histx = [left, bottom_h, width, 0.25] # dimensions of x-histogram
+#    rect_histy = [left_h, bottom, 0.25, height] # dimensions of y-histogram
+    
+    # Set up the size of the figure
+    fig = plt.figure(1, figsize=(9.5,9))
+    
[…]

### 2013-02-15 paper/ms.tex
updated caption

@@
 \includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1d.eps}
-\caption{Upper panels: Histograms of the radial and tangential
+\caption{Upper panels: 2D Histograms of the radial and tangential
   velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
-  panels: Histograms of the orbital angular momentum ($l_{\rm orb}$)
+  panels: 2D Histograms of the orbital angular momentum ($l_{\rm orb}$)
   and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
-  calculated considering the halos as point masses. The left (right) panels
-  correspond to the full (reduced) pair sample. The locii for the highest
-  density regions are in the $v_{\rm r}-v_{\rm t}$ plane are listed in
+  calculated considering the halos as point masses. The bar indicates
+  the number of pairs in each cell. The locii for the high
+  density regions in the $v_{\rm r}-v_{\rm t}$ plane are listed in
   Table \ref{table:2}. 

### 2013-02-15 code/make2DHistogram.py
new 2D histogram implementation

@@ -40,8 +40,8 @@ def MakeSimple2DPlot(xdata,ydata,filename=None,clues=[], clues_label='',xlims=-9
     mtitle = ''
     
     # Define the locations for the axes
-    left, width = 0.12, 0.80
-    bottom, height = 0.12, 0.80
+    left, width = 0.02, 0.95
+    bottom, height = 0.00, 0.98
     bottom_h = left_h = left+width+0.02
     
     # Set up the geometry of the three plots

### 2013-03-01 paper/reply_referee.txt
referee reply

Dear Editor,

We have appreciated the positive referee's comments and the suggestions
for improvement in the paper.

In what follows below the dashes, the referees comments are indented
and our responses are not commented.

Jaime E. Forero-Romero (on behalf of the authors)

----

	First, it would be useful if the authors were more explicit about how
	they are defining "consistency". Reading the text fairly carefully, I
	think that I was able to discern that you are using 1-sigma
	constraints on v_tan to define consistency, but this needs to be
	stated very clearly. For example, what is the quoted c.l. for the HST
	limit on V_tan listed in Table 1? This should be stated clearly in the
	table itself. In the caption of Table 3, again describe exactly how
	you are defining consistency. It would be useful to know how these
	numbers change as you very the definition (1-sigma, 2, 3?). 
	I discern from a footnote that the contours drawn in Figure 1 are
	1-sigma contours. Please state this clearly in the caption
	and/or in the text that describes the figure.

We have now included explicit mention of the 1-sigma criteria in Table
1, Table 3 and the caption of Figure 1.

   	 Second, the authors make a good case pairs that have Vr/Vtan values
	 similar to those observed are somewhat rare in the simulations and
	 suggest that this might possibly imply that the LG is not typical. It
	 should be possible to ask some basic questions about the small # of
	 systems that appear consistent with the data: does it matter? Do they
	 have outlying halo spins, concentrations, or formation times? What
	 about their larger-scale environments?

The answers to these questions are being clarified in a paper under
preparation by one of the coauthors (S. Bustamante).

The largest difficulty to study the pairs consistent with Vr/Vtan is
their low number. There are only 8 of them for the general sample and
_none_ in the reduced sample. This makes any statistical claim hard to
obtain. However we have found that these pairs prefer special large
scale environments defined by the tidal and velocity shear tensor
fields, namely they live in filaments. We are also investigating their
alignment with respect to the directions along this filaments. We will
include in that study the referee sugestions to study the
concentration and spin parameter.

We consider that this low number matters to the extent that it shows
how unlikely is to have Local Group like kinematic configurations in
LCDM, which might be relevant to explain other features in the matter
density field such as the special alignment of satellites in the Milky
Way and Andromeda or might hint towards a constraint on the expected
MW assembly history in LCDM.

### 2013-03-01 paper/reply_referee.txt
referee reply
updated version

@@
 In what follows below the dashes, the referees comments are indented
-and our responses are not commented.
+and our responses are not indented.
+
+The refereee raised concerns on Figures that look cut off. It seems to
+be a technical problem on the side of the ApJLetter submissions
+software. The raw eps figures and the pdf we produce (latex, dvips,
+ps2pdf) look complete (see:
+https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
+click on 'View Raw').  We kindly as the Editor to pay attention to
+this issue and inform us if we should make any further change to our
+figures.
+
+With best regards,
 
@@
 We have now included explicit mention of the 1-sigma criteria in Table
-1, Table 3 and the caption of Figure 1.
+1, Table 3 and the caption of Figure 1. There is also a new paragraph in Section 3.1 dicussing the results for 3-sigma uncertainties.
 
@@
 _none_ in the reduced sample. This makes any statistical claim hard to
-obtain. However we have found that these pairs prefer special large
+obtain. However, we have found that these pairs prefer special large
 scale environments defined by the tidal and velocity shear tensor
@@
 MW assembly history in LCDM.
+
+   	    Third, while this is minor in terms of content, I noticed
+	    a number of	grammatical issues. I strongly suggest a
+	    re-read for type-os and wording.
+
+We have fixed all the instances noted by the referee in the report.
+
+Concerning the Figures that look cut off, it seems a technical problem
+of the ApJLetter software. The raw eps figures and the pdf we produce
+(latex, dvips, ps2pdf) look complete (see:
+https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
+click on 'View Raw')

### 2013-03-01 paper/ms.tex
matches resubmitted version
updated version

@@
 context of the $\Lambda$CDM. The comparison of the ensembles of
-simulated pairs with the observed LG has been done with respect to the
+simulated pairs with the observed LG {\bf at the 1-$\sigma$ level in
+  the uncertainties} has been done with respect to the
 observed radial and tangential velocities, the reduced orbital energy
@@
 tidal disruption features \citep{pandas09} and the disk dominated
-morphology \citep{Kazantzidis2008} are examples on how LG studies
+morphology \citep{Kazantzidis2008} are examples of how LG studies
 are linked to the cosmological context. Detailed studies on the
@@
 collision. With this information it is possible to quantify how common
-is this kinematic configuration in a $\Lambda$CDM Universe.
+such a kinematic configuration is in a $\Lambda$CDM Universe.
 
-This Letter presents such study. We use a large volume, high resolution
+This Letter presents such a study. We use a large volume, high resolution
 dark matter only N-body simulation in the concordance $\Lambda$CDM
@@
 This Letter is structured as follows. In the next section we present
-the N-body simulations, the criteria we use to select LG-like halo
+the N-body simulations and the criteria we use to select LG-like halo
 pairs. In Section \ref{sec:results} we present the results for the
@@
 \vspace{1mm}
-Notes:\\
+Notes:
 \begin{enumerate}
@@
   components. $\sigma_{\bf x}$ represents the uncertainty on the
-  components of vector ${\bf x}$. 
+  components of vector ${\bf x}$. {\bf The uncertainties correspond to
+    1-sigma values.}  
 \item The values for the individual halo
@@
 correspond to a $5\%$ in each component consistent with the
-uncertainties in the distance \citep[see references
+1-$\sigma$ uncertainties in the distance \citep[see references
   in][]{vanderMarel08}.
@@
 \begin{center}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1a.eps}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1b.eps}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1c.eps}
-\includegraphics[keepaspectratio=true,width=0.46\textwidth]{fig_1d.eps}
+\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1a.eps}
+\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1b.eps}
+\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1c.eps}
+\includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1d.eps}
 \caption{Upper panels: 2D Histograms of the radial and tangential
@@
   and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
-  calculated considering the halos as point masses. The bar indicates
+  calculated considering the halos as point masses. {\bf Left (right) panels
+    correspond to the full (reduced) sample}. The bar indicates
   the number of pairs in each cell. The locii for the high
@@
   The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
-  the observed observational and the circles represent the positions
-  of the pairs from the constrained simulations. The contour line in
-  the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the Monte
-  Carlo generated points to estimated the uncertainties from the
-  observational values summarized in Table 1.}
+  the observations {\bf with 1-$\sigma$ uncertainties}. The contour
+  line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
+  Monte   Carlo generated points to estimated the uncertainties from
+  the   observational values summarized in Table 1. The circles
+  represent the positions of the pairs from the constrained simulations.
+}
 \label{fig:rt}
@@
 with highest & (\%) with highest\\ 
-property & with observations & with observations & likelihood in
-$\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
+property & with observations (1-$\sigma$) & with
+observations (1-$\sigma$)& likelihood in $\Lambda$CDM & likelihood in $\Lambda$CDM\\ 
  & (full sample) & (reduced sample) & (full sample) & (reduced sample)\\ \hline
@@
 \vspace{1mm}
-Note: In columns 4 and 5 the number of pairs around the preferred region
+Notes:\\
+\begin{enumerate}
+\item 
+{\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
+  uncerainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
+    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$ is
+  defined from the constrain $r_{\rm t}<0.32$ derived from the
+  1$-\sigma$ observational uncertainties in the radial and tangential
+  velocities.} 
+\item
+In columns 4 and 5 the number of pairs around the preferred region
 in $v_{\rm r}-v_{\rm t}$ in $\Lambda$CDM (as shown in
 Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
-calculated within the same observational absolute uncertainty.
+calculated within the same observational 1-$\sigma$ absolute uncertainty.
+\end{enumerate}
 \label{table:3}
 \end{table*}
+
+%Obs (full -3sigma) 127
+%Theory (full 3sigma) 130
+%Obs (reduced -3sigma) 11
+%Theory (reduced 3sigma) 17
 
@@
 Bolshoi Simulation have radial and tangential velocities notably
-different from the observational constraints.
+different from the observational constraints at the 1-$\sigma$ level.
 
@@
 
-The number of pairs compatible with LG observations are listed in
+The number of pairs compatible with LG observations {\bf at the
+  1-$\sigma$ level} are listed in
 Table \ref{table:3} (columns 2 and 3). In the same table (columns 4
@@
 configurations for the $\Lambda$CDM LG-like objects. This holds for
-the full and the reduced sample of objects.
+the full and the reduced sample of objects. {\bf This
+  conclusions are valid at 1-$\sigma$ level in the observational
+  uncertainties. Taking 3-$\sigma$ levels (i.e. uncertainties on the
+  order of $\sigma_{\rm tan}=51$\kms and $\sigma_{\rm
+    rad}=12$\kms the observational values are as likely as the
+  preferred $\Lambda$CDM values, although only $\sim 6\%$ of the pairs
+in the full and reduced samples would be consistent with observations.}
 
@@
 
-The expected observational value for $\lambda$ is lightly inconsistent with
+The expected observational value for $\lambda$ is slightly inconsistent with
 the statistics derived from halo pairs in $\Lambda$CDM. The spin
@@
 constrained LGs are $\log_{10}(\lambda_{\rm  CLUES})=-2.21$, $-1.72$
-and $-1.65$, with the latter value belonging belonging toto the sole
+and $-1.65$, with the latter value belonging to the sole
 constrained LG of the reduced sample.
@@
 velocity components, we find that none of the halos in the reduced sample
-is compatible with observations. This makes pairs with the preferred
+are compatible with observations. This makes pairs with the preferred
 $\Lambda$CDM values at least $13$ times more common than pairs
@@
 momentum and energy and a marked difference with the precise balance
-between today's radial and tangential velocities. Under the
+between today's radial and tangential velocities, {\bf i.e. head-on
+  collisions of MW-like halos are not common in $\Lambda$CDM}.
+
+Under the
 approximation of conservation of the orbital angular momentum and
@@
 comparison to the initial conditions of any other pair of dark matter
-halos in the $\Lambda$CDM cosmology. This could be related to a
-particular place of the Local Group in the Cosmic Web
-\citep{Forero-Romero09,Hoffman12}, opening a new window into the
+halos in the $\Lambda$CDM cosmology. {\bf A thorough characterization
+  of the small pair sample consistent with observations in terms
+  of halo properties such as concentration, spin and place in the cosmic web
+  \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
+  prep.)}.
+
+{\bf The results presented in this Letter open a new window into the
   question of how unique, if at all, is the LG in a cosmological
-context. This will continue to be studied within the CLUES framework.
+  context. This will continue to be studied within the CLUES project
+  framework.}
 
@@
 \section*{Acknowledgments}  
-JEF-R acknowledges financial support from the Vicerrector\'{\i}a de
-Investigaciones at Universidad de los Andes through its {\it Fondo de
+JEF-R acknowledges financial support from Universidad de los Andes through its {\it Fondo de
   Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber
-Foundation through its fellowship administered by the International
-Astronomical Union. JEF-R also acknowledges early discussions with
+Foundation through its fellowship administered by the IAU. JEF-R also acknowledges early discussions with
 Alejandro Garc\'ia that motivated this work. YH has been supported by
@@
 The data and source code and instructions to replicate the results of
-this paper can be found as a github repository {\texttt{https://github.com/forero/LG\_Kinematics/}}. Thanks to the IPython
+this paper can be found here {\texttt{https://github.com/forero/LG\_Kinematics/}}. Thanks to the IPython
 community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing

### 2013-03-05 paper/ms.tex
minor updates

@@
 \vspace{1mm}
-Note: the velocity uncertainties correspond to the minimal scale required to
-obtain robust statistics in the upper panels of Figure \ref{fig:rt}
+Note: the velocity uncertainties correspond to the minimal bin size required to
+obtain robust statistics in Figure \ref{fig:rt}
 \label{table:2}
@@
 configurations for the $\Lambda$CDM LG-like objects. This holds for
-the full and the reduced sample of objects. {\bf This
+the full and the reduced sample of objects. {\bf These
   conclusions are valid at 1-$\sigma$ level in the observational
-  uncertainties. Taking 3-$\sigma$ levels (i.e. uncertainties on the
-  order of $\sigma_{\rm tan}=51$\kms and $\sigma_{\rm
-    rad}=12$\kms the observational values are as likely as the
-  preferred $\Lambda$CDM values, although only $\sim 6\%$ of the pairs
-in the full and reduced samples would be consistent with observations.}
+  uncertainties. An agreement between both samples of simulated LGs and the observed LG is reached only at the 3-sigma level.}
 

### 2013-03-06 paper/reply_referee.txt
line break
minor updates
last version before resubmission

@@
 
-The refereee raised concerns on Figures that look cut off. It seems to
+Changes in the paper are boldfaced.
+
+The referee raised concerns on Figures that look cut off. It seems to
 be a technical problem on the side of the ApJLetter submissions
@@
 https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
-click on 'View Raw').  We kindly as the Editor to pay attention to
+click on 'View Raw').  We kindly ask the Editor to pay attention to
 this issue and inform us if we should make any further change to our
@@
 
-We have now included explicit mention of the 1-sigma criteria in Table
-1, Table 3 and the caption of Figure 1. There is also a new paragraph in Section 3.1 dicussing the results for 3-sigma uncertainties.
+We have now included explicit mention of the 1-sigma criteria in the
+abstract, Table 1, Table 3 and the caption of Figure 1. There is also
+a new paragraph in Section 3.1 dicussing the results for 3-sigma
+uncertainties.
 
@@
 scale environments defined by the tidal and velocity shear tensor
-fields, namely they live in filaments. We are also investigating their
+fields: they live in filaments. We are also investigating their
 alignment with respect to the directions along this filaments. We will
@@
 
-We have fixed all the instances noted by the referee in the report.
-
-Concerning the Figures that look cut off, it seems a technical problem
-of the ApJLetter software. The raw eps figures and the pdf we produce
-(latex, dvips, ps2pdf) look complete (see:
-https://github.com/forero/LG_Kinematics/blob/master/paper/ms.pdf ,
-click on 'View Raw')
+We have fixed all the instances noted by the referee in the report and
+made a new revesion to remove other typos.

### 2013-03-06 paper/ms.tex
last version before resubmission

@@
   the uncertainties} has been done with respect to the
-observed radial and tangential velocities, the reduced orbital energy
+radial and tangential velocities, the reduced orbital energy
 ($e_{\rm tot}$), angular momentum ($l_{\rm orb}$) and the
@@
 
-\section{Simulation and Environment}
+\section{Simulation and Pair Samples}
 \label{sec:methods}
@@
 conclusions of this paper hold in the case of halos defined by a FOF
-algorithm with a linking length 0.17 times the interparticle
+algorithm with a linking length 0.17 times the mean interparticle
 distance.
@@
 A more reduced sample from the full sample has been constructed
-so as to fit obey the observational bounds on the masses and
+so as to obey the observational bounds on the masses and
 separation of the two main halos. These amount to:   
@@
   components of vector ${\bf x}$. {\bf The uncertainties correspond to
-    1-sigma values.}  
+    1-$\sigma$ values.}  
 \item The values for the individual halo
@@
 \item The value for $\log_{10}\lambda$ is obtained in this Letter
-from a Monte Carlo simulation as described in Section
-\ref{subsection:e-l}. 
+from a Monte Carlo simulation as described in Section 3.
 \end{enumerate}
@@
 \vspace{1mm}
-Note: the velocity uncertainties correspond to the minimal bin size required to
+Note: the velocity uncertainties correspond to the minimum bin size required to
 obtain robust statistics in Figure \ref{fig:rt}
@@
 \includegraphics[keepaspectratio=true,width=0.42\textwidth]{fig_1d.eps}
-\caption{Upper panels: 2D Histograms of the radial and tangential
+\caption{Upper panels: 2D histograms of the radial and tangential
   velocities for LG-like halo pairs in the Bolshoi simulation.   Lower
-  panels: 2D Histograms of the orbital angular momentum ($l_{\rm orb}$)
+  panels: 2D histograms of the orbital angular momentum ($l_{\rm orb}$)
   and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
@@
   The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
-  the observations {\bf with 1-$\sigma$ uncertainties}. The contour
+  the {\bf 1-$\sigma$ uncertainties} in observations. Equivalently, the contour
   line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
-  Monte   Carlo generated points to estimated the uncertainties from
+  Monte   Carlo generated points from
   the   observational values summarized in Table 1. The circles
@@
 {\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
-  uncerainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
-    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$ is
+  uncertainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
+    orb}$ and $\log_{10}\lambda$. In the case of $r_{\rm t}$, is
   defined from the constrain $r_{\rm t}<0.32$ derived from the
@@
 Fig. \ref{fig:rt} and summarized in Table \ref{table:2}) are
-calculated within the same observational 1-$\sigma$ absolute uncertainty.
+calculated using the same observational 1-$\sigma$ absolute
+uncertainty on $v_{\rm r}$ and $v_{\rm t}$.
 \end{enumerate}
@@
 \end{table*}
-
-%Obs (full -3sigma) 127
-%Theory (full 3sigma) 130
-%Obs (reduced -3sigma) 11
-%Theory (reduced 3sigma) 17
 
@@
 space for $\Lambda$CDM are at least $13$ times more common than pairs
-with the observed velocities for the Local group. We highlight that
+with the observed velocities for the LG. We highlight that
 this is a lower bound given that in the reduced sample none of the
@@
 eccentric orbit of the observed LG  constitutes an unlikely
-configurations for the $\Lambda$CDM LG-like objects. This holds for
+configuration for the $\Lambda$CDM LG-like objects. This holds for
 the full and the reduced sample of objects. {\bf These
   conclusions are valid at 1-$\sigma$ level in the observational
-  uncertainties. An agreement between both samples of simulated LGs and the observed LG is reached only at the 3-sigma level.}
+  uncertainties. An equivalence in the abundance between the pairs
+  around the observational values and those around the
+  preferred $\Lambda$CDM velocities is reached only at
+  the 4-$\sigma$ level on the observational uncertainties.}
 
@@
 The two-body problem of point-like masses can serve as a proxy for the
-dynamics of the LG and be used as a tool for  studying the LG within
+dynamics of the LG and can be used as a tool for  studying the LG within
 the framework of the standard cosmological model. Within the model the
 dynamics of the LG is governed by the (center of mass)  reduced
-angular momentum $l_{\rm orb}= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
+angular momentum $l_{\rm orb}= |{{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
 reduced energy  $e_{\rm tot}=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
@@
 momentum from the observed properties in the LG listed in
-Table\ref{table:1}. Figure \ref{fig:rt} also presents the results in the
-plane $e_{\rm tot}-l_{\rm orb}$ following the   plotting conventions from
-Figure \ref{fig:rt}. \footnote{The  1-$\sigma$ contour in the lower
+Table\ref{table:1}. Lower panels in Figure \ref{fig:rt} also presents the results in the
+plane $e_{\rm tot}-l_{\rm orb}$\footnote{The  1-$\sigma$ contour in the lower
   panels of Figure \ref{fig:rt}
 does not cross the zero level, as can be naively expected from Figure
-\ref{fig:rt}. This happens for the $\approx 80\%$ level
-contour.}  As expected, the $e_{\rm tot}$  and $l_{\rm orb}$
+\ref{fig:rt}. This happens for the $\approx 80\%$ contour level.}.  As
+expected, the $e_{\rm tot}$  and $l_{\rm orb}$ 
 constraints are less restrictive compared with the radial and
@@
 consistent with the 1-$\sigma$ observational uncertainty.  However a
-reduction by a factor of $2$ in the uncertainty of the tangential
-velocity showing that the tangential velocity is below $17$\kms would
-bring these pairs and the $\Lambda$CDM expectation outside the
-1-$\sigma$ observational uncertainty.
+an hypothetical increase by a factor of $2$ in accuracy of the
+tangential velocity, showing that it is below $17$\kms, would bring
+these pairs and the $\Lambda$CDM expectation outside this uncertainty
+region.
 
@@
 between today's radial and tangential velocities, {\bf i.e. head-on
-  collisions of MW-like halos are not common in $\Lambda$CDM}.
+  collisions of MW-like halos, as described by 1-$\sigma$
+  uncertainties in observations, are not common in $\Lambda$CDM}.
 
@@
 comparison to the initial conditions of any other pair of dark matter
-halos in the $\Lambda$CDM cosmology. {\bf A thorough characterization
-  of the small pair sample consistent with observations in terms
-  of halo properties such as concentration, spin and place in the cosmic web
+halos in the $\Lambda$CDM cosmology. {\bf Investigating this
+  perspective, by a thorough characterization  of the small pair
+  sample consistent with observations in terms   of halo properties
+  such as concentration, spin and place in the cosmic web
   \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
@@
   question of how unique, if at all, is the LG in a cosmological
-  context. This will continue to be studied within the CLUES project
-  framework.}
+  context. This will continue to be studied within the framework of
+  the CLUES project.}
 

### 2013-03-07 paper/ms.tex
accepted version :)

@@
 context of the $\Lambda$CDM. The comparison of the ensembles of
-simulated pairs with the observed LG {\bf at the 1-$\sigma$ level in
-  the uncertainties} has been done with respect to the
+simulated pairs with the observed LG at the 1-$\sigma$ level in
+  the uncertainties has been done with respect to the
 radial and tangential velocities, the reduced orbital energy
@@
   components. $\sigma_{\bf x}$ represents the uncertainty on the
-  components of vector ${\bf x}$. {\bf The uncertainties correspond to
-    1-$\sigma$ values.}  
+  components of vector ${\bf x}$. The uncertainties correspond to
+    1-$\sigma$ values.
 \item The values for the individual halo
@@
   and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
-  calculated considering the halos as point masses. {\bf Left (right) panels
-    correspond to the full (reduced) sample}. The bar indicates
+  calculated considering the halos as point masses. Left (right) panels
+    correspond to the full (reduced) sample. The bar indicates
   the number of pairs in each cell. The locii for the high
@@
   The half ellipse in the $v_{\rm r}-v_{\rm t}$ plane corresponds to
-  the {\bf 1-$\sigma$ uncertainties} in observations. Equivalently, the contour
+  the 1-$\sigma$ uncertainties in observations. Equivalently, the contour
   line in   the $e_{\rm tot}-l_{\rm orb}$ plane encloses $68\%$ of the
@@
 \item 
-{\bf In columns 2 and 3 consistency is defined from 1-$\sigma$
+In columns 2 and 3 consistency is defined from 1-$\sigma$
   uncertainties for $v_{\rm r}$-$v_{\rm t}$, $e_{\rm tot}$-$l_{\rm
@@
   1$-\sigma$ observational uncertainties in the radial and tangential
-  velocities.} 
+  velocities.
 \item
@@
 
-The number of pairs compatible with LG observations {\bf at the
-  1-$\sigma$ level} are listed in
+The number of pairs compatible with LG observations  at the
+  1-$\sigma$ level are listed in
 Table \ref{table:3} (columns 2 and 3). In the same table (columns 4
@@
 configuration for the $\Lambda$CDM LG-like objects. This holds for
-the full and the reduced sample of objects. {\bf These
+the full and the reduced sample of objects. These
   conclusions are valid at 1-$\sigma$ level in the observational
@@
   preferred $\Lambda$CDM velocities is reached only at
-  the 4-$\sigma$ level on the observational uncertainties.}
+  the 4-$\sigma$ level on the observational uncertainties.
 
@@
 momentum and energy and a marked difference with the precise balance
-between today's radial and tangential velocities, {\bf i.e. head-on
+between today's radial and tangential velocities, i.e. head-on
   collisions of MW-like halos, as described by 1-$\sigma$
-  uncertainties in observations, are not common in $\Lambda$CDM}.
+  uncertainties in observations, are not common in $\Lambda$CDM.
 
@@
 comparison to the initial conditions of any other pair of dark matter
-halos in the $\Lambda$CDM cosmology. {\bf Investigating this
+halos in the $\Lambda$CDM cosmology. Investigating this
   perspective, by a thorough characterization  of the small pair
@@
   \citep{Forero-Romero09,Hoffman12} is  underway (Bustamante et al. in
-  prep.)}.
+  prep.).
 
-{\bf The results presented in this Letter open a new window into the
+The results presented in this Letter open a new window into the
   question of how unique, if at all, is the LG in a cosmological
   context. This will continue to be studied within the framework of
-  the CLUES project.}
+  the CLUES project.
 
